---
title: "**KSU paired amplicons: Beta-null deviations**"
subtitle: "16S and ITS OTU tables (EDGE samples excluded), by Grass Host"
author: "Kelli Feeser"
date: '2023-09-20'
output:
  html_document: 
    code_folding: hide
    css: test.css
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 2
  html_notebook:
    code_folding: hide
    df_print: paged
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 2
editor_options:
  chunk_output_type: inline
---

\

------------------------------------------------------------------------

# Set-up & data import -------------------------------------------------------

\

```{r LANL proxies, eval=T, include=FALSE}
# LANL proxies
Sys.setenv('HTTP_PROXY'='http://proxyout.lanl.gov:8080')
Sys.setenv('HTTPS_PROXY'='http://proxyout.lanl.gov:8080')
```

## install packages {.hidden}

```{r eval=FALSE, warning=FALSE, include=T}
install.packages("~/Downloads/densEstBayes_1.0-2.2.tar.gz", repos = NULL, type="source")
install.packages("densEstBayes")
install.packages("~/Downloads/reldist_1.7-2.tar.gz", repos = NULL, type="source")
# install.packages("~/Downloads/dotCall64_1.0-2.tar.gz", repos = NULL, type="source")
# install.packages("bipartite")
# install.packages("~/Downloads/statnet.common_4.1.4.tar", repos = NULL, type="source")

install.packages("~/Downloads/carData_3.0-3.tar", repos = NULL, type="source")
install.packages("~/Downloads/pbkrtest_0.4-7.tar", repos = NULL, type="source")
install.packages("~/Downloads/car_3.0-0.tar", repos = NULL, type="source")

install.packages("~/Downloads/rstatix_0.6.0.tar", repos = NULL, type="source")
install.packages("~/Downloads/ggpubr_0.3.0.tar", repos = NULL, type="source")
```

\

## load packages and sources {.hidden}

```{r eval=T, include=F}
library("phyloseq"); packageVersion("phyloseq") #‘1.44.0’
library("reldist"); packageVersion("reldist") #‘1.7.2’
library("vegan"); packageVersion("vegan") #‘2.6.4’
library("bipartite"); packageVersion("bipartite") #‘3.4.4’
library("ggplot2"); packageVersion("ggplot2") #‘3.4.4’
library("ggpubr"); packageVersion("ggpubr") #‘0.6.0’
library("stats"); packageVersion("stats") #‘4.3.1’
library("corrplot"); packageVersion("corrplot") #‘0.92’
library("dplyr"); packageVersion("dplyr") #‘1.1.3’
library("reshape2"); packageVersion("reshape2") #‘1.4.4’
library("ggtext"); packageVersion("ggtext") #‘0.1.2’

source("PANullDevFctsOikos.r")

#save the two R files below in your working directory
# source("MetacommunityDynamicsFctsOikos.r")
# source("PANullDevFctsOikos.r")


library(carData)
library(rstatix)
library(cowplot)
```

```{r, include=F, eval=T}
combined_metadata<-read.csv(file="../metadata/combined_metadata.csv")
```

-   ANGE (*Andropogon gerardii* / Big bluestem)\
-   SCSC (*Schizachyrium scoparium* / Little bluestem)\
-   BOER (*B. eriopoda* / Black grama)\
-   BOGR (*B. gracilis* / Blue grama)\
-   BUDA (*Bouteloua dactyloides* / Buffalograss)\

# output-top-predvar {.hidden}

## BOGR

```{r output-top-predvar-BOGR, include=FALSE, cache=T}
Bac.BOGR_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Bac.BOGR_BnullDev_abund_bysample.csv")
Fun.BOGR_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Fun.BOGR_BnullDev_abund_bysample.csv")

Bac.BOGR_dev_vals<-Bac.BOGR_dev_vals_raw
Fun.BOGR_dev_vals<-Fun.BOGR_dev_vals_raw
  
names(Bac.BOGR_dev_vals)[1] <- "X16S_ID"
names(Bac.BOGR_dev_vals)[2] <- "Bac.Observed"
names(Bac.BOGR_dev_vals)[3] <- "Bac.beta_div_metacomm_mean"
names(Bac.BOGR_dev_vals)[4] <- "Bac.abund_null_dev_stoch"
names(Bac.BOGR_dev_vals)[5] <- "Bac.abund_null_dev_det"

names(Fun.BOGR_dev_vals)[1] <- "ITS_ID"
names(Fun.BOGR_dev_vals)[2] <- "Fun.Observed"
names(Fun.BOGR_dev_vals)[3] <- "Fun.beta_div_metacomm_mean"
names(Fun.BOGR_dev_vals)[4] <- "Fun.abund_null_dev_stoch"
names(Fun.BOGR_dev_vals)[5] <- "Fun.abund_null_dev_det"


Bac.BOGR_dev_vals_dat <- merge(Bac.BOGR_dev_vals, combined_metadata, by = "X16S_ID")
Fun.BOGR_dev_vals_dat <- merge(Fun.BOGR_dev_vals, combined_metadata, by = "ITS_ID")

BOGR_dev_vals_dat <-merge(Bac.BOGR_dev_vals_dat, Fun.BOGR_dev_vals_dat)

BOGR_dev_vals_dat$Bin<-factor(BOGR_dev_vals_dat$Bin)
levels(BOGR_dev_vals_dat$Bin)
BOGR_dev_vals_dat$Bin<-factor(BOGR_dev_vals_dat$Bin, levels=c("North","North Central", "South Central", "South"))

BOGR_dev_vals_dat$Gradient<-factor(BOGR_dev_vals_dat$Gradient)
levels(BOGR_dev_vals_dat$Gradient)
BOGR_dev_vals_dat$Gradient<-factor(BOGR_dev_vals_dat$Gradient, levels=c("West","Middle", "East"))


Bac.BOGR_dev_vals_dat_longish <- data.frame(Bac.BOGR_dev_vals_dat, 
                                            "Domain" = rep("Bacteria",
                                                    nrow(Bac.BOGR_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Bac.BOGR_dev_vals_dat_longish)[2:5] <- names(Bac.BOGR_dev_vals_raw)[2:5]

Fun.BOGR_dev_vals_dat_longish <- data.frame(Fun.BOGR_dev_vals_dat, 
                                            "Domain" = rep("Fungi",
                                                    nrow(Fun.BOGR_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Fun.BOGR_dev_vals_dat_longish)[2:5] <- names(Fun.BOGR_dev_vals_raw)[2:5]

BOGR_dev_vals_dat_long <- rbind(Bac.BOGR_dev_vals_dat_longish,Fun.BOGR_dev_vals_dat_longish)

BOGR_dev_vals_dat_numeric<-subset(BOGR_dev_vals_dat, 
                                  select = -c(X16S_ID,sample_name,ITS_ID,
                                              Site, Site_Grass, Replicate,
                                              Site.Rep, Grass, Edge, SiteName,
                                              Grass_Rep, CollectionDate, MAT_degC,
                                              MAP_mm, Gradient, Bin, Grassland,
                                              julianday))

BOGR.cor.res_rho_no = cor(BOGR_dev_vals_dat_numeric[,c(24:25,28:29)],
                            BOGR_dev_vals_dat_numeric[,c(1:21)], 
                            method = "spearman")

BOGR.cor.res_rhot <- t(BOGR.cor.res_rho_no)
BOGR.cor.res_rho = data.frame("predictor" = row.names(BOGR.cor.res_rhot),
                              "bac_rho" = BOGR.cor.res_rhot[,1],
                              "fun_rho" = BOGR.cor.res_rhot[,3])


top_dev_pred_Bac.BOGR <- with(BOGR.cor.res_rho, subset(predictor, bac_rho > 0.2 |  bac_rho < -0.2))

top_dev_pred_Fun.BOGR <- with(BOGR.cor.res_rho, subset(predictor, fun_rho > 0.2 |  fun_rho < -0.2))

BOGR.cor.res_pval = cor.mtest(BOGR_dev_vals_dat_numeric, method = "spearman")$p

  # view p-val associated b/w comm and pred.var
    # BOGR.cor.res_pval["Bac.abund_null_dev_stoch","avg_SLA"]

top_dev_pred_Bac.BOGR_dat_unmelt <- subset(BOGR_dev_vals_dat, select = c("sample_name", "Bac.abund_null_dev_stoch", paste(top_dev_pred_Bac.BOGR)))
top_dev_pred_Fun.BOGR_dat_unmelt <- subset(BOGR_dev_vals_dat, select = c("sample_name", "Fun.abund_null_dev_stoch", paste(top_dev_pred_Fun.BOGR)))

top_dev_pred_Bac.BOGR_dat <- melt(top_dev_pred_Bac.BOGR_dat_unmelt, measure.vars=c(paste(top_dev_pred_Bac.BOGR)))

top_dev_pred_Fun.BOGR_dat <- melt(top_dev_pred_Fun.BOGR_dat_unmelt, measure.vars=c(paste(top_dev_pred_Fun.BOGR)))
```

\

## BOER

```{r output-top-predvar-BOER, include=FALSE, cache=T}
Bac.BOER_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Bac.BOER_BnullDev_abund_bysample.csv")
Fun.BOER_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Fun.BOER_BnullDev_abund_bysample.csv")

Bac.BOER_dev_vals<-Bac.BOER_dev_vals_raw
Fun.BOER_dev_vals<-Fun.BOER_dev_vals_raw
  
names(Bac.BOER_dev_vals)[1] <- "X16S_ID"
names(Bac.BOER_dev_vals)[2] <- "Bac.Observed"
names(Bac.BOER_dev_vals)[3] <- "Bac.beta_div_metacomm_mean"
names(Bac.BOER_dev_vals)[4] <- "Bac.abund_null_dev_stoch"
names(Bac.BOER_dev_vals)[5] <- "Bac.abund_null_dev_det"

names(Fun.BOER_dev_vals)[1] <- "ITS_ID"
names(Fun.BOER_dev_vals)[2] <- "Fun.Observed"
names(Fun.BOER_dev_vals)[3] <- "Fun.beta_div_metacomm_mean"
names(Fun.BOER_dev_vals)[4] <- "Fun.abund_null_dev_stoch"
names(Fun.BOER_dev_vals)[5] <- "Fun.abund_null_dev_det"


Bac.BOER_dev_vals_dat <- merge(Bac.BOER_dev_vals, combined_metadata, by = "X16S_ID")
Fun.BOER_dev_vals_dat <- merge(Fun.BOER_dev_vals, combined_metadata, by = "ITS_ID")

BOER_dev_vals_dat <-merge(Bac.BOER_dev_vals_dat, Fun.BOER_dev_vals_dat)

BOER_dev_vals_dat$Bin<-factor(BOER_dev_vals_dat$Bin)
levels(BOER_dev_vals_dat$Bin)
BOER_dev_vals_dat$Bin<-factor(BOER_dev_vals_dat$Bin, levels=c("North","North Central", "South Central", "South"))

BOER_dev_vals_dat$Gradient<-factor(BOER_dev_vals_dat$Gradient)
levels(BOER_dev_vals_dat$Gradient)
BOER_dev_vals_dat$Gradient<-factor(BOER_dev_vals_dat$Gradient, levels=c("West","Middle", "East"))


Bac.BOER_dev_vals_dat_longish <- data.frame(Bac.BOER_dev_vals_dat, 
                                            "Domain" = rep("Bacteria",
                                                    nrow(Bac.BOER_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Bac.BOER_dev_vals_dat_longish)[2:5] <- names(Bac.BOER_dev_vals_raw)[2:5]

Fun.BOER_dev_vals_dat_longish <- data.frame(Fun.BOER_dev_vals_dat, 
                                            "Domain" = rep("Fungi",
                                                    nrow(Fun.BOER_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Fun.BOER_dev_vals_dat_longish)[2:5] <- names(Fun.BOER_dev_vals_raw)[2:5]

BOER_dev_vals_dat_long <- rbind(Bac.BOER_dev_vals_dat_longish,Fun.BOER_dev_vals_dat_longish)

BOER_dev_vals_dat_numeric<-subset(BOER_dev_vals_dat, 
                                  select = -c(X16S_ID,sample_name,ITS_ID,
                                              Site, Site_Grass, Replicate,
                                              Site.Rep, Grass, Edge, SiteName,
                                              Grass_Rep, CollectionDate, MAT_degC,
                                              MAP_mm, Gradient, Bin, Grassland,
                                              julianday))

BOER.cor.res_rho_no = cor(BOER_dev_vals_dat_numeric[,c(24:25,28:29)],
                            BOER_dev_vals_dat_numeric[,c(1:21)], 
                            method = "spearman")

BOER.cor.res_rhot <- t(BOER.cor.res_rho_no)
BOER.cor.res_rho = data.frame("predictor" = row.names(BOER.cor.res_rhot),
                              "bac_rho" = BOER.cor.res_rhot[,1],
                              "fun_rho" = BOER.cor.res_rhot[,3])


top_dev_pred_Bac.BOER <- with(BOER.cor.res_rho, subset(predictor, bac_rho > 0.2 |  bac_rho < -0.2))

top_dev_pred_Fun.BOER <- with(BOER.cor.res_rho, subset(predictor, fun_rho > 0.2 |  fun_rho < -0.2))

BOER.cor.res_pval = cor.mtest(BOER_dev_vals_dat_numeric, method = "spearman")$p

  # view p-val associated b/w comm and pred.var
    # BOER.cor.res_pval["Bac.abund_null_dev_stoch","avg_SLA"]

top_dev_pred_Bac.BOER_dat_unmelt <- subset(BOER_dev_vals_dat, select = c("sample_name", "Bac.abund_null_dev_stoch", paste(top_dev_pred_Bac.BOER)))
top_dev_pred_Fun.BOER_dat_unmelt <- subset(BOER_dev_vals_dat, select = c("sample_name", "Fun.abund_null_dev_stoch", paste(top_dev_pred_Fun.BOER)))

top_dev_pred_Bac.BOER_dat <- melt(top_dev_pred_Bac.BOER_dat_unmelt, measure.vars=c(paste(top_dev_pred_Bac.BOER)))

top_dev_pred_Fun.BOER_dat <- melt(top_dev_pred_Fun.BOER_dat_unmelt, measure.vars=c(paste(top_dev_pred_Fun.BOER)))
```

\

## BUDA

```{r output-top-predvar-BUDA, include=FALSE, cache=T}
Bac.BUDA_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Bac.BUDA_BnullDev_abund_bysample.csv")
Fun.BUDA_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Fun.BUDA_BnullDev_abund_bysample.csv")

Bac.BUDA_dev_vals<-Bac.BUDA_dev_vals_raw
Fun.BUDA_dev_vals<-Fun.BUDA_dev_vals_raw
  
names(Bac.BUDA_dev_vals)[1] <- "X16S_ID"
names(Bac.BUDA_dev_vals)[2] <- "Bac.Observed"
names(Bac.BUDA_dev_vals)[3] <- "Bac.beta_div_metacomm_mean"
names(Bac.BUDA_dev_vals)[4] <- "Bac.abund_null_dev_stoch"
names(Bac.BUDA_dev_vals)[5] <- "Bac.abund_null_dev_det"

names(Fun.BUDA_dev_vals)[1] <- "ITS_ID"
names(Fun.BUDA_dev_vals)[2] <- "Fun.Observed"
names(Fun.BUDA_dev_vals)[3] <- "Fun.beta_div_metacomm_mean"
names(Fun.BUDA_dev_vals)[4] <- "Fun.abund_null_dev_stoch"
names(Fun.BUDA_dev_vals)[5] <- "Fun.abund_null_dev_det"


Bac.BUDA_dev_vals_dat <- merge(Bac.BUDA_dev_vals, combined_metadata, by = "X16S_ID")
Fun.BUDA_dev_vals_dat <- merge(Fun.BUDA_dev_vals, combined_metadata, by = "ITS_ID")

BUDA_dev_vals_dat <-merge(Bac.BUDA_dev_vals_dat, Fun.BUDA_dev_vals_dat)

BUDA_dev_vals_dat$Bin<-factor(BUDA_dev_vals_dat$Bin)
levels(BUDA_dev_vals_dat$Bin)
BUDA_dev_vals_dat$Bin<-factor(BUDA_dev_vals_dat$Bin, levels=c("North","North Central", "South Central", "South"))

BUDA_dev_vals_dat$Gradient<-factor(BUDA_dev_vals_dat$Gradient)
levels(BUDA_dev_vals_dat$Gradient)
BUDA_dev_vals_dat$Gradient<-factor(BUDA_dev_vals_dat$Gradient, levels=c("West","Middle", "East"))


Bac.BUDA_dev_vals_dat_longish <- data.frame(Bac.BUDA_dev_vals_dat, 
                                            "Domain" = rep("Bacteria",
                                                    nrow(Bac.BUDA_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Bac.BUDA_dev_vals_dat_longish)[2:5] <- names(Bac.BUDA_dev_vals_raw)[2:5]

Fun.BUDA_dev_vals_dat_longish <- data.frame(Fun.BUDA_dev_vals_dat, 
                                            "Domain" = rep("Fungi",
                                                    nrow(Fun.BUDA_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Fun.BUDA_dev_vals_dat_longish)[2:5] <- names(Fun.BUDA_dev_vals_raw)[2:5]

BUDA_dev_vals_dat_long <- rbind(Bac.BUDA_dev_vals_dat_longish,Fun.BUDA_dev_vals_dat_longish)

BUDA_dev_vals_dat_numeric<-subset(BUDA_dev_vals_dat, 
                                  select = -c(X16S_ID,sample_name,ITS_ID,
                                              Site, Site_Grass, Replicate,
                                              Site.Rep, Grass, Edge, SiteName,
                                              Grass_Rep, CollectionDate, MAT_degC,
                                              MAP_mm, Gradient, Bin, Grassland,
                                              julianday))

BUDA.cor.res_rho_no = cor(BUDA_dev_vals_dat_numeric[,c(24:25,28:29)],
                            BUDA_dev_vals_dat_numeric[,c(1:21)], 
                            method = "spearman")

BUDA.cor.res_rhot <- t(BUDA.cor.res_rho_no)
BUDA.cor.res_rho = data.frame("predictor" = row.names(BUDA.cor.res_rhot),
                              "bac_rho" = BUDA.cor.res_rhot[,1],
                              "fun_rho" = BUDA.cor.res_rhot[,3])


top_dev_pred_Bac.BUDA <- with(BUDA.cor.res_rho, subset(predictor, bac_rho > 0.2 |  bac_rho < -0.2))

top_dev_pred_Fun.BUDA <- with(BUDA.cor.res_rho, subset(predictor, fun_rho > 0.2 |  fun_rho < -0.2))

BUDA.cor.res_pval = cor.mtest(BUDA_dev_vals_dat_numeric, method = "spearman")$p

  # view p-val associated b/w comm and pred.var
    # BUDA.cor.res_pval["Bac.abund_null_dev_stoch","avg_SLA"]

top_dev_pred_Bac.BUDA_dat_unmelt <- subset(BUDA_dev_vals_dat, select = c("sample_name", "Bac.abund_null_dev_stoch", paste(top_dev_pred_Bac.BUDA)))
top_dev_pred_Fun.BUDA_dat_unmelt <- subset(BUDA_dev_vals_dat, select = c("sample_name", "Fun.abund_null_dev_stoch", paste(top_dev_pred_Fun.BUDA)))

top_dev_pred_Bac.BUDA_dat <- melt(top_dev_pred_Bac.BUDA_dat_unmelt, measure.vars=c(paste(top_dev_pred_Bac.BUDA)))

top_dev_pred_Fun.BUDA_dat <- melt(top_dev_pred_Fun.BUDA_dat_unmelt, measure.vars=c(paste(top_dev_pred_Fun.BUDA)))
```

\

## ANGE

```{r output-top-predvar-ANGE, include=FALSE, cache=T}
Bac.ANGE_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Bac.ANGE_BnullDev_abund_bysample.csv")
Fun.ANGE_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Fun.ANGE_BnullDev_abund_bysample.csv")

Bac.ANGE_dev_vals<-Bac.ANGE_dev_vals_raw
Fun.ANGE_dev_vals<-Fun.ANGE_dev_vals_raw
  
names(Bac.ANGE_dev_vals)[1] <- "X16S_ID"
names(Bac.ANGE_dev_vals)[2] <- "Bac.Observed"
names(Bac.ANGE_dev_vals)[3] <- "Bac.beta_div_metacomm_mean"
names(Bac.ANGE_dev_vals)[4] <- "Bac.abund_null_dev_stoch"
names(Bac.ANGE_dev_vals)[5] <- "Bac.abund_null_dev_det"

names(Fun.ANGE_dev_vals)[1] <- "ITS_ID"
names(Fun.ANGE_dev_vals)[2] <- "Fun.Observed"
names(Fun.ANGE_dev_vals)[3] <- "Fun.beta_div_metacomm_mean"
names(Fun.ANGE_dev_vals)[4] <- "Fun.abund_null_dev_stoch"
names(Fun.ANGE_dev_vals)[5] <- "Fun.abund_null_dev_det"


Bac.ANGE_dev_vals_dat <- merge(Bac.ANGE_dev_vals, combined_metadata, by = "X16S_ID")
Fun.ANGE_dev_vals_dat <- merge(Fun.ANGE_dev_vals, combined_metadata, by = "ITS_ID")

ANGE_dev_vals_dat <-merge(Bac.ANGE_dev_vals_dat, Fun.ANGE_dev_vals_dat)

ANGE_dev_vals_dat$Bin<-factor(ANGE_dev_vals_dat$Bin)
levels(ANGE_dev_vals_dat$Bin)
ANGE_dev_vals_dat$Bin<-factor(ANGE_dev_vals_dat$Bin, levels=c("North","North Central", "South Central", "South"))

ANGE_dev_vals_dat$Gradient<-factor(ANGE_dev_vals_dat$Gradient)
levels(ANGE_dev_vals_dat$Gradient)
ANGE_dev_vals_dat$Gradient<-factor(ANGE_dev_vals_dat$Gradient, levels=c("West","Middle", "East"))


Bac.ANGE_dev_vals_dat_longish <- data.frame(Bac.ANGE_dev_vals_dat, 
                                            "Domain" = rep("Bacteria",
                                                    nrow(Bac.ANGE_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Bac.ANGE_dev_vals_dat_longish)[2:5] <- names(Bac.ANGE_dev_vals_raw)[2:5]

Fun.ANGE_dev_vals_dat_longish <- data.frame(Fun.ANGE_dev_vals_dat, 
                                            "Domain" = rep("Fungi",
                                                    nrow(Fun.ANGE_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Fun.ANGE_dev_vals_dat_longish)[2:5] <- names(Fun.ANGE_dev_vals_raw)[2:5]

ANGE_dev_vals_dat_long <- rbind(Bac.ANGE_dev_vals_dat_longish,Fun.ANGE_dev_vals_dat_longish)

ANGE_dev_vals_dat_numeric<-subset(ANGE_dev_vals_dat, 
                                  select = -c(X16S_ID,sample_name,ITS_ID,
                                              Site, Site_Grass, Replicate,
                                              Site.Rep, Grass, Edge, SiteName,
                                              Grass_Rep, CollectionDate, MAT_degC,
                                              MAP_mm, Gradient, Bin, Grassland,
                                              julianday))

ANGE.cor.res_rho_no = cor(ANGE_dev_vals_dat_numeric[,c(24:25,28:29)],
                            ANGE_dev_vals_dat_numeric[,c(1:21)], 
                            method = "spearman")

ANGE.cor.res_rhot <- t(ANGE.cor.res_rho_no)
ANGE.cor.res_rho = data.frame("predictor" = row.names(ANGE.cor.res_rhot),
                              "bac_rho" = ANGE.cor.res_rhot[,1],
                              "fun_rho" = ANGE.cor.res_rhot[,3])


top_dev_pred_Bac.ANGE <- with(ANGE.cor.res_rho, subset(predictor, bac_rho > 0.2 |  bac_rho < -0.2))

top_dev_pred_Fun.ANGE <- with(ANGE.cor.res_rho, subset(predictor, fun_rho > 0.2 |  fun_rho < -0.2))

ANGE.cor.res_pval = cor.mtest(ANGE_dev_vals_dat_numeric, method = "spearman")$p

  # view p-val associated b/w comm and pred.var
    # ANGE.cor.res_pval["Bac.abund_null_dev_stoch","avg_SLA"]

top_dev_pred_Bac.ANGE_dat_unmelt <- subset(ANGE_dev_vals_dat, select = c("sample_name", "Bac.abund_null_dev_stoch", paste(top_dev_pred_Bac.ANGE)))
top_dev_pred_Fun.ANGE_dat_unmelt <- subset(ANGE_dev_vals_dat, select = c("sample_name", "Fun.abund_null_dev_stoch", paste(top_dev_pred_Fun.ANGE)))

top_dev_pred_Bac.ANGE_dat <- melt(top_dev_pred_Bac.ANGE_dat_unmelt, measure.vars=c(paste(top_dev_pred_Bac.ANGE)))

top_dev_pred_Fun.ANGE_dat <- melt(top_dev_pred_Fun.ANGE_dat_unmelt, measure.vars=c(paste(top_dev_pred_Fun.ANGE)))
```

\

## SCSC

```{r output-top-predvar-SCSC, include=FALSE, cache=T}
Bac.SCSC_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Bac.SCSC_BnullDev_abund_bysample.csv")
Fun.SCSC_dev_vals_raw<-read.csv(file="../processed_data/BetaNullDeviation/Fun.SCSC_BnullDev_abund_bysample.csv")

Bac.SCSC_dev_vals<-Bac.SCSC_dev_vals_raw
Fun.SCSC_dev_vals<-Fun.SCSC_dev_vals_raw
  
names(Bac.SCSC_dev_vals)[1] <- "X16S_ID"
names(Bac.SCSC_dev_vals)[2] <- "Bac.Observed"
names(Bac.SCSC_dev_vals)[3] <- "Bac.beta_div_metacomm_mean"
names(Bac.SCSC_dev_vals)[4] <- "Bac.abund_null_dev_stoch"
names(Bac.SCSC_dev_vals)[5] <- "Bac.abund_null_dev_det"

names(Fun.SCSC_dev_vals)[1] <- "ITS_ID"
names(Fun.SCSC_dev_vals)[2] <- "Fun.Observed"
names(Fun.SCSC_dev_vals)[3] <- "Fun.beta_div_metacomm_mean"
names(Fun.SCSC_dev_vals)[4] <- "Fun.abund_null_dev_stoch"
names(Fun.SCSC_dev_vals)[5] <- "Fun.abund_null_dev_det"


Bac.SCSC_dev_vals_dat <- merge(Bac.SCSC_dev_vals, combined_metadata, by = "X16S_ID")
Fun.SCSC_dev_vals_dat <- merge(Fun.SCSC_dev_vals, combined_metadata, by = "ITS_ID")

SCSC_dev_vals_dat <-merge(Bac.SCSC_dev_vals_dat, Fun.SCSC_dev_vals_dat)

SCSC_dev_vals_dat$Bin<-factor(SCSC_dev_vals_dat$Bin)
levels(SCSC_dev_vals_dat$Bin)
SCSC_dev_vals_dat$Bin<-factor(SCSC_dev_vals_dat$Bin, levels=c("North","North Central", "South Central", "South"))

SCSC_dev_vals_dat$Gradient<-factor(SCSC_dev_vals_dat$Gradient)
levels(SCSC_dev_vals_dat$Gradient)
SCSC_dev_vals_dat$Gradient<-factor(SCSC_dev_vals_dat$Gradient, levels=c("West","Middle", "East"))


Bac.SCSC_dev_vals_dat_longish <- data.frame(Bac.SCSC_dev_vals_dat, 
                                            "Domain" = rep("Bacteria",
                                                    nrow(Bac.SCSC_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Bac.SCSC_dev_vals_dat_longish)[2:5] <- names(Bac.SCSC_dev_vals_raw)[2:5]

Fun.SCSC_dev_vals_dat_longish <- data.frame(Fun.SCSC_dev_vals_dat, 
                                            "Domain" = rep("Fungi",
                                                    nrow(Fun.SCSC_dev_vals_dat)),
                                            stringsAsFactors = T)
names(Fun.SCSC_dev_vals_dat_longish)[2:5] <- names(Fun.SCSC_dev_vals_raw)[2:5]

SCSC_dev_vals_dat_long <- rbind(Bac.SCSC_dev_vals_dat_longish,Fun.SCSC_dev_vals_dat_longish)

SCSC_dev_vals_dat_numeric<-subset(SCSC_dev_vals_dat, 
                                  select = -c(X16S_ID,sample_name,ITS_ID,
                                              Site, Site_Grass, Replicate,
                                              Site.Rep, Grass, Edge, SiteName,
                                              Grass_Rep, CollectionDate, MAT_degC,
                                              MAP_mm, Gradient, Bin, Grassland,
                                              julianday))

SCSC.cor.res_rho_no = cor(SCSC_dev_vals_dat_numeric[,c(24:25,28:29)],
                            SCSC_dev_vals_dat_numeric[,c(1:21)], 
                            method = "spearman")

SCSC.cor.res_rhot <- t(SCSC.cor.res_rho_no)
SCSC.cor.res_rho = data.frame("predictor" = row.names(SCSC.cor.res_rhot),
                              "bac_rho" = SCSC.cor.res_rhot[,1],
                              "fun_rho" = SCSC.cor.res_rhot[,3])


top_dev_pred_Bac.SCSC <- with(SCSC.cor.res_rho, subset(predictor, bac_rho > 0.2 |  bac_rho < -0.2))

top_dev_pred_Fun.SCSC <- with(SCSC.cor.res_rho, subset(predictor, fun_rho > 0.2 |  fun_rho < -0.2))

SCSC.cor.res_pval = cor.mtest(SCSC_dev_vals_dat_numeric, method = "spearman")$p

  # view p-val associated b/w comm and pred.var
    # SCSC.cor.res_pval["Bac.abund_null_dev_stoch","avg_SLA"]

top_dev_pred_Bac.SCSC_dat_unmelt <- subset(SCSC_dev_vals_dat, select = c("sample_name", "Bac.abund_null_dev_stoch", paste(top_dev_pred_Bac.SCSC)))
top_dev_pred_Fun.SCSC_dat_unmelt <- subset(SCSC_dev_vals_dat, select = c("sample_name", "Fun.abund_null_dev_stoch", paste(top_dev_pred_Fun.SCSC)))

top_dev_pred_Bac.SCSC_dat <- melt(top_dev_pred_Bac.SCSC_dat_unmelt, measure.vars=c(paste(top_dev_pred_Bac.SCSC)))

top_dev_pred_Fun.SCSC_dat <- melt(top_dev_pred_Fun.SCSC_dat_unmelt, measure.vars=c(paste(top_dev_pred_Fun.SCSC)))
```

\
\

## find intersect of pred var {.hidden}

```{r, include=F, eval=F}
intersect(top_dev_pred_Bac.BOER, top_dev_pred_Bac.BOGR)

intersect(intersect(top_dev_pred_Bac.BOER, top_dev_pred_Bac.BOGR), top_dev_pred_Bac.BUDA)

intersect(intersect(top_dev_pred_Bac.BOER, top_dev_pred_Bac.BOGR), top_dev_pred_Bac.SCSC)

intersect(intersect(intersect(top_dev_pred_Bac.BOER, top_dev_pred_Bac.BOGR), top_dev_pred_Bac.SCSC), top_dev_pred_Bac.BUDA)

Bac.predvar_all <- intersect(top_dev_pred_Bac.ANGE, top_dev_pred_Bac.BOGR, top_dev_pred_Bac.BOER, top_dev_pred_Bac.BUDA, top_dev_pred_Bac.SCSC)
```

## combine top pred var {.hidden}

```{r combine-top-predvar, include=FALSE, cache=TRUE}

#combine all top_dev_pred_Bac.*_dat
top_dev_pred_Bac.BOGR_dat$Grass<-"BOGR"
top_dev_pred_Bac.BOER_dat$Grass<-"BOER"
top_dev_pred_Bac.BUDA_dat$Grass<-"BUDA"
top_dev_pred_Bac.ANGE_dat$Grass<-"ANGE"
top_dev_pred_Bac.SCSC_dat$Grass<-"SCSC"

top_dev_pred_Bac_dat <- rbind(top_dev_pred_Bac.ANGE_dat, top_dev_pred_Bac.BOGR_dat,
                                   top_dev_pred_Bac.BOER_dat, top_dev_pred_Bac.BUDA_dat,
                                   top_dev_pred_Bac.SCSC_dat)

#combine all top_dev_pred_Fun.*_dat
top_dev_pred_Fun.BOGR_dat$Grass<-"BOGR"
top_dev_pred_Fun.BOER_dat$Grass<-"BOER"
top_dev_pred_Fun.BUDA_dat$Grass<-"BUDA"
top_dev_pred_Fun.ANGE_dat$Grass<-"ANGE"
top_dev_pred_Fun.SCSC_dat$Grass<-"SCSC"

top_dev_pred_Fun_dat <- rbind(top_dev_pred_Fun.ANGE_dat, top_dev_pred_Fun.BOGR_dat,
                                   top_dev_pred_Fun.BOER_dat, top_dev_pred_Fun.BUDA_dat,
                                   top_dev_pred_Fun.SCSC_dat)

SCSC.cor.res_rho
SCSC.cor.res_pval
```

## combine rho mats {.hidden}

```{r combine-rho-mats, include=FALSE, cache=TRUE}
#combine all *.cor.res_rho
BOGR.cor.res_rho$Grass<-"BOGR"
BOER.cor.res_rho$Grass<-"BOER"
BUDA.cor.res_rho$Grass<-"BUDA"
ANGE.cor.res_rho$Grass<-"ANGE"
SCSC.cor.res_rho$Grass<-"SCSC"

all.cor.res_rho <- rbind(ANGE.cor.res_rho, BOGR.cor.res_rho,
                                   BOER.cor.res_rho, BUDA.cor.res_rho,
                                   SCSC.cor.res_rho)
```

## combine pvals {.hidden}

```{r combine-rho-pval-mats, include=FALSE, cache=TRUE}
#combine all *.cor.res_pval
# BOGR.cor.res_pval$Grass<-"BOGR"
# BOER.cor.res_pval$Grass<-"BOER"
# BUDA.cor.res_pval$Grass<-"BUDA"
# ANGE.cor.res_pval$Grass<-"ANGE"
# SCSC.cor.res_pval$Grass<-"SCSC"

# all.cor.res_pval <- rbind(ANGE.cor.res_pval, BOGR.cor.res_pval,
#                                   BOER.cor.res_pval, BUDA.cor.res_pval,
#                                   SCSC.cor.res_pval)
```

# plot top pred var {.tabset .tabset-pills}




```{r, include=F}
hist(top_dev_pred_Bac_dat$Bac.abund_null_dev_stoch, xlim = c(-0.3,0.6), freq=F)
hist(top_dev_pred_Fun_dat$Fun.abund_null_dev_stoch, xlim = c(-0.3,0.6), freq=F)
hist(combined_metadata$herbivory_perc, freq=F)
```

```{r, include=F}
qqplot(top_dev_pred_Bac_dat$Bac.abund_null_dev_stoch, qchisq(ppoints(top_dev_pred_Bac_dat$Bac.abund_null_dev_stoch), df = 3)); abline(0, 1, col = 2, lty = 2)

```


\

## **Bacteria** {.unnumbered}

𝛃~[NULL]{.smallcaps}~ deviation values:\
- min: `r round(min(top_dev_pred_Bac_dat$Bac.abund_null_dev_stoch),2)`\
- max: `r round(max(top_dev_pred_Bac_dat$Bac.abund_null_dev_stoch),2)`\
- median: `r round(median(top_dev_pred_Bac_dat$Bac.abund_null_dev_stoch),2)`\
- mean: `r round(mean(top_dev_pred_Bac_dat$Bac.abund_null_dev_stoch),2)`

```{r, fig.height=7, fig.width=9, echo=F, message=F}
top_dev_pred_Bac_dat2 <- subset(top_dev_pred_Bac_dat, 
                                !(variable %in% c("perc_moisture","GWC",
                                                  "coll_GDD","ppt2015",
                                                  "ppt30yr","GDD2015",
                                                  "GDD30yr")))
top_dev_pred_Bac_dat2$variable<-factor(top_dev_pred_Bac_dat2$variable)

#top_dev_pred_Bac_dat2[top_dev_pred_Bac_dat2$variable=="Longitude"]

top_dev_pred_Bac_dat3 <- top_dev_pred_Bac_dat2
#top_dev_pred_Bac_dat3$variable<-as.factor(top_dev_pred_Bac_dat3$variable)
levels(top_dev_pred_Bac_dat3$variable) <- c(levels(top_dev_pred_Bac_dat3$variable), 'Longitude')

top_dev_pred_Bac_dat3$variable <- factor(top_dev_pred_Bac_dat2$variable, 
                                         levels = c("Latitude", "Longitude","Elevation_m",
                                            "SOM", "pH","soil_moisture","phos",
                                            "ammonium","nitrate",
                                            "GDD3yr","ppt3yr",
                                            "herbivory_perc","avg_SLA","avg_SRL"))

ggplot(data=top_dev_pred_Bac_dat3, aes(x=Bac.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_grid(variable ~ Grass, scales = "free_y", drop = F)+
  labs(col="Metadata factor \ncorrelated with \ndeviation values",
       title = "Impact of all significant metadata* on &beta;<sub><i>NULL</i></sub> deviation values<span> by Grass Host</span>",
       subtitle = "* rho > | 0.2 |") +
  guides(col=guide_legend(ncol=1))+
  stat_smooth(method = "loess")+
  theme_bw() +
  theme(
    plot.background = element_blank(),
    plot.title = element_markdown(lineheight = 1.1),
    panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    strip.text.y = element_text(color="black", size=10, angle = 0),
    strip.text.x = element_text(color="black", size=12),
    legend.position="right",
    axis.title.x = element_text(color="black", size=10),
    axis.title.y = element_text(color="black", size=10),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

# summary(lm(Bac.abund_null_dev_stoch~avg_SLA, data=top_dev_pred_Bac.BOGR_dat_unmelt))
# summary(lm(Bac.abund_null_dev_stoch~avg_SRL, data=top_dev_pred_Bac.BOGR_dat_unmelt))
```

## **Fungi** {.unnumbered}

𝛃~[NULL]{.smallcaps}~ deviation values:\
- min: `r round(min(top_dev_pred_Fun_dat$Fun.abund_null_dev_stoch),2)`\
- max: `r round(max(top_dev_pred_Fun_dat$Fun.abund_null_dev_stoch),2)`\
- median: `r round(median(top_dev_pred_Fun_dat$Fun.abund_null_dev_stoch),2)`\
- mean: `r round(mean(top_dev_pred_Fun_dat$Fun.abund_null_dev_stoch),2)`

```{r, fig.height=7, fig.width=9, echo=F, message=F}
top_dev_pred_Fun_dat2 <- subset(top_dev_pred_Fun_dat, 
                                !(variable %in% c("perc_moisture","GWC",
                                                  "coll_GDD","ppt2015",
                                                  "ppt30yr","GDD2015",
                                                  "GDD30yr")))
top_dev_pred_Fun_dat2$variable<-factor(top_dev_pred_Fun_dat2$variable)

#top_dev_pred_Fun_dat2[top_dev_pred_Fun_dat2$variable=="Longitude"]

top_dev_pred_Fun_dat3 <- top_dev_pred_Fun_dat2
#top_dev_pred_Fun_dat3$variable<-as.factor(top_dev_pred_Fun_dat3$variable)
levels(top_dev_pred_Fun_dat3$variable) <- c(levels(top_dev_pred_Fun_dat3$variable), 'Longitude')

top_dev_pred_Fun_dat3$variable <- factor(top_dev_pred_Fun_dat2$variable, 
                                         levels = c("Latitude", "Longitude","Elevation_m",
                                            "SOM", "pH","soil_moisture","phos",
                                            "ammonium","nitrate",
                                            "GDD3yr","ppt3yr",
                                            "herbivory_perc","avg_SLA","avg_SRL"))

ggplot(data=top_dev_pred_Fun_dat3, aes(x=Fun.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_grid(variable ~ Grass, scales = "free_y")+
  labs(col="Metadata factor \ncorrelated with \ndeviation values",
       title = "Impact of all significant metadata* on &beta;<sub><i>NULL</i></sub> deviation values<span>by Grass Host</span>",
       subtitle = "* rho > | 0.2 |") +
  guides(col=guide_legend(ncol=1))+
  stat_smooth(method = "loess")+
  theme_bw() +
  theme(
    plot.background = element_blank(),
     plot.title = element_markdown(lineheight = 1.1),
    panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    strip.text.y = element_text(color="black", size=10, angle = 0),
    strip.text.x = element_text(color="black", size=12),
    legend.position="right",
    axis.title.x = element_text(color="black", size=10),
    axis.title.y = element_text(color="black", size=10),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

# summary(lm(Fun.abund_null_dev_stoch~avg_SLA, data=top_dev_pred_Fun.BOGR_dat_unmelt))
# summary(lm(Fun.abund_null_dev_stoch~avg_SRL, data=top_dev_pred_Fun.BOGR_dat_unmelt))
```

#  {.unnumbered}

*Climate Data:* \
- GDD = growing degree days (determined over 3- and 30-year windows and 2015) \
- ppt = mean annual precipitation (determined a 30-year) \
\

*Plant Traits:* \
- avg_SLA: specific leaf area \
- avg_SRL: specific root length \
  - Large values of specific leaf area (SLA) and specific root length (SRL) reflect large plant investment in resource acquisition; small values indicate investment in plant tissue longevity \
- herbivory_perc: average site level damage estimate for herbivory, averaged over all species and individuals at the site; could be used to indicate herbivory pressure at the site level \
  - ranges from 1:17 \


# plot top pred var by grass {.tabset .tabset-pills .unnumbered}

```{r}
top_dev_pred_Bac_dat2 <- subset(top_dev_pred_Bac_dat, 
                                !(variable %in% c("perc_moisture","GWC",
                                                  "coll_GDD","ppt2015",
                                                  "ppt30yr","GDD2015",
                                                  "GDD30yr")))
top_dev_pred_Bac_dat2$variable<-factor(top_dev_pred_Bac_dat2$variable)

#top_dev_pred_Bac_dat2[top_dev_pred_Bac_dat2$variable=="Longitude"]

top_dev_pred_Bac_dat3 <- top_dev_pred_Bac_dat2
#top_dev_pred_Bac_dat3$variable<-as.factor(top_dev_pred_Bac_dat3$variable)
levels(top_dev_pred_Bac_dat3$variable) <- c(levels(top_dev_pred_Bac_dat3$variable), 'Longitude')

top_dev_pred_Bac_dat3$variable <- factor(top_dev_pred_Bac_dat2$variable, 
                                         levels = c("Latitude", "Longitude","Elevation_m",
                                            "SOM", "pH","soil_moisture","phos",
                                            "ammonium","nitrate",
                                            "GDD3yr","ppt3yr",
                                            "herbivory_perc","avg_SLA","avg_SRL"))
```

## Bacteria: BOER {.unnumbered}
```{r, message=F, echo=F}
ggplot(data=subset(top_dev_pred_Bac_dat3, Grass=="BOER"), aes(y=Bac.abund_null_dev_stoch, x=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_x")+
  stat_smooth(method = "loess")+
  theme_bw() +
  theme(
    plot.background = element_blank(),
     plot.title = element_markdown(lineheight = 1.1),
    panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=10),
    axis.text.y = element_text(color="black", size=10),
    strip.text.y = element_text(color="black", size=10, angle = 0),
    strip.text.x = element_text(color="black", size=12),
    legend.position="right",
    axis.title.x = element_text(color="black", size=10),
    axis.title.y = element_text(color="black", size=10),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

```

```{r}
subset(BOER.cor.res_rho, predictor %in% subset(top_dev_pred_Bac_dat3, Grass=="BOER")$variable)[,1:2]
```


## Fungi: BOER {.unnumbered}
```{r, message=F, echo=F}
ggplot(data=subset(top_dev_pred_Fun_dat3, Grass=="BOER"), aes(x=Fun.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_y")+
  stat_smooth(method = "loess")
```

```{r}
subset(BOER.cor.res_rho, predictor %in% subset(top_dev_pred_Fun_dat3, Grass=="BOER")$variable)[,c(1,3)]
```


## Bacteria: BOGR {.unnumbered}
```{r, message=F, echo=F}
ggplot(data=subset(top_dev_pred_Bac_dat3, Grass=="BOGR"), aes(x=Bac.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_y")+
  stat_smooth(method = "loess")
```

```{r}
subset(BOGR.cor.res_rho, predictor %in% subset(top_dev_pred_Bac_dat3, Grass=="BOGR")$variable)[,1:2]
```


## Fungi: BOGR {.unnumbered}
```{r, message=F, echo=F}
ggplot(data=subset(top_dev_pred_Fun_dat3, Grass=="BOGR"), aes(x=Fun.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_y")+
  stat_smooth(method = "loess")
```

```{r}
subset(BOGR.cor.res_rho, predictor %in% subset(top_dev_pred_Bac_dat3, Grass=="BOGR")$variable)[,c(1,3)]
```

## Bacteria: BUDA {.unnumbered}
```{r, message=F, echo=F, eval=T, cache=T}
ggplot(data=subset(top_dev_pred_Bac_dat3, Grass=="BUDA"), aes(x=Bac.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_y")+
  stat_smooth(method = "loess")
```

```{r}
subset(BUDA.cor.res_rho, predictor %in% subset(top_dev_pred_Bac_dat3, Grass=="BUDA")$variable)[,1:2]
```


## Fungi: BUDA {.unnumbered}
```{r, message=F, echo=F, eval=T, cache=T}
ggplot(data=subset(top_dev_pred_Fun_dat3, Grass=="BUDA"), aes(x=Fun.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_y")+
  stat_smooth(method = "loess")
```

```{r}
subset(BUDA.cor.res_rho, predictor %in% subset(top_dev_pred_Bac_dat3, Grass=="BUDA")$variable)[,c(1,3)]
```


## Bacteria: ANGE {.unnumbered}
```{r, message=F, echo=F, eval=T, cache=T}
ggplot(data=subset(top_dev_pred_Bac_dat3, Grass=="ANGE"), aes(x=Bac.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_y")+
  stat_smooth(method = "loess")
```

```{r}
subset(ANGE.cor.res_rho, predictor %in% subset(top_dev_pred_Bac_dat3, Grass=="ANGE")$variable)[,1:2]
```


## Fungi: ANGE {.unnumbered}
```{r, message=F, echo=F, eval=T, cache=T}
ggplot(data=subset(top_dev_pred_Fun_dat3, Grass=="ANGE"), aes(x=Fun.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_y")+
  stat_smooth(method = "loess")
```

```{r}
subset(ANGE.cor.res_rho, predictor %in% subset(top_dev_pred_Bac_dat3, Grass=="ANGE")$variable)[,c(1,3)]
```

## Bacteria: SCSC {.unnumbered}
```{r, message=F, echo=F, eval=T, cache=T}
ggplot(data=subset(top_dev_pred_Bac_dat3, Grass=="SCSC"), aes(x=Bac.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_y")+
  stat_smooth(method = "loess")
```

```{r}
subset(SCSC.cor.res_rho, predictor %in% subset(top_dev_pred_Bac_dat3, Grass=="SCSC")$variable)[,1:2]
```


## Fungi: SCSC {.unnumbered}
```{r, message=F, echo=F, eval=T, cache=T}
ggplot(data=subset(top_dev_pred_Fun_dat3, Grass=="SCSC"), aes(x=Fun.abund_null_dev_stoch, y=value, col=variable)) +
  geom_point() +
  facet_wrap(.~variable, scales = "free_y")+
  stat_smooth(method = "loess")
```

```{r}
subset(SCSC.cor.res_rho, predictor %in% subset(top_dev_pred_Bac_dat3, Grass=="SCSC")$variable)[,c(1,3)]
```
\


```{r}
cor(combined_metadata$ammonium,combined_metadata$phos, use="pairwise.complete.obs")
```

```{r}
cor(combined_metadata$ammonium,combined_metadata$nitrate, use="pairwise.complete.obs")
```

```{r}
cor(combined_metadata$soil_moisture,combined_metadata$GDD3yr, use="pairwise.complete.obs")
cor(combined_metadata$soil_moisture,combined_metadata$phos, use="pairwise.complete.obs")
cor(combined_metadata$soil_moisture,combined_metadata$ammonium, use="pairwise.complete.obs")

cor(combined_metadata$ammonium,combined_metadata$GDD3yr, use="pairwise.complete.obs")

```

```{r}
cor(combined_metadata$soil_moisture,combined_metadata$pH, use="pairwise.complete.obs")
```


```{r, include=FALSE}
Bac.by.sample.dat<-read.csv(file="../processed_data/BetaNullDeviation/Bac_BnullDev_abund_bysample.csv", stringsAsFactors = T)
Fun.by.sample.dat<-read.csv(file="../processed_data/BetaNullDeviation/Fun_BnullDev_abund_bysample.csv", stringsAsFactors = T)
combined_metadata_E<-subset(combined_metadata, Edge=="no")

Bac.by.sample.dat_cmeta<-merge(Bac.by.sample.dat, combined_metadata_E, by = "X16S_ID")
Fun.by.sample.dat_cmeta <- merge(Fun.by.sample.dat, combined_metadata_E, by = "ITS_ID")


dom_joiner <- subset(combined_metadata_E, select = c(X16S_ID, ITS_ID, sample_name))

Bac.by.sample.dat_wname<-merge(Bac.by.sample.dat, dom_joiner, by = "X16S_ID")
Fun.by.sample.dat_wname<-merge(Fun.by.sample.dat, dom_joiner, by = "ITS_ID", all.y=T)

Bac.by.sample.dat_wname2 <- subset(Bac.by.sample.dat_wname, select = -c(ITS_ID, X16S_ID))
#row.names(Bac.by.sample.dat_wname2) <- Bac.by.sample.dat_wname2$sample_name
Fun.by.sample.dat_wname2 <- subset(Fun.by.sample.dat_wname, select = -c(ITS_ID, X16S_ID))

AllD.by.sample.dat_rbind<-rbind(Bac.by.sample.dat_wname2, Fun.by.sample.dat_wname2)

AllD.by.sample.dat_cmetaA<-merge(AllD.by.sample.dat_rbind, combined_metadata_E, 
                                by = "sample_name", all.x=T)
AllD.by.sample.dat_cmeta <- subset(AllD.by.sample.dat_cmetaA, 
                                   select = -c(Grass.x, ITS_ID, X16S_ID, Edge,
                                               MAP_mm, MAT_degC,perc_moisture,GWC,
                                               coll_GDD,ppt2015,ppt30yr,
                                               GDD2015,GDD30yr, julianday, 
                                               CollectionDate))
names(AllD.by.sample.dat_cmeta)[11] <- "Grass"

AllD.by.sample.dat_cmeta$Domain <- factor(AllD.by.sample.dat_cmeta$Domain)
levels(AllD.by.sample.dat_cmeta$Domain)

#AllD.by.sample.dat_cmeta_long <- melt(AllD.by.sample.dat_cmeta, id.vars = c(Domain),
 #                                     measure.vars = c())

# AllD.by.sample.dat_wname<-merge(Bac.by.sample.dat_wname, Fun.by.sample.dat, 
#                                 by = "ITS_ID", all=T,
#                                 suffixes = c("Bac","Fun"))
# AllD.by.sample.dat_wname2 <- subset(AllD.by.sample.dat_wname, select = -c(GrassFun, ITS_ID, X16S_ID))
# AllD.by.sample.dat_wname_long <- melt(AllD.by.sample.dat_wname2, id.vars = c(DomainBac, DomainFun),
#                                       measure.vars = c())
# AllD.by.sample.dat<-merge(Bac.by.sample.dat, Fun.by.sample.dat, by = "sample_name", all = T,
#                           suffixes = c("Bac","Fun"))
```


```{r, message=F}
# round(IQR(subset(AllD.by.sample.dat_cmeta, Domain=="Bacteria")$abund_null_dev_stoch),3)
# round(IQR(subset(AllD.by.sample.dat_cmeta, Domain=="Fungi")$abund_null_dev_stoch),3)

soil_moisture_dev_dom<-ggplot(data=subset(AllD.by.sample.dat_cmeta, !is.na(Domain)), aes(x=soil_moisture, y=abund_null_dev_stoch, col=Domain)) +
  geom_point(alpha=0.85)+
  scale_color_manual(values = c("#006e90", "#f18f01"))+
  geom_hline(yintercept=0, col="gray40", lty=2, show.legend = F)+
  labs(x="Soil Moisture", y="Deviation Value")+
  # stat_smooth(inherit.aes = F, method = "lm", level=0.8, 
  #             aes(x=soil_moisture, y=abund_null_dev_stoch), col="black", linewidth=2, alpha=0.85)+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.85, linewidth=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="right",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=16),
    legend.text  = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

pH_dev_dom<-ggplot(data=subset(AllD.by.sample.dat_cmeta, !is.na(Domain)), aes(x=pH, y=abund_null_dev_stoch, col=Domain)) +
  geom_point(alpha=0.85)+
  scale_color_manual(values = c("#006e90", "#f18f01"))+
  geom_hline(yintercept=0, col="gray40", lty=2, show.legend = F)+
  labs(x="pH", y="")+
  # stat_smooth(inherit.aes = F, method = "lm", level=0.8, 
  #             aes(x=soil_moisture, y=abund_null_dev_stoch), col="black", linewidth=2, alpha=0.85)+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.85, linewidth=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

nitrate_dev_dom<-ggplot(data=subset(AllD.by.sample.dat_cmeta, !is.na(Domain)), aes(x=nitrate, y=abund_null_dev_stoch, col=Domain)) +
  geom_point(alpha=0.85)+
  scale_color_manual(values = c("#006e90", "#f18f01"))+
  geom_hline(yintercept=0, col="gray40", lty=2, show.legend = F)+
  labs(x="Nitrate", y="")+
  # stat_smooth(inherit.aes = F, method = "lm", level=0.8, 
  #             aes(x=soil_moisture, y=abund_null_dev_stoch), col="black", linewidth=2, alpha=0.85)+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.85, linewidth=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

avg_SLA_dev_dom<-ggplot(data=subset(AllD.by.sample.dat_cmeta, !is.na(Domain)), aes(x=avg_SLA, y=abund_null_dev_stoch, col=Domain)) +
  geom_point(alpha=0.85)+
  scale_color_manual(values = c("#006e90", "#f18f01"))+
  geom_hline(yintercept=0, col="gray40", lty=2, show.legend = F)+
  labs(x="Specific Leaf Area", y="Deviation Value")+
  # stat_smooth(inherit.aes = F, method = "lm", level=0.8, 
  #             aes(x=soil_moisture, y=abund_null_dev_stoch), col="black", linewidth=2, alpha=0.85)+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.85, linewidth=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

avg_SRL_dev_dom<-ggplot(data=subset(AllD.by.sample.dat_cmeta, !is.na(Domain)), aes(x=avg_SRL, y=abund_null_dev_stoch, col=Domain)) +
  geom_point(alpha=0.85)+
  scale_color_manual(values = c("#006e90", "#f18f01"))+
  geom_hline(yintercept=0, col="gray40", lty=2, show.legend = F)+
  labs(x="Specific Root Length", y="")+
  # stat_smooth(inherit.aes = F, method = "lm", level=0.8, 
  #             aes(x=soil_moisture, y=abund_null_dev_stoch), col="black", linewidth=2, alpha=0.85)+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.85, linewidth=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

dev_dom_leg<-get_legend(soil_moisture_dev_dom)
soil_moisture_dev_dom_noleg<-soil_moisture_dev_dom+theme(legend.position = "none")
```


```{r, message=F}
plot_grid(soil_moisture_dev_dom_noleg, pH_dev_dom, nitrate_dev_dom, avg_SLA_dev_dom, avg_SRL_dev_dom,dev_dom_leg)
```



```{r}
ggplot(data=subset(AllD.by.sample.dat_cmeta, !is.na(Domain)), aes(x=soil_moisture, y=abund_null_dev_stoch, col=Domain)) +
  geom_point(alpha=0.85)+
  scale_color_manual(values = c("#006e90", "#f18f01"))+
  geom_hline(yintercept=0, col="gray40", lty=2, show.legend = F)+
  labs(x="Soil Moisture", y="Deviation Value")+
  # stat_smooth(inherit.aes = F, method = "lm", level=0.8, 
  #             aes(x=soil_moisture, y=abund_null_dev_stoch), col="black", linewidth=2, alpha=0.85)+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.85, linewidth=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )
```



# soil moisture {.tabset .tabset-pills .unnumbered}

## **Bacteria** {.unnumbered}
```{r, message=F}
b.smg<-ggplot(data=Bac.by.sample.dat_cmeta, aes(x=soil_moisture, y=abund_null_dev_stoch, col=Grass.x)) +
  geom_point(alpha=0.5)+
  geom_hline(yintercept=0, col="gray40", lty=2)+
  scale_color_manual(values = bac_grass_cols)+
  scale_y_continuous(limits = c(-0.4,.6))+
  labs(y="Deviation values", x="Soil Moisture", col="Grass Host")+
  # stat_smooth(inherit.aes = F, method = "lm", level=0.8, 
  #             aes(x=soil_moisture, y=abund_null_dev_stoch), col="black", linewidth=2, alpha=0.85)+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.85, linewidth=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~soil_moisture, data = Bac.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~soil_moisture, data = Bac.by.sample.dat_cmeta)))$coefficients[2,1],2)` \
p-value: `r (summary(lm(abund_null_dev_stoch~soil_moisture, data = Bac.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

## **Fungi** {.unnumbered}
```{r, messgae=F}
f.smg<-ggplot(data=Fun.by.sample.dat_cmeta, aes(x=soil_moisture, y=abund_null_dev_stoch, col=Grass.x)) +
  geom_point(alpha=0.5)+
  geom_hline(yintercept=0, col="gray40", lty=2)+
  scale_y_continuous(limits = c(-0.4,.6))+
  scale_color_manual(values = bac_grass_cols)+
  labs(y="Deviation values", x="Soil Moisture", col="Grass Host")+
  # stat_smooth(inherit.aes = F, method = "lm", level=0.8, 
  #             aes(x=soil_moisture, y=abund_null_dev_stoch), col="black", linewidth=2, alpha=0.85)+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.85, linewidth=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="right",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~soil_moisture, data = Fun.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~soil_moisture, data = Fun.by.sample.dat_cmeta)))$coefficients[2,1],2)` \
p-value: `r (summary(lm(abund_null_dev_stoch~soil_moisture, data = Fun.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

```{r}
plot_grid(b.smg,f.smg, rel_widths = c(.44,.55))
```



# pH {.tabset .tabset-pills .unnumbered}

## **Bacteria** {.unnumbered}
```{r}
ggplot(data=Bac.by.sample.dat_cmeta, aes(y=pH, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=pH, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~pH, data = Bac.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~pH, data = Bac.by.sample.dat_cmeta)))$coefficients[2,1],2)` \
p-value: `r (summary(lm(abund_null_dev_stoch~pH, data = Bac.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

## **Fungi** {.unnumbered}
```{r}
ggplot(data=Fun.by.sample.dat_cmeta, aes(y=pH, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=pH, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~pH, data = Fun.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~pH, data = Fun.by.sample.dat_cmeta)))$coefficients[2,1],3)` \
p-value: `r (summary(lm(abund_null_dev_stoch~pH, data = Fun.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

# nitrate {.tabset .tabset-pills .unnumbered}

## **Bacteria** {.unnumbered}
```{r}
ggplot(data=Bac.by.sample.dat_cmeta, aes(y=nitrate, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alnitratea=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=nitrate, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~nitrate, data = Bac.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~nitrate, data = Bac.by.sample.dat_cmeta)))$coefficients[2,1],2)` \
p-value: `r (summary(lm(abund_null_dev_stoch~nitrate, data = Bac.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

## **Fungi** {.unnumbered}
```{r}
ggplot(data=Fun.by.sample.dat_cmeta, aes(y=nitrate, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=nitrate, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~nitrate, data = Fun.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~nitrate, data = Fun.by.sample.dat_cmeta)))$coefficients[2,1],2)` \
p-value: `r (summary(lm(abund_null_dev_stoch~nitrate, data = Fun.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

# herbivory_perc {.tabset .tabset-pills .unnumbered}

## **Bacteria** {.unnumbered}
```{r}
ggplot(data=Bac.by.sample.dat_cmeta, aes(y=herbivory_perc, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=herbivory_perc, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~herbivory_perc, data = Bac.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~herbivory_perc, data = Bac.by.sample.dat_cmeta)))$coefficients[2,1],3)` \
p-value: `r (summary(lm(abund_null_dev_stoch~herbivory_perc, data = Bac.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

## **Fungi** {.unnumbered}
```{r}
ggplot(data=Fun.by.sample.dat_cmeta, aes(y=herbivory_perc, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=herbivory_perc, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~herbivory_perc, data = Fun.by.sample.dat_cmeta)))$adj.r.squared,3)` \
slope: `r round((summary(lm(abund_null_dev_stoch~herbivory_perc, data = Fun.by.sample.dat_cmeta)))$coefficients[2,1],3)` \
p-value: `r (summary(lm(abund_null_dev_stoch~herbivory_perc, data = Fun.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

# avg_SLA {.tabset .tabset-pills .unnumbered}

## **Bacteria** {.unnumbered}
```{r}
ggplot(data=Bac.by.sample.dat_cmeta, aes(y=avg_SLA, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=avg_SLA, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~avg_SLA, data = Bac.by.sample.dat_cmeta)))$adj.r.squared,3)` \
slope: `r round((summary(lm(abund_null_dev_stoch~avg_SLA, data = Bac.by.sample.dat_cmeta)))$coefficients[2,1],4)` \
p-value: `r (summary(lm(abund_null_dev_stoch~avg_SLA, data = Bac.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

## **Fungi** {.unnumbered}
```{r}
ggplot(data=Fun.by.sample.dat_cmeta, aes(y=avg_SLA, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=avg_SLA, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~avg_SLA, data = Fun.by.sample.dat_cmeta)))$adj.r.squared,3)` \
slope: `r round((summary(lm(abund_null_dev_stoch~avg_SLA, data = Fun.by.sample.dat_cmeta)))$coefficients[2,1],4)` \
p-value: `r (summary(lm(abund_null_dev_stoch~avg_SLA, data = Fun.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

# avg_SRL {.tabset .tabset-pills .unnumbered}

## **Bacteria** {.unnumbered}
```{r}
ggplot(data=Bac.by.sample.dat_cmeta, aes(y=avg_SRL, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=avg_SRL, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~avg_SRL, data = Bac.by.sample.dat_cmeta)))$adj.r.squared,3)` \
slope: `r round((summary(lm(abund_null_dev_stoch~avg_SRL, data = Bac.by.sample.dat_cmeta)))$coefficients[2,1],5)` \
p-value: `r (summary(lm(abund_null_dev_stoch~avg_SRL, data = Bac.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

## **Fungi** {.unnumbered}
```{r}
ggplot(data=Fun.by.sample.dat_cmeta, aes(y=avg_SRL, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=avg_SRL, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~avg_SRL, data = Fun.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~avg_SRL, data = Fun.by.sample.dat_cmeta)))$coefficients[2,1],2)` \
p-value: `r (summary(lm(abund_null_dev_stoch~avg_SRL, data = Fun.by.sample.dat_cmeta)))$coefficients[2,4]` \
\


# {-}


# exploring decoupled pred var

```{r}
Bac.AllG<-readRDS(file = "../processed_data/clean_rds_saves/BacArch_min2_rare9048_v1.rds")
  # otu_table()   OTU Table:         [ 23156 taxa and 533 samples ]
  # sample_data() Sample Data:       [ 533 samples by 39 sample variables ]
Fun.AllG<-readRDS(file = "../processed_data/clean_rds_saves/Fun_min2_rare10116_v2.rds")
  # otu_table()   OTU Table:         [ 6749 taxa and 509 samples ]
  # sample_data() Sample Data:       [ 509 samples by 39 sample variables ]
Bac.AllG.k5.ord <- readRDS(file="../processed_data/clean_rds_saves/Bac.AllG.k5.ord.rds")
Fun.AllG.k5.ord <- readRDS(file="../processed_data/clean_rds_saves/Fun.AllG.k5.ord.rds")

```


## soil_moisture
negative to Bac, positive to Fun
```{r}
 adonis2(phyloseq::distance(Bac.AllG, method="bray") ~ Bin, 
          data = (as(sample_data(get(comm)), "data.frame")))
```

```{r plot nmds AllG col soil_moisture, include=F, eval=T}
nmds.Bac.AllG.colsoil_moisture_wleg<-plot_ordination(Bac.AllG, Bac.AllG.k5.ord, color = rev("soil_moisture")) +
  scale_color_gradient(high = "#132B43",low = "#56B1F7",name="Soil Moisture") +
  geom_point(size = 3, alpha = 0.9)+
  #stat_ellipse(type = "t", level = 0.9) +
  ggtitle("NMDS by Soil Moisture", subtitle = "16S") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="bottom",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

leg_nmds.colsoil_moisture<-get_legend(nmds.Bac.AllG.colsoil_moisture_wleg)
nmds.Bac.AllG.colsoil_moisture<-nmds.Bac.AllG.colsoil_moisture_wleg+theme(legend.position="none")

nmds.Fun.AllG.colsoil_moisture<-plot_ordination(Fun.AllG, Fun.AllG.k5.ord, color = rev("soil_moisture")) +
  scale_color_gradient(high = "#132B43",low = "#56B1F7") +
  geom_point(size = 3, alpha = 0.9) +
  # stat_ellipse(type = "t", level = 0.9) +
  labs(y="") +
  ggtitle("", subtitle = "ITS") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

nmds.AllG.colsoil_moisture_upper<-plot_grid(nmds.Bac.AllG.colsoil_moisture,nmds.Fun.AllG.colsoil_moisture, ncol=2)
plot_grid(nmds.AllG.colsoil_moisture_upper,leg_nmds.colsoil_moisture, nrow=2, rel_heights = c(0.85,0.15))

# ggsave(filename = "../analyses/ordinations/NMDS_AllG_soil_moisture.png", plot = last_plot(), device = "png", height = 4, width = 7, units = "in")
```


## pH
negative to Bac, no relationship to Fun

## nitrate
positive to Bac, negative to Fun

```{r, include=F, eval=F}
allM<-BOGR_dev_vals_dat_long

allM$Grass<-factor(allM$Grass)
levels(allM$Grass)
allM$Grass<-factor(allM$Grass, levels=c("ANGE","SCSC", "BOER", "BOGR", "BUDA"))

allM$Bin<-factor(allM$Bin)
levels(allM$Bin)
allM$Bin<-factor(allM$Bin, levels=c("North","North Central", "South Central", "South"))

allM$Gradient<-factor(allM$Gradient)
levels(allM$Gradient)
allM$Gradient<-factor(allM$Gradient, levels=c("West","Middle", "East"))

min(allM$abund_null_dev_stoch)
max(allM$abund_null_dev_stoch)

hist(allM$Bnull_dev)

bin_comparisons <- list( c("North", "North Central"), c("North", "South Central"), c("North", "South"),
                         c("North Central", "South Central"), c("North Central", "South"), 
                         c("South Central", "South"))

grass_comparisons <- list( c("ANGE", "SCSC"), c("ANGE", "BOER"), c("ANGE", "BOGR"), c("ANGE", "BUDA"), 
                           c("SCSC", "BOER"), c("SCSC", "BOGR"), c("SCSC", "BUDA"),
                           c("BOER", "BOGR"), c("BOER", "BUDA"),
                           c("BOGR", "BUDA"))

gradient_comparisons <- list( c("West", "Middle"), c("West", "East"), c("Middle", "East"))

domain_comparisons <- list( c("Bacteria", "Fungi"))

ggboxplot(subset(allM, Domain=="Bacteria"), x = "Bin", y = "abund_null_dev_stoch", 
               facet.by = "Gradient",
               color = "Bin", palette = bac_bin_cols_n5, ylab = "Beta null deviation value",
               xlab = "\nLatitudinal Bin",
               title = "Bacterial Communities: Stochastic", add = "jitter", ylim=c(-0.3, 0.4)) +
    stat_compare_means(comparisons = bin_comparisons, label = "p.signif",
                      #label.y = c(.2,.3, .4,.2)
                       )+
    theme(axis.text.x = element_text(color="black", angle = 90, hjust=1,vjust=.5),
          legend.position = "bottom",
          ) +
  geom_hline(yintercept =0,  linetype = 2, color="grey30")

# ggsave(file="Bnull_dev/BA_grass.bin.png", plot = last_plot(), width=8, height=6, units = "in", device="png")

ggboxplot(subset(allM, Domain=="Fungi"), x = "Bin", y = "Bnull_dev", 
               #facet.by = "Bin",
               color = "Grass", palette = bac_cols_n5, ylab = "Beta null deviation value",
               xlab = "\nLatitudinal Bin",
               title = "Fungal Communities", add = "jitter"
          #, ylim=c(-0.4, 1)
          ) +
    stat_compare_means(comparisons = bin_comparisons, label = "p.signif",
                      #label.y = c(.2,.3, .4,.2)
                       )+
    theme(axis.text.x = element_text(color="black", angle = 90, hjust=1,vjust=.5),
          legend.position = "bottom",
          ) +
  geom_hline(yintercept =0,  linetype = 2, color="grey30")
# ggsave(file="Bnull_dev/F_grass.bin.png", plot = last_plot(), width=8, height=6, units = "in", device="png")


# +
#     geom_hline(yintercept =0,  linetype = 2, color="grey30")+
#     geom_vline(xintercept =2.5,  linetype =1)
# p <- p +    annotate("text", x = .65, y = 0.5, label = "Phase I")+
#     annotate("text", x = 2.8, y = 0.5, label = "Phase II")

ggboxplot(subset(allM, Domain=="Bacteria"), x = "Bin", y = "Bnull_dev", 
               facet.by = "Grassland",
               color = "Grass", palette = bac_cols_n5, ylab = "Beta null deviation value",
               xlab = "\nLatitudinal Bin",
               title = "Bacterial Communities", add = "jitter", ylim=c(-0.3, 0.4)) +
    stat_compare_means(comparisons = bin_comparisons, label = "p.signif",
                      #label.y = c(.2,.3, .4,.2)
                       )+
    theme(axis.text.x = element_text(color="black", angle = 90, hjust=1,vjust=.5),
          legend.position = "bottom",
          ) +
  geom_hline(yintercept =0,  linetype = 2, color="grey30")

# ggsave(file="Bnull_dev/BA_grass.grassland.png", plot = last_plot(), units = "in", device="png")

ggboxplot(subset(allM, Domain=="Fungi"), x = "Bin", y = "Bnull_dev", 
               facet.by = "Grassland",
               color = "Grass", palette = bac_cols_n5, ylab = "Beta null deviation value",
               xlab = "\nLatitudinal Bin",
               title = "Fungal Communities", add = "jitter"
          , ylim=c(0, 0.9)
          ) +
    stat_compare_means(comparisons = bin_comparisons, label = "p.signif",
                      #label.y = c(.2,.3, .4,.2)
                       )+
    theme(axis.text.x = element_text(color="black", angle = 90, hjust=1,vjust=.5),
          legend.position = "bottom",
          ) +
  geom_hline(yintercept =0,  linetype = 2, color="grey30")

# ggsave(file="Bnull_dev/F_grass.grassland.png", plot = last_plot(), units = "in", device="png")

```
