# avg_SRL {.tabset .tabset-pills .unnumbered}

## **Bacteria** {.unnumbered}
```{r}
ggplot(data=Bac.by.sample.dat_cmeta, aes(y=avg_SRL, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alavg_SRLa=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=avg_SRL, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~avg_SRL, data = Bac.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~avg_SRL, data = Bac.by.sample.dat_cmeta)))$coefficients[2,1],2)` \
p-value: `r (summary(lm(abund_null_dev_stoch~avg_SRL, data = Bac.by.sample.dat_cmeta)))$coefficients[2,4]` \
\

## **Fungi** {.unnumbered}
```{r}
ggplot(data=Fun.by.sample.dat_cmeta, aes(y=avg_SRL, x=abund_null_dev_stoch, col=Grass.x)) +
  geom_point()+
  stat_smooth(method = "loess", level=0.8, se=F, alpha=0.5)+
  stat_smooth(inherit.aes = F, method = "lm", level=0.8, aes(y=avg_SRL, x=abund_null_dev_stoch), col="black", linewidth=2)
```

adj.r.squared: `r round((summary(lm(abund_null_dev_stoch~avg_SRL, data = Fun.by.sample.dat_cmeta)))$adj.r.squared,2)` \
slope: `r round((summary(lm(abund_null_dev_stoch~avg_SRL, data = Fun.by.sample.dat_cmeta)))$coefficients[2,1],2)` \
p-value: `r (summary(lm(abund_null_dev_stoch~avg_SRL, data = Fun.by.sample.dat_cmeta)))$coefficients[2,4]` \
\