


# Replace ITS taxonomy table with updated

```{r, eval=F}
tax_table(Fun_min2_rare10116)<-UNITE_alleuk.its.all.newphyla.taxa
saveRDS(Fun_min2_rare10116, file = "clean_rds_saves/Fun_min2_rare10116_noEDGE_v2.rds")
saveRDS(Fun_min2_rare10116, file = "clean_rds_saves/Fun_min2_rare10116_v2.rds")
```



# Subset by Grass Hosts

**Foundational Grasses** (n=6) \
6 plant hosts (family Poaceae): \
  - ANGE (*Andropogon gerardii* / Big bluestem) \
  - BUDA (*Bouteloua dactyloides* / Buffalograss) \
  - BOER (*B. eriopoda* / Black grama) \
  - BOGR (*B. gracilis* / Blue grama) \
  - BOCU (*B. curtipendula* / Sideoats grama) - note: only present in EDGE sites \
  - SCSC (*Schizachyrium scoparium* / Little bluestem)  \
\

## Subset 16S non-EDGE sites (post rarefaction)

```{r, echo=T}
BacArch_min2_rare9048<-readRDS(file = "clean_rds_saves/BacArch_min2_rare9048_v1.rds")
```

### 16S Big bluestem

```{r, eval=F}
BacArch_min2_rare9048_noEDGE_ANGE <- subset_samples(BacArch_min2_rare9048, Grass=="ANGE")

BacArch_min2_rare9048_noEDGE_ANGE <- prune_taxa(taxa_sums(BacArch_min2_rare9048_noEDGE_ANGE) > 0, BacArch_min2_rare9048_noEDGE_ANGE)

BacArch_min2_rare9048_noEDGE_ANGE
# otu_table()   OTU Table:         [ 14268 taxa and 106 samples ]

saveRDS(BacArch_min2_rare9048_noEDGE_ANGE, file = "clean_rds_saves/BacArch_min2_rare9048_noEDGE_ANGE_v1.rds")
```

### 16S Buffalograss

```{r, eval=F}
BacArch_min2_rare9048_noEDGE_BUDA <- subset_samples(BacArch_min2_rare9048, Grass=="BUDA")

BacArch_min2_rare9048_noEDGE_BUDA <- prune_taxa(taxa_sums(BacArch_min2_rare9048_noEDGE_BUDA) > 0, BacArch_min2_rare9048_noEDGE_BUDA)

BacArch_min2_rare9048_noEDGE_BUDA
# otu_table()   OTU Table:         [ 11048 taxa and 83 samples ]

saveRDS(BacArch_min2_rare9048_noEDGE_BUDA, file = "clean_rds_saves/BacArch_min2_rare9048_noEDGE_BUDA_v1.rds")
```

### 16S Black grama
```{r, eval=F}
BacArch_min2_rare9048_noEDGE_BOER <- subset_samples(BacArch_min2_rare9048, Grass=="BOER")

BacArch_min2_rare9048_noEDGE_BOER <- prune_taxa(taxa_sums(BacArch_min2_rare9048_noEDGE_BOER) > 0, BacArch_min2_rare9048_noEDGE_BOER)

BacArch_min2_rare9048_noEDGE_BOER
# otu_table()   OTU Table:         [ 8374 taxa and 84 samples ]

saveRDS(BacArch_min2_rare9048_noEDGE_BOER, file = "clean_rds_saves/BacArch_min2_rare9048_noEDGE_BOER_v1.rds")
```



### 16S Little bluestem
```{r, eval=F}
BacArch_min2_rare9048_noEDGE_SCSC <- subset_samples(BacArch_min2_rare9048, Grass=="SCSC")

BacArch_min2_rare9048_noEDGE_SCSC <- prune_taxa(taxa_sums(BacArch_min2_rare9048_noEDGE_SCSC) > 0, BacArch_min2_rare9048_noEDGE_SCSC)

BacArch_min2_rare9048_noEDGE_SCSC
# otu_table()   OTU Table:         [ 17568 taxa and 174 samples ]

saveRDS(BacArch_min2_rare9048_noEDGE_SCSC, file = "clean_rds_saves/BacArch_min2_rare9048_noEDGE_SCSC_v1.rds")
```

\


## Subset ITS non-EDGE sites (post rarefaction)
```{r, echo=T}
#old phyla
Fun_min2_rare10116<-readRDS(file = "clean_rds_saves/Fun_min2_rare10116_v1.rds")

#new phyla
Fun_min2_rare10116<-readRDS(file = "clean_rds_saves/Fun_min2_rare10116_v2.rds")
# otu_table()   OTU Table:         [ 6749 taxa and 509 samples ]

get_taxa_unique(Fun_min2_rare10116, taxonomic.rank = "Phylum")
```

### ITS Big bluestem
```{r, eval=F}
Fun_min2_rare10116_noEDGE_ANGE <- subset_samples(Fun_min2_rare10116, Grass=="ANGE")

Fun_min2_rare10116_noEDGE_ANGE <- prune_taxa(taxa_sums(Fun_min2_rare10116_noEDGE_ANGE) > 0, Fun_min2_rare10116_noEDGE_ANGE)

Fun_min2_rare10116_noEDGE_ANGE
# otu_table()   OTU Table:         [ 2985 taxa and 98 samples ]



saveRDS(Fun_min2_rare10116_noEDGE_ANGE, file = "clean_rds_saves/Fun_min2_rare10116_noEDGE_ANGE_v1.rds")
```

### ITS Buffalograss
```{r, eval=F}
Fun_min2_rare10116_noEDGE_BUDA <- subset_samples(Fun_min2_rare10116, Grass=="BUDA")

Fun_min2_rare10116_noEDGE_BUDA <- prune_taxa(taxa_sums(Fun_min2_rare10116_noEDGE_BUDA) > 0, Fun_min2_rare10116_noEDGE_BUDA)

Fun_min2_rare10116_noEDGE_BUDA
# otu_table()   OTU Table:         [ 2369 taxa and 80 samples ]

saveRDS(Fun_min2_rare10116_noEDGE_BUDA, file = "clean_rds_saves/Fun_min2_rare10116_noEDGE_BUDA_v1.rds")
```

### ITS Black grama
```{r, eval=F}
Fun_min2_rare10116_noEDGE_BOER <- subset_samples(Fun_min2_rare10116, Grass=="BOER")

Fun_min2_rare10116_noEDGE_BOER <- prune_taxa(taxa_sums(Fun_min2_rare10116_noEDGE_BOER) > 0, Fun_min2_rare10116_noEDGE_BOER)

Fun_min2_rare10116_noEDGE_BOER
# otu_table()   OTU Table:         [ 1662 taxa and 78 samples ]

saveRDS(Fun_min2_rare10116_noEDGE_BOER, file = "clean_rds_saves/Fun_min2_rare10116_noEDGE_BOER_v1.rds")
```


### ITS Blue grama
```{r, eval=F}
Fun_min2_rare10116_noEDGE_BOGR <- subset_samples(Fun_min2_rare10116, Grass=="BOGR")

Fun_min2_rare10116_noEDGE_BOGR <- prune_taxa(taxa_sums(Fun_min2_rare10116_noEDGE_BOGR) > 0, Fun_min2_rare10116_noEDGE_BOGR)

Fun_min2_rare10116_noEDGE_BOGR
# otu_table()   OTU Table:         [ 2668 taxa and 92 samples ]

saveRDS(Fun_min2_rare10116_noEDGE_BOGR, file = "clean_rds_saves/Fun_min2_rare10116_noEDGE_BOGR_v1.rds")
```

### ITS Little bluestem
```{r, eval=F}
Fun_min2_rare10116_noEDGE_SCSC <- subset_samples(Fun_min2_rare10116, Grass=="SCSC")

Fun_min2_rare10116_noEDGE_SCSC <- prune_taxa(taxa_sums(Fun_min2_rare10116_noEDGE_SCSC) > 0, Fun_min2_rare10116_noEDGE_SCSC)

Fun_min2_rare10116_noEDGE_SCSC
# otu_table()   OTU Table:         [ 4347 taxa and 161 samples ]

saveRDS(Fun_min2_rare10116_noEDGE_SCSC, file = "clean_rds_saves/Fun_min2_rare10116_noEDGE_SCSC_v1.rds")
```


\
\


\
\


# Export savepoint for clean phyloseq objects / OTU table

```{r eval=FALSE}
# all non-EDGE samples
saveRDS(BacArch_min2_rare9048, file = "clean_rds_saves/BacArch_min2_rare9048_v1.rds")
saveRDS(Fun_min2_rare10116, file = "clean_rds_saves/Fun_min2_rare10116_v1.rds")

# *Blue grama* samples, excluding EDGE samples
saveRDS(BacArch_min2_rare9048_noEDGE_BOGR, file = "clean_rds_saves/BacArch_min2_rare9048_noEDGE_BOGR_v1.rds")
saveRDS(Fun_min2_rare10116_noEDGE_BOGR, file = "clean_rds_saves/Fun_min2_rare10116_noEDGE_BOGR_v1.rds")

```




# Taxonomy tables

## Unique BacArch Phyla detected

```{r echo-clean-phyla-setup-16S, echo=TRUE,message=FALSE,results='hide'}
BacArch_min2_rare9048.Phyla.A <- tax_glom(BacArch_min2_rare9048, taxrank="Phylum")

BacArch_min2_rare9048.Phyla <- prune_taxa(taxa_sums(BacArch_min2_rare9048.Phyla.A) > 0, BacArch_min2_rare9048.Phyla.A)

BA_Phyla<-get_taxa_unique(BacArch_min2_rare9048.Phyla, taxonomic.rank="Phylum")
uniq_BAphyla<-data.frame(matrix(ncol=4,nrow=length(BA_Phyla)))
row.names(uniq_BAphyla)<-BA_Phyla
colnames(uniq_BAphyla)<-c("BacArch_Phylum","reads_assigned","percentage_BacArch_reads", "OTU_count")
uniq_BAphyla[,1]<-BA_Phyla
uniq_BAphyla[,2]<-taxa_sums(BacArch_min2_rare9048.Phyla)
# uniq_BAphyla[,2]<-sample_sums(BacArch_min2_rare9048.Phyla)
uniq_BAphyla[,3]<-taxa_sums(BacArch_min2_rare9048.Phyla)/sum(sample_sums(BacArch_min2_rare9048.Phyla))
```

```{r, include=FALSE}
uniq_BAphyla[1,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[1,1]))
uniq_BAphyla[2,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[2,1]))
uniq_BAphyla[3,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[3,1]))
uniq_BAphyla[4,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[4,1]))
uniq_BAphyla[5,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[5,1]))
uniq_BAphyla[6,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[6,1]))
uniq_BAphyla[7,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[7,1]))
uniq_BAphyla[8,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[8,1]))
uniq_BAphyla[9,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[9,1]))
uniq_BAphyla[10,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[10,1]))
uniq_BAphyla[11,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[11,1]))
uniq_BAphyla[12,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[12,1]))
uniq_BAphyla[13,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[13,1]))
uniq_BAphyla[14,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[14,1]))
uniq_BAphyla[15,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[15,1]))
uniq_BAphyla[16,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[16,1]))
uniq_BAphyla[17,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[17,1]))
uniq_BAphyla[18,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[18,1]))
uniq_BAphyla[19,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[19,1]))
uniq_BAphyla[20,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[20,1]))
uniq_BAphyla[21,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[21,1]))
uniq_BAphyla[22,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[22,1]))
uniq_BAphyla[23,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[23,1]))
uniq_BAphyla[24,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[24,1]))
uniq_BAphyla[25,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[25,1]))
uniq_BAphyla[26,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[26,1]))
uniq_BAphyla[27,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[27,1]))
uniq_BAphyla[28,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[28,1]))
uniq_BAphyla[29,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[29,1]))
uniq_BAphyla[30,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[30,1]))
uniq_BAphyla[31,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[31,1]))
uniq_BAphyla[32,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[32,1]))
uniq_BAphyla[33,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[33,1]))
uniq_BAphyla[34,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[34,1]))
uniq_BAphyla[35,4]<-ntaxa(subset_taxa(BacArch_min2_rare9048, Phylum==uniq_BAphyla[35,1]))

```


\

```{r echo-clean-phyla-setup-ITS, echo=TRUE,message=FALSE,results='hide'}
Fun_min2_rare10116.Phyla.A <- tax_glom(Fun_min2_rare10116, taxrank="Phylum")

Fun_min2_rare10116.Phyla <- prune_taxa(taxa_sums(Fun_min2_rare10116.Phyla.A) > 0, Fun_min2_rare10116.Phyla.A)
```

```{r, include=F, eval=F}
unrare_BacArch<-unrare_BacArch.NO_chlrpst.mito_Unclcyano.bac.arch

unrarefied_16S_NOuncl_Domain_noEuk.Phyla.A <- tax_glom(unrarefied_16S_NOuncl_Domain_noEuk, taxrank="Phylum")

unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B <- prune_taxa(taxa_sums(unrarefied_16S_NOuncl_Domain_noEuk.Phyla.A) > 0, unrarefied_16S_NOuncl_Domain_noEuk.Phyla.A)

unrarefied_16S_NOuncl_Domain_noEuk.Phyla.r  = transform_sample_counts(unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B, function(x) x / sum(x) )

unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B<-saveRDS(unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B, file="../processed_data/intermediate_rds_saves/unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B.rds")
```


```{r unique-bac-arch-phyla-set-up, eval=FALSE, echo=F, message=FALSE, warning=FALSE, include=T, results="asis"}

unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B<-readRDS(file="../processed_data/intermediate_rds_saves/unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B.rds")

BA_Phyla<-get_taxa_unique(unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B, taxonomic.rank="Phylum")
uniq_BAphyla<-data.frame(matrix(ncol=4,nrow=length(BA_Phyla)))
row.names(uniq_BAphyla)<-BA_Phyla
colnames(uniq_BAphyla)<-c("BacArch_Phylum","reads_assigned","percentage_BacArch_reads", "OTU_count")
uniq_BAphyla[,1]<-BA_Phyla
uniq_BAphyla[,2]<-taxa_sums(unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B)
# uniq_BAphyla[,2]<-sample_sums(unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B)
uniq_BAphyla[,3]<-taxa_sums(unrarefied_16S_NOuncl_Domain_noEuk.Phyla.B)/sum(sample_sums(unrarefied_16S_NOuncl_Domain_noEuk))




saveRDS(uniq_BAphyla, file = "../processed_data/intermediate_rds_saves/uniq_BAphyla.rds")

```
\
\

```{r eval=T, echo=F}
uniq_BAphyla<-readRDS(file = "../processed_data/intermediate_rds_saves/uniq_BAphyla.rds")
```


