---
title: '**RAM paired amplicons taxa queries:** Distribution and diversity of *Darksidea* and *Monosporascus* sp. with variable input data'
author: "Kelli Feeser"
subtitle: "`r params$INPUT_DATA_DESCR` "
date: "`r Sys.Date()`"
params:
  PROJABBV: "noEDGEnoMP"
  INPUT_DATA_DESCR: "Input data subset to a) exclude EDGE samples, and b) exclude fungal samples with [m]{.underline}issing bacterial [p]{.underline}air ('noEDGEnoMP')"
output:
  bookdown::html_document2:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 3
    fig.caption: yes
    keep_md: yes
  html_notebook:
    code_folding: hide
    df_print: paged
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 3
    fig.caption: yes
  prettydoc::html_pretty:
    toc: yes
    fig.caption: yes
editor_options:
  chunk_output_type: inline
---

\

::: homelink
<a href="https://kellifeeser.github.io/ksu-paired-amplicon-workflow/index.html" target="_blank" style="text-align:right">Back to Home</a>
:::

\

------------------------------------------------------------------------

Document last updated: `r Sys.Date()`

------------------------------------------------------------------------

\

# Set-up project naming from reproducible code {.unlisted .unnumbered}

```{r project-naming, include=F, eval=T}
PROJECT_PAGENAME <- 'RAM paired amplicons'
PROJECT_SUBPAGENAME <- 'Queries of taxa of interest with variable input data'
INPUT_DATA_DESCR <- "Input data subset to a) exclude EDGE samples, and b) exclude fungal samples with [m]{.underline}issing bacterial pair ('noEDGEnoMP')"
PROJABBV <- 'noEDGEnoMP'

bindingIsLocked("params", env = .GlobalEnv)
unlockBinding("params", env = .GlobalEnv)
params$INPUT_DATA_DESCR <- INPUT_DATA_DESCR

# DATASET1 <- 'Bac'
# DATASET2 <- 'Fun'

RDS1 <- 'Fun_wholecommunity.rds'
# RDS2 <- ''

#create paths for directories (sans prefix)
FIG_PATHDIR=paste0("../docs/figures/darksidea_plots/",PROJABBV)
CACHE_PATHDIR=paste0("../docs/github_darksidea_cache/github_darksidea_",PROJABBV,"_files/")
# occupancy/abundance/tax frequencies
OAF_PATHDIR=paste0("../docs/processed_data/OTU_occurrencefrequency/",PROJABBV)


FIG_PATH=paste0("../docs/figures/darksidea_plots/",PROJABBV,"/RAM_",PROJABBV,"-")
CACHE_PATH=paste0("../docs/github_darksidea_cache/github_darksidea_",PROJABBV,"_files/RAM_",PROJABBV,"_")
# occupancy/abundance/tax frequencies
OAF_PATH=paste0("../docs/processed_data/OTU_occurrencefrequency/",PROJABBV,"/RAM_",PROJABBV,"_Freq_OTU_Occurrence_")
```

```{r global-options, include = FALSE}
library(knitr)

# set up global R options
  # to list the current values of all options but better to F1
    #options()
    # easier way:
      #?options
  # to access the value of a single option
    #getOption("width")

# figure guidelines
  # save as vector (.ai, .eps, .pdf)
  # resolution: min 300 dpi
  # dimensions: 
    # Nature Research Journals
      # 1 col width = 88mm (3.45 in)
      # 2 col width = 180mm (7.08 in)
    # All Other Journals
      # 1 col width = 85mm (3.34 in)
      # 2 col width = 175mm (6.88 in)
  # file size should be < 50 MB

# text
  # font: sans-serif (helvetica or arial)
  # size: min 5pt, max 7pt

## ~~create directories~~ ##
###########################################
dir.create(FIG_PATHDIR, showWarnings = T,recursive=T)
dir.create(CACHE_PATHDIR, showWarnings = T,recursive=T)
dir.create(OAF_PATHDIR, showWarnings = T,recursive=T)
###########################################


# set up knitr global chunk options
  # See str(knitr::opts_chunk$get()) for a list of default chunk options.
#knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(echo = FALSE,
                      fig.path = FIG_PATH,
                      cache.path = CACHE_PATH,
                      eval.after = "fig.cap",
                      fig.pos = '!h'
                      )
```

```{r LANL-proxies, eval=T, include=FALSE}
# LANL proxies
Sys.setenv('HTTP_PROXY'='http://proxyout.lanl.gov:8080')
Sys.setenv('HTTPS_PROXY'='http://proxyout.lanl.gov:8080')
```

# Set-up {.unlisted .unnumbered .hidden}

## load packages

```{r load packages, eval=T, include=F}
library("ggplot2"); packageVersion("ggplot2") #‘3.4.4’
library("vegan"); packageVersion("vegan") # # ‘2.6.4’
library("phyloseq"); packageVersion("phyloseq") #‘1.44.0’
library("scales"); packageVersion("scales") #‘1.2.1’
library("swatches"); packageVersion("swatches") #‘0.5.0’
library("ggsci"); packageVersion("ggsci") #‘3.0.0’
library("ggpubr"); packageVersion("ggpubr") #‘0.6.0’
library("viridis"); packageVersion("viridis") #‘0.6.4’
library("RColorBrewer"); packageVersion("RColorBrewer") #‘1.1.3’
library("cowplot"); packageVersion("cowplot") #‘1.1.1’
library("dplyr"); packageVersion("dplyr") #‘1.1.3’
library("plyr"); packageVersion("plyr") #‘1.8.9’
library("knitr"); packageVersion("knitr") #‘1.44’
#install.packages('kableExtra', dependencies=TRUE, repos='http://cran.rstudio.com/')
#library("kableExtra"); packageVersion("kableExtra") #‘1.3.4’
library("DT"); packageVersion("DT") #‘0.30’
library("tidyverse"); packageVersion("tidyverse") #‘2.0.0’
library("prettydoc"); packageVersion("prettydoc") #‘0.4.1’
library("magrittr"); packageVersion("magrittr") #‘2.0.3’
library("broom"); packageVersion("broom") #‘1.0.5’
library("kableExtra")

# library("devtools")
# devtools::install_github("vmikk/metagMisc")
# library("metagMisc"); packageVersion("metagMisc") #‘2.0.3’



  # if (!require("BiocManager", quietly = TRUE))
  #     install.packages("BiocManager")
  # BiocManager::install(version = "3.11")
  # 
  # BiocManager::install("microbiome")
library("microbiome"); packageVersion("microbiome") #‘1.22.0’

source("ksu_colors.R")
source("../../black-sheep-club/code/sheepish_scripts.R")
```

## load rds

```{r load-rds, include=FALSE,eval=TRUE}
###.###.######.###.###
# phylo #
###.###.###
Fun<-readRDS(file = paste0("../processed_data/clean_rds_saves/",RDS1))
Fun
###.###.######.###.###
```

(ref:OA-AllG-allOTUs-fig-caption) Occupancy-abundance figure overlayed with generalist framework results from OTU down-selection for co-occurrence (top 2.5% of distribution). Plot contains data from all samples and OTUs. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data. Note both axes are shown on a log scale.

(ref:OA-BOGR-allOTUs-fig-caption) Occupancy-abundance figure overlayed with generalist framework results from OTU down-selection for co-occurrence (top 2.5% of distribution). Plot contains data from all OTUs associated with *Blue grama* hosts. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data. Note both axes are shown on a log scale.

(ref:OA-BOER-allOTUs-fig-caption) Occupancy-abundance figure overlayed with generalist framework results from OTU down-selection for co-occurrence (top 2.5% of distribution). Plot contains data from all OTUs associated with *Black grama* hosts. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data. Note both axes are shown on a log scale.

(ref:OA-BUDA-allOTUs-fig-caption) Occupancy-abundance figure overlayed with generalist framework results from OTU down-selection for co-occurrence (top 2.5% of distribution). Plot contains data from all OTUs associated with *Buffalograss* hosts. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data. Note both axes are shown on a log scale.

# Data Overview {.unnumbered}

`r INPUT_DATA_DESCR`

------------------------------------------------------------------------

### OTU tables {.unnumbered}

```{r make-Darksidea-stats-Fun_Genus_Mean_Abund, eval=T, include=F, cache=F}
dat_Genus.r <- Fun %>%
  tax_glom(taxrank = "Genus") %>%
  transform_sample_counts(function(x) {x / sum(x)} ) %>%
  psmelt()
# dat_Genus.r$Genus <- factor(dat_Genus.r$Genus)
# saveRDS(dat_Genus.r, file="../processed_data/clean_rds_saves/dat_Genus.r.rds")

# dat_Genus.r <- readRDS(file="../processed_data/clean_rds_saves/dat_Genus.r.rds")
dat_Genus.r$Genus <- factor(dat_Genus.r$Genus)

Darksidea_Genus_stats <- dat_Genus.r %>% 
  as_tibble %>%
  group_by(Genus) %>%
  reframe(
  av = mean(Abundance)*100,
  sd = sd(Abundance,na.rm = T)*100) %>% 
  filter(Genus=="Darksidea")

dat_OTU.r <- Fun %>% 
  transform_sample_counts(function(x) {x / sum(x)} ) %>% 
  psmelt()
dat_OTU.r$Genus <- factor(dat_OTU.r$Genus)
dat_OTU.r$OTU <- factor(dat_OTU.r$OTU)

Darksidea_OTU_stats <- dat_OTU.r %>% 
  as_tibble %>%
  filter(Genus=="Darksidea") %>% 
  group_by(OTU) %>%
  reframe(
  av = mean(Abundance)*100,
  sd = sd(Abundance,na.rm = T)*100) 

###############
# Monosporascus
Monosporascus_Genus_stats <- dat_Genus.r %>% 
  as_tibble %>%
  group_by(Genus) %>%
  reframe(
  av = mean(Abundance)*100,
  sd = sd(Abundance,na.rm = T)*100) %>% 
  filter(Genus=="Monosporascus")

Monosporascus_OTU_stats <- dat_OTU.r %>% 
  as_tibble %>%
  filter(Genus=="Monosporascus") %>% 
  group_by(OTU) %>%
  reframe(
  av = mean(Abundance)*100,
  sd = sd(Abundance,na.rm = T)*100) 
###############

Fun_merged <- merge_samples(Fun, "Edge",fun=mean)
Fun_merged.Genus.A <- tax_glom(Fun_merged, taxrank="Genus")
Fun_merged.Genus <- prune_taxa(taxa_sums(Fun_merged.Genus.A) > 0, Fun_merged.Genus.A)
Fun_merged.Genus.r  = transform_sample_counts(Fun_merged.Genus, function(x) x / sum(x) )
  #731

sample_names(Fun_merged.Genus.r)<-"Fun_Genus_Mean"
Fun_Genus_Mean_Abund <- merge(tax_table(Fun_merged.Genus.r)[,c("Domain","Phylum","Class","Order","Family","Genus")], t(otu_table(Fun_merged.Genus.r)), by=0, all=TRUE)
colnames(Fun_Genus_Mean_Abund)[1]<-"Fun_OTU"
Fun_Genus_Mean_Abund$Fun_Genus_Mean<-as.numeric(Fun_Genus_Mean_Abund$Fun_Genus_Mean*100)
Fun_Genus_Mean_Abund <-Fun_Genus_Mean_Abund[order(Fun_Genus_Mean_Abund$Fun_Genus_Mean, decreasing = T),]

Fun_Genus_Mean_Abund2 <- Fun_Genus_Mean_Abund 

colnames(Fun_Genus_Mean_Abund2)[8] <- "Mean rel abund (%)"
# Fun_Genus_Mean_Abund2$Rank_RA <- rank(Fun_Genus_Mean_Abund2$`Abundance (%)`,ties.method	= "max")

Fun_Genus_Mean_Abund2$Rank_RA<-(dplyr::dense_rank(desc(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`)))

Fun_Genus_Mean_Abund2$Percentile<-(dplyr::percent_rank(desc(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`)))*100


# round mean rel abundance
    # Round values less than 0.0001 to 5 decimal places
    # Round values between 0.0001 and 0.001 to 4 decimal places
    # Round values between 0.001 and 0.01 to 3 decimal places
    # Round values between 0.01 and 0.1 to 2 decimal places
    # Round values between 0.1 and 10 to 1 decimal place
    # Round values greater than 10 to 0 decimal places.


Fun_Genus_Mean_Abund2$`Mean rel abund (%)` <- ifelse(
  Fun_Genus_Mean_Abund2$`Mean rel abund (%)` < 0.0001,
  round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 5),
  ifelse(
    Fun_Genus_Mean_Abund2$`Mean rel abund (%)` < 0.001,
    round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 4),
    ifelse(
      Fun_Genus_Mean_Abund2$`Mean rel abund (%)` < 0.01,
      round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 3),
      ifelse(
        Fun_Genus_Mean_Abund2$`Mean rel abund (%)` < 0.1,
        round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 2),
        ifelse(
          Fun_Genus_Mean_Abund2$`Mean rel abund (%)` > 10,
          round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 0),
          round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 1)
        )
      )
    )
  )
)


#Fun_Genus_Mean_Abund2$Percentile <- round(Fun_Genus_Mean_Abund2$Percentile,1)

MAX_RANK_VALUE<-max(Fun_Genus_Mean_Abund2[,9])

colnames(Fun_Genus_Mean_Abund2)[9] <- paste0("Rank_RA (max ",MAX_RANK_VALUE,")")
```

```{r calc_occurrence-data-manipulation, include=F,eval=T,cache=F}

Fun_Sev <-subset_samples(Fun, Site=="SEV")
Fun_Sev_no1x <- prune_taxa(taxa_sums(Fun_Sev) > 0, Fun_Sev)

# taxa_frequency = 
  # "count" - just counts OTU occurrence (number of samples)
  # "percentage" - (default) Frequency of occurrence 
      # proportion of sampling units that contains the OTU 
      # (sample units in which OTU occurred / total # of sample units)
  # "relfreq" - Relative occurrence frequency
      # aka relative frequency of OTU 
      # (frequency of each OTU [see 'percentage'] / sum of the frequency of all OTUs)


occ_otu_null <- calc_occurrence(Fun, justdf = T)
occ_otu_site <- calc_occurrence(Fun, variable = "Site", justdf = T)
occ_otu_grass_allsites <- calc_occurrence(Fun, "Grass", justdf = T)
occ_otu_grass_SEVonly <- calc_occurrence(Fun_Sev, variable = "Grass", justdf = T)


tax_table1<-as.data.frame(tax_table(Fun))
tax_table <- tibble::rownames_to_column(tax_table1, "OTU")

colnames(occ_otu_site)[1] <- "OTU"

colnames(occ_otu_null)[1] <- "OTU"
colnames(occ_otu_null)[2] <- "global_occurrence"

colnames(occ_otu_grass_allsites)[1] <- "OTU"
colnames(occ_otu_grass_allsites)[2] <- "ANGE_allsites"
colnames(occ_otu_grass_allsites)[3] <- "BOER_allsites"
colnames(occ_otu_grass_allsites)[4] <- "BOGR_allsites"
colnames(occ_otu_grass_allsites)[5] <- "BUDA_allsites"
colnames(occ_otu_grass_allsites)[6] <- "SCSC_allsites"


colnames(occ_otu_grass_SEVonly)[1] <- "OTU"
colnames(occ_otu_grass_SEVonly)[2] <- "BOER_SEVonly"
colnames(occ_otu_grass_SEVonly)[3] <- "BOGR_SEVonly"



OTU_freq_occurr_bySite_Grass<-
  tax_table %>% 
  left_join(occ_otu_null,by = "OTU") %>% 
  left_join(occ_otu_grass_SEVonly,by = "OTU") %>% 
  left_join(occ_otu_grass_allsites,by = "OTU") %>% 
  left_join(occ_otu_site,by = "OTU")

#head(OTU_freq_occurr_bySite_Grass)

OTU_freq_occurr_bySite_Grass_Monosporascus <- subset(OTU_freq_occurr_bySite_Grass, Genus == "Monosporascus")
OTU_freq_occurr_bySite_Grass_Monosporascus$Domain<-as.factor(OTU_freq_occurr_bySite_Grass_Monosporascus$Domain)
OTU_freq_occurr_bySite_Grass_Monosporascus$OTU<-as.factor(OTU_freq_occurr_bySite_Grass_Monosporascus$OTU)

OTU_freq_occurr_bySite_Grass_Xylariales <- subset(OTU_freq_occurr_bySite_Grass, Order == "Xylariales")
OTU_freq_occurr_bySite_Grass_Darksidea <- subset(OTU_freq_occurr_bySite_Grass, Genus == "Darksidea")

# write to docs/processed_data
write.csv(OTU_freq_occurr_bySite_Grass, file=paste0(OAF_PATH,"bySitexGrass_OTU_alltaxa.csv"))

write.csv(OTU_freq_occurr_bySite_Grass_Monosporascus, file=paste0(OAF_PATH,"bySitexGrass_genus_Monosporascus.csv"))

write.csv(OTU_freq_occurr_bySite_Grass_Xylariales, file=paste0(OAF_PATH,"bySitexGrass_order_Xylariales.csv"))

write.csv(OTU_freq_occurr_bySite_Grass_Darksidea, file=paste0(OAF_PATH,"bySitexGrass_genus_Darksidea.csv"))

# write to ../processed_data
# write.csv(OTU_freq_occurr_bySite_Grass, file="../processed_data/OTU_occurrencefrequency/noEDGEnoMP/RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_OTU_alltaxa.csv")

# write.csv(OTU_freq_occurr_bySite_Grass_Monosporascus, file="../processed_data/OTU_occurrencefrequency/noEDGEnoMP/RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_genus_Monosporascus.csv")

# write.csv(OTU_freq_occurr_bySite_Grass_Xylariales, file="../processed_data/OTU_occurrencefrequency/noEDGEnoMP/RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_order_Xylariales.csv")

# write.csv(OTU_freq_occurr_bySite_Grass_Darksidea, file="../processed_data/OTU_occurrencefrequency/noEDGEnoMP/RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_genus_Darksidea.csv")

```

Taxonomy assigned using SINTAX algorithm (USEARCH v11)\
Reference database: most current UNITE UNITE USEARCH/UTAX release for eukaryotes

-   version no 9.0, release date 2023-07-18

-   taxon group: All eukaryotes

-   location: <https://dx.doi.org/10.15156/BIO/2938084>

-   citation: Abarenkov, Kessy; Zirk, Allan; Piirmann, Timo; Pöhönen, Raivo; Ivanov, Filipp; Nilsson, R. Henrik; Kõljalg, Urmas (2023): UNITE USEARCH/UTAX release for eukaryotes. Version 18.07.2023. UNITE Community. <https://doi.org/10.15156/BIO/2938084>

All EDGE samples have been excluded\
ITS OTU table rarefied to `r prettyNum(max(sample_sums(Fun)),big.mark = ",")` seqs/sample

**Fungal (ITS) Dataset:**

-   `r prettyNum(ntaxa(Fun),big.mark = ",")` OTUs
-   `r prettyNum(nsamples(Fun),big.mark = ",")` samples\
-   `r prettyNum(length(get_taxa_unique(Fun, taxonomic.rank = "Genus")),big.mark = ",")` unique genera detected, inc. *Darksidea*\

------------------------------------------------------------------------

### Available metadata {.unnumbered}

Note that numeric metadata has been mean standardized.

```{r mean-std-metadata, eval=TRUE,include=FALSE}
#######
# Fun #
#######
#Fun_m.std <- perform_mean_standardization(Fun)
# coll_GDD,GDD30yr,GDD3yr,GDD2015,ppt30yr,ppt3yr,ppt2015,soil_moisture,GWC,perc_moisture,pH,phos,ammonium,nitrate,SOM,herbivory_perc,avg_SLA,avg_SRL

#saveRDS(Fun_m.std, file="../processed_data/clean_rds_saves/Fun_m.std.rds")

# Fun_m.std <- readRDS(file="../processed_data/clean_rds_saves/Fun_m.std.rds")


# saveRDS(Fun.Genus, file="../processed_data/clean_rds_saves/Fun.Genus.rds")
# Fun.Genus_m.std <- perform_mean_standardization(Fun.Genus)
# saveRDS(Fun.Genus_m.std, file="../processed_data/clean_rds_saves/Fun.Genus_m.std.rds")
# Fun.Genus_m.std <- readRDS(file="../processed_data/clean_rds_saves/Fun.Genus_m.std.rds")

#######
# Bac #
#######
# Bac_m.std <- perform_mean_standardization(Bac)
# coll_GDD,GDD30yr,GDD3yr,GDD2015,ppt30yr,ppt3yr,ppt2015,soil_moisture,GWC,perc_moisture,pH,phos,ammonium,nitrate,SOM,herbivory_perc,avg_SLA,avg_SRL

# saveRDS(Bac_m.std, file="../processed_data/clean_rds_saves/Bac_m.std.rds")

# Bac_m.std <- readRDS(file="../processed_data/clean_rds_saves/Bac_m.std.rds")


```

```{r, echo=F, eval=F}
sd_Fun<-sample_data(Fun)
colnames(sd_Fun)
```

[*Sampling Variables:*]{.underline}

-   **Grass host** (n=5)
    -   ANGE (*Andropogon gerardii* / Big bluestem)
    -   SCSC (*Schizachyrium scoparium* / Little bluestem)
    -   BOER (*Bouteloua eriopoda* / Black grama)
    -   BOGR (*B. gracilis* / Blue grama)
    -   BUDA (*B. dactyloides* / Buffalograss)
-   **Grassland type** (n=4)
    -   desert, shortgrass, mixed grass, and tallgrass
-   **Site** (n=21)
-   GPS coordinates - **latitude** and **longitude**
-   '**Longitudinal Gradient**' (n=4)
    -   North, North Central, South Central, South
-   '**Latitudinal Bin**' (n=3)
    -   East, Middle, West
-   **Elevation** (in meters)\

[*Climate Data:*]{.underline}

-   **GDD** = growing degree days (determined over 3- and 30-year windows and in 2015)
-   **ppt** = mean annual precipitation (determined over 3- and 30-year windows and in 2015)\
    \*Data from PRISM Climate Group 2019 (Oregon State University)\

[*Edaphic Variables:*]{.underline}

-   **ammonium** (NH~4~^+^)
-   **nitrate** (NO~3~^--^)
-   **SOM** (via LoI)
-   **phosphorous** (total P)
-   **pH**
-   **soil_moisture** = percentage by weight of water in soil (by sample, Rudgers et al. called this a climate variable)\
    \*soils combined within each Grass Host x Site combination\

[*Plant Traits:*]{.underline}

-   **avg_SLA**: specific leaf area
-   **avg_SRL**: specific root length
    -   Large values of specific leaf area (SLA) and specific root length (SRL) reflect large plant investment in resource acquisition; small values indicate investment in plant tissue longevity
-   **herbivory_perc**: average site level damage estimate for herbivory, averaged over all species and individuals at the site; could be used to indicate herbivory pressure at the site level
    -   ranges from 1:17\

------------------------------------------------------------------------

\

## Links to download csv files of OTU occurrence frequencies by Site-by-Grass host combinations {.unnumbered}

`colnames` and/or `indices` = explanation

-   `OTU` = OTU ID number (fungal)

-   `Domain` : `Species` `([,c(3:9)])` = taxonomic classifications for each OTU

-   `global_occurrence` `[,10]` = proportion of samples with OTU detected over of total number of samples (max of 1)

-   Sevilleta-site only (`r nsamples(Fun_Sev)` samples total, of which `r nrow(subset(sample_data(Fun_Sev), Grass=="BOER"))` were associated with *Black grama* and `r nrow(subset(sample_data(Fun_Sev), Grass=="BOGR"))` with *Blue grama*)

    -   `BOER_SEVonly` `[,11]`

    -   `BOGR_SEVonly` `[,12]`

-   By grass host `([,c(13:17)])` = splitting samples by grass host, proportion of samples with OTU detected over of total number of samples per host. Denominator ranged from 75 - 157 (samples per grass host).

    -   `ANGE_allsites`, `BOER_allsites`, `BOGR_allsites`, `BUDA_allsites`, `SCSC_allsites`

-   By site `([,c(18:39)])` = splitting samples by site, proportion of samples with OTU detected over of total number of samples per site. Denominator ranged from 6 - 42 (samples per site).\

\

[OTU-level frequency of occurrence]{.underline}

All OTUs

-   [RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_OTU_alltaxa.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/master/processed_data/OTU_occurrencefrequency/noEDGEnoMP/RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_OTU_alltaxa.csv "RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_OTU_alltaxa.csv")

OTUs assigned to genus *Darksidea* (n = 3)

-   [RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_genus_Darksidea.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/master/processed_data/OTU_occurrencefrequency/noEDGEnoMP/RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_genus_Darksidea.csv "RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_genus_Darksidea.csv")

OTUs assigned to genus *Monosporascus* (n = 2)

-   [RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_genus_Monosporascus.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/master/processed_data/OTU_occurrencefrequency/noEDGEnoMP/RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_genus_Monosporascus.csv "RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_genus_Monosporascus.csv")

OTUs assigned to order *Xylariales* (order containing *Monosporascus*)(n = 99)

-   [RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_order_Xylariales.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/master/processed_data/OTU_occurrencefrequency/noEDGEnoMP/RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_order_Xylariales.csv "RAM_noEDGEnoMP_Freq_OTU_Occurrence_bySitexGrass_order_Xylariales.csv")

\
\
\

------------------------------------------------------------------------

## OTU-level relative abundance {.unnumbered}

Relative abundance of each OTU were averaged across all samples for these results.

***Darksidea***

This dataset contains `r ntaxa(subset_taxa(Fun, Genus=="Darksidea"))` OTUs belonging to genus *Darksidea*.

At the OTU/species-level, all *Darksidea* OTUs were classified as "`r sort(get_taxa_unique(subset_taxa(Fun, Genus=="Darksidea"), taxonomic.rank = "Species"))`".\

Mean and sd of *Darksidea* OTUs abundance across all samples:

-   `r Darksidea_OTU_stats %>% dplyr::slice(1) %$% OTU`: `r round(Darksidea_OTU_stats %>% dplyr::slice(1) %$% av,3)` ± `r round(Darksidea_OTU_stats %>% dplyr::slice(1) %$% sd,2)`%
-   `r Darksidea_OTU_stats %>% dplyr::slice(2) %$% OTU`: `r prettyNum(Darksidea_OTU_stats %>% dplyr::slice(2) %$% av,digits=2)` ± `r round(Darksidea_OTU_stats %>% dplyr::slice(2) %$% sd,3)`%
-   `r Darksidea_OTU_stats %>% dplyr::slice(3) %$% OTU`: `r prettyNum(Darksidea_OTU_stats %>% dplyr::slice(3) %$% av,digits=2)` ± `r round(Darksidea_OTU_stats %>% dplyr::slice(3) %$% sd,2)`%
    -   Global sum: `r prettyNum(sum(Darksidea_OTU_stats$av),digits=2)`%

***Monosporascus***

This dataset contains `r ntaxa(subset_taxa(Fun, Genus=="Monosporascus"))` OTUs belonging to genus *Monosporascus*.

At the OTU/species-level, all *Monosporascus* OTUs were classified as "`r sort(get_taxa_unique(subset_taxa(Fun, Genus=="Monosporascus"), taxonomic.rank = "Species"))`".

Mean and sd of *Monosporascus* OTUs abundance across all samples:

-   `r Monosporascus_OTU_stats %>% dplyr::slice(1) %$% OTU`: `r prettyNum(Monosporascus_OTU_stats %>% dplyr::slice(1) %$% av,digits=2)` ± `r prettyNum(Monosporascus_OTU_stats %>% dplyr::slice(1) %$% sd,digits=1)`%
-   `r Monosporascus_OTU_stats %>% dplyr::slice(2) %$% OTU`: `r prettyNum(Monosporascus_OTU_stats %>% dplyr::slice(2) %$% av,digits=2)` ± `r round(Monosporascus_OTU_stats %>% dplyr::slice(2) %$% sd,3)`%
    -   Global sum: `r prettyNum(sum(Monosporascus_OTU_stats$av),digits=2)`%

------------------------------------------------------------------------

## Genus-level relative abundance {.unnumbered}

Relative abundance of each OTU were averaged across all samples for these results.\

***Darksidea***

This dataset contains `r ntaxa(subset_taxa(Fun, Genus=="Darksidea"))` OTUs belonging to genus *Darksidea*.

At the species-level, all *Darksidea* OTUs were classified as "`r sort(get_taxa_unique(subset_taxa(Fun, Genus=="Darksidea"), taxonomic.rank = "Species"))`".

Across all samples, *Darksidea* had an average relative abundance of `r round(Darksidea_Genus_stats$av,3)` ± `r round(Darksidea_Genus_stats$sd,2)`%.\

***Monosporascus***

This dataset contains `r ntaxa(subset_taxa(Fun, Genus=="Monosporascus"))` OTUs belonging to genus *Monosporascus*.

At the species-level, all *Monosporascus* OTUs were classified as "`r sort(get_taxa_unique(subset_taxa(Fun, Genus=="Monosporascus"), taxonomic.rank = "Species"))`".

Across all samples, *Monosporascus* had an average relative abundance of `r prettyNum(Monosporascus_Genus_stats$av,digits=2)` ± `r round(Monosporascus_Genus_stats$sd,3)`% (sum of all relative abundances is 100%).\
\

```{r, echo=F, eval=T, warning=F}
datatable(Fun_Genus_Mean_Abund2, caption = "SEARCHABLE table of genus-level taxonomy relative abundance & abundance rank and percentile, across all non-EDGE samples.    Notes: 'Abundance' is expressed as mean relative abundance percentage of genera across all non-EDGE samples. 'Rank_RA' indicates the rank of each genus by mean global relative abundance (rank of 1 is the most abundant, 416 is the least [gapless variant of ties.method = 'min']). 'Percentile' indicates the proportion of taxa with a higher mean relative abundance (0 is the most abundant).",
          rownames = FALSE,
          filter = 'top',
          #width=680,
          #style = "font-size: 8px",
          options = list(searching = T,
                         order = list(list(7, 'desc')), #sort by decreasing abundance
                         initComplete = JS(
              "function(settings, json) {",
              "  var table = settings.oInstance.api();",
              "  var darksideaIndex = table.column(6).data().indexOf('Darksidea');",
              "  var darksideaRow = table.row(darksideaIndex).node();",
              "  var fusariumIndex = table.column(6).data().indexOf('Fusarium');",
              "  var fusariumRow = table.row(fusariumIndex).node();",
              "  var monosporascusIndex = table.column(6).data().indexOf('Monosporascus');",
              "  var monosporascusRow = table.row(monosporascusIndex).node();",
              "  var tbody = table.table().body();",
              "  $(darksideaRow).prependTo(tbody);",
              "  $(darksideaRow).css('background-color', '#94D747');",
              "  $(fusariumRow).insertAfter(darksideaRow);",
              "  $(fusariumRow).css('background-color', '#EF9F43');",
              "  $(monosporascusRow).insertAfter(fusariumRow);", # Placing Monosporascus after Fusarium
              "  $(monosporascusRow).css('background-color', '#9CD2E8');", # 
              "}"
            ),
            scrollX=TRUE,
                         columnDefs = list(list(visible = FALSE, targets = c('Fun_OTU', 'Domain', 'Order')), #domain
                                           list(width = '16%', targets = c('Class', 'Family', 'Genus')), #family
                                           list(width = '25%', targets = 'Mean rel abund (%)'), #mean RA
                                           list(width = '25%', targets = colnames(Fun_Genus_Mean_Abund2)[9]) 
                                           #list(width = '125%', targets = 8) #rank RA
                                           #,list(width = '110%', targets = 9) #percentile
                                           )
                         )
          ) %>% formatRound('Percentile', 1)

```

\
\

```{r extract_rao-function, include=F,eval=T}
extract_rao <- function(phylo_obj, spe.gen_res) {
  
  colnames(spe.gen_res)[1] <- "OTU"
  
  # convert to relative abundance
  phylo_obj.r  = transform_sample_counts(phylo_obj, function(x) x / sum(x) )
  phylo_obj.r_comm.tab<-as.data.frame(phylo2vegan_OTU(phylo_obj.r))
  
  # create presence/absence matrix
  phylo_obj.r_comm.tab.bin<-ceiling(phylo_obj.r_comm.tab/max(phylo_obj.r_comm.tab))
  
  phylo_obj_OA<-data.frame(matrix(ncol = 3, nrow = ncol(phylo_obj.r_comm.tab.bin)))
  x <- c("RelAbund", "Occurrence", "sign")
  colnames(phylo_obj_OA) <- x
  
  # report mean abundance across all samples (only if present)
  phylo_obj_OA$RelAbund<-(average_if_gt_zero(phylo_obj.r_comm.tab)*100)
  
  # report occurrence
  phylo_obj_OA$Occurrence<-(colSums(phylo_obj.r_comm.tab.bin)/dim(phylo_obj.r_comm.tab.bin)[1])
  row.names(phylo_obj_OA)<-names(phylo_obj.r_comm.tab)
  phylo_obj_OA$OTU <- row.names(phylo_obj_OA)
  
  # add spe.gen results
  phylo_obj_OA_sign<-merge(phylo_obj_OA,spe.gen_res[,-(2:5)],
                           by = "OTU", all.x=T, no.dups=F) %>%
    select(-sign.x)

  colnames(phylo_obj_OA_sign)[4] <- "Sign"

  phylo_obj_OA_sign_tax<-cbind(phylo_obj_OA_sign,as(tax_table(phylo_obj), "matrix"))

  phylo_obj_OA_sign_tax$Genus<-factor(phylo_obj_OA_sign_tax$Genus)

  phylo_obj_OA_sign_tax$Sign <- phylo_obj_OA_sign_tax$Sign %>% replace_na('Singletons')
  
  phylo_obj_OA_sign_tax$Sign <- factor(phylo_obj_OA_sign_tax$Sign)
  
  phylo_obj_OA_sign_tax$Sign <- factor(phylo_obj_OA_sign_tax$Sign, 
                                       levels = c("Generalists","Specialists","NS","Singletons"))
  
  phylo_obj_OA_sign_tax
  
  # print(levels(phylo_obj_OA_sign_tax$Sign))
  
  }
```

```{r darksidea-data-manipulation, include=F, eval=TRUE, warning=F, cache=T}

# AllG
fun_AllG <- Fun
fun_AllG_OTU_spegen_res <- read.csv(file="../analyses/spe.gen/GvS_AllGrasses/GvS_Res_ITS_AllGrasses_nperm100.csv")
fun_AllG_OTU_OA_sign_tax <- extract_rao(fun_AllG, fun_AllG_OTU_spegen_res)

# BOGR
fun_BOGR <- readRDS(file = "../processed_data/clean_rds_saves/Fun_min2_rare10116_noEDGE_BOGR_v1.rds")
fun_BOGR_OTU_spegen_res <- read.csv(file="../analyses/spe.gen/GvS_Bluegrama/GvS_Res_ITS_BOGR_nperm500.csv")
fun_BOGR_OTU_OA_sign_tax <- extract_rao(fun_BOGR, fun_BOGR_OTU_spegen_res)

# BOER
fun_BOER <- readRDS(file = "../processed_data/clean_rds_saves/Fun_min2_rare10116_noEDGE_BOER_v1.rds")
fun_BOER_OTU_spegen_res <- read.csv(file="../analyses/spe.gen/GvS_Blackgrama/GvS_Res_ITS_BOER_nperm500.csv")
fun_BOER_OTU_OA_sign_tax <- extract_rao(fun_BOER, fun_BOER_OTU_spegen_res)

# BUDA
fun_BUDA <- readRDS(file = "../processed_data/clean_rds_saves/Fun_min2_rare10116_noEDGE_BUDA_v1.rds")
fun_BUDA_OTU_spegen_res <- read.csv(file="../analyses/spe.gen/GvS_Buffalograss/GvS_Res_ITS_BUDA_nperm500.csv")
fun_BUDA_OTU_OA_sign_tax <- extract_rao(fun_BUDA, fun_BUDA_OTU_spegen_res)

# ANGE
fun_ANGE <- readRDS(file = "../processed_data/clean_rds_saves/Fun_min2_rare10116_noEDGE_ANGE_v1.rds")
fun_ANGE_OTU_spegen_res <- read.csv(file="../analyses/spe.gen/GvS_BigBluestem/GvS_Res_ITS_ANGE_nperm500.csv")
fun_ANGE_OTU_OA_sign_tax <- extract_rao(fun_ANGE, fun_ANGE_OTU_spegen_res)

# SCSC
fun_SCSC <- readRDS(file = "../processed_data/clean_rds_saves/Fun_min2_rare10116_noEDGE_SCSC_v1.rds")
fun_SCSC_OTU_spegen_res <- read.csv(file="../analyses/spe.gen/GvS_LittleBluestem/GvS_Res_ITS_SCSC_nperm500.csv")
fun_SCSC_OTU_OA_sign_tax <- extract_rao(fun_SCSC, fun_SCSC_OTU_spegen_res)

```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Occupancy-abundance, OTU-level {.tabset .unnumbered}

[only showing results for EDGE samples excluded, no missing pairs]

\

## ***Darksidea*** {.tabset .tabset-pills .unnumbered}

*Darksidea* sp. only detected under *Blue grama* and *Black grama* hosts. Figures below show data from all samples and sites, or samples associated with specific grass hosts as indicated.

### All Grasses {.unnumbered}

```{r, echo=F,eval=T,fig.width=8, fig.cap='(ref:OA-AllG-allOTUs-fig-caption)', cache=T}
# OTUs - abundance vs occurrence plot 
ggplot(fun_AllG_OTU_OA_sign_tax, aes(y = (RelAbund/100), x = Occurrence, color = Sign, shape = Sign)) +
  geom_jitter(size = 1.8, alpha=0.4,width = 0.02,height = 0.025) +
  scale_shape_manual(
    values = c(
      "Darksidea" = 17,
      "Generalists" = 16,
      "Specialists" = 16,
      "NS" = 16,
      "Singletons" = 1),
    guide = "legend") +
  scale_color_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singletons" = "grey"),
    guide = "legend"
  ) +
  geom_point(
    data = subset(fun_AllG_OTU_OA_sign_tax, Genus == "Darksidea"),
    aes(y = (RelAbund/100), x = Occurrence, size = "Darksidea"),
    color = "black", shape = 17, size = 8, alpha=1
  ) +
  labs(
    y = "Mean relative abundance, if present",
    x = "Frequency of OTU occurrence by sample\n(proportion of samples with OTU present)",
    title = "Occupancy-abundance: fungal OTUs from all grass hosts"
    ,subtitle = expression("OTUs belonging to " * italic("Darksidea") * " represented as black triangles (n=3)")
    # ,caption = "Note that both axes are shown on a log scale. Plot contains data from all samples and OTUs. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data."
  ) +
  scale_y_log10(labels = scales::percent,limits = c(0.00009, 0.5), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.5)) +
  scale_x_log10(limits = c(0.0015, 1), breaks=c(0.002, 0.01, 0.1, 0.2, 1)) +
  # scale_size_manual(
  #   values = c(italic("Darksidea") = 5),
  #   guide = "legend"
  # ) +
  guides(
    color = guide_legend(title = NULL,override.aes = list(size=3,alpha = 0.8)),
    shape = guide_legend(title = NULL),
    size = guide_legend(title = NULL)
  ) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(color="black", size=13),
    axis.text.y = element_text(color="black", size=13),
    legend.text = element_text(color="black", size=11),
    legend.title = element_blank(),
    legend.position=c(.8,.87),
    legend.key.height = unit(0.4, 'cm'),
    plot.title = element_text(color="black", size=14),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

```

### All Grasses (no *Darksidea*) {.unnumbered}

```{r, echo=F,eval=T,fig.width=8, fig.cap='(ref:OA-AllG-allOTUs-fig-caption)', cache=T}
# OTUs - abundance vs occurrence plot 
ggplot(fun_AllG_OTU_OA_sign_tax, aes(y = (RelAbund/100), x = Occurrence, color = Sign, shape = Sign)) +
  geom_jitter(size = 1.8, alpha=0.4,width = 0.02,height = 0.025) +
  scale_shape_manual(
    values = c(
      "Darksidea" = 17,
      "Generalists" = 16,
      "Specialists" = 16,
      "NS" = 16,
      "Singletons" = 1),
    guide = "legend") +
  scale_color_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singletons" = "grey"),
    guide = "legend"
  # ) +
  # geom_point(
  #   data = subset(fun_AllG_OTU_OA_sign_tax, Genus == "Darksidea"),
  #   aes(y = (RelAbund/100), x = Occurrence, size = "Darksidea"),
  #   color = "black", shape = 17, size = 8, alpha=1
  ) +
  labs(
    y = "Mean relative abundance, if present",
    x = "Frequency of OTU occurrence by sample\n(proportion of samples with OTU present)",
    title = "Occupancy-abundance: fungal OTUs from all grass hosts"
    # ,subtitle = expression("OTUs belonging to " * italic("Darksidea") * " represented as black triangles (n=3)")
    # ,caption = "Note that both axes are shown on a log scale. Plot contains data from all samples and OTUs. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data."
  ) +
  scale_y_log10(labels = scales::percent,limits = c(0.00009, 0.5), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.5)) +
  scale_x_log10(limits = c(0.0015, 1), breaks=c(0.002, 0.01, 0.1, 0.2, 1)) +
  # scale_size_manual(
  #   values = c(italic("Darksidea") = 5),
  #   guide = "legend"
  # ) +
  guides(
    color = guide_legend(title = NULL,override.aes = list(size=3,alpha = 0.8)),
    shape = guide_legend(title = NULL),
    size = guide_legend(title = NULL)
  ) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(color="black", size=13),
    axis.text.y = element_text(color="black", size=13),
    legend.text = element_text(color="black", size=11),
    legend.title = element_blank(),
    legend.position=c(.8,.87),
    legend.key.height = unit(0.4, 'cm'),
    plot.title = element_text(color="black", size=14),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

# ggsave(filename = "../figures/darksidea_plots/OccAbund_GvS_ITS_AllGrasses.pdf", plot = last_plot(), device = "pdf",width=10, height = 5, units = "in", dpi=300)
```

### *Blue grama* {.unnumbered}

```{r, echo=F,eval=T,fig.width=8, fig.cap='(ref:OA-BOGR-allOTUs-fig-caption)', cache=T}
# OTUs - abundance vs occurrence plot 
ggplot(fun_BOGR_OTU_OA_sign_tax, aes(y = (RelAbund/100), x = Occurrence, color = Sign, shape = Sign)) +
  geom_jitter(size = 1.8, alpha=0.4,width = 0.02,height = 0.025) +
  scale_shape_manual(
    values = c(
      "Darksidea" = 17,
      "Generalists" = 16,
      "Specialists" = 16,
      "NS" = 16,
      "Singletons" = 1),
    guide = "legend") +
  scale_color_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singletons" = "grey"),
    guide = "legend"
  ) +
  geom_point(
    data = subset(fun_BOGR_OTU_OA_sign_tax, Genus == "Darksidea"),
    aes(y = (RelAbund/100), x = Occurrence, size = "Darksidea"),
    color = "black", shape = 17, size = 8, alpha=1
  ) +
  labs(
    y = "Mean relative abundance, if present",
    x = "Frequency of OTU occurrence by sample\n(proportion of samples with OTU present)",
    title = expression("Occupancy-abundance: fungal OTUs from " * italic("Blue grama")),
    subtitle = expression("OTUs belonging to " * italic("Darksidea") * " represented as black triangles (n=2)")
    # ,caption = "Note that both axes are shown on a log scale. Plot contains data from all samples and OTUs. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data."
  ) +
  scale_y_log10(labels = scales::percent,limits = c(0.00004, 0.6), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.6)) +
  scale_x_log10(limits = c(0.009, 1), breaks=c(0.01, 0.025, 0.1, 0.2, 1)) +
                #,expand = expansion(mult = c(0.055, 0))) +
  # scale_size_manual(
  #   values = c(italic("Darksidea") = 5),
  #   guide = "legend"
  # ) +
  guides(
    color = guide_legend(title = NULL,override.aes = list(size=3,alpha = 0.8)),
    shape = guide_legend(title = NULL),
    size = guide_legend(title = NULL)
  ) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(color="black", size=13),
    axis.text.y = element_text(color="black", size=13),
    legend.text = element_text(color="black", size=11),
    legend.title = element_blank(),
    legend.position=c(.8,.87),
    legend.key.height = unit(0.4, 'cm'),
    plot.title = element_text(color="black", size=14),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )
```

### *Blue grama* (no *Darksidea*) {.unnumbered}

```{r, echo=F,eval=T,fig.width=8, fig.cap='(ref:OA-BOGR-allOTUs-fig-caption)', cache=T}
# OTUs - abundance vs occurrence plot 
ggplot(fun_BOGR_OTU_OA_sign_tax, aes(y = (RelAbund/100), x = Occurrence, color = Sign, shape = Sign)) +
  geom_jitter(size = 1.8, alpha=0.4,width = 0.02,height = 0.025) +
  scale_shape_manual(
    values = c(
      "Darksidea" = 17,
      "Generalists" = 16,
      "Specialists" = 16,
      "NS" = 16,
      "Singletons" = 1),
    guide = "legend") +
  scale_color_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singletons" = "grey"),
    guide = "legend"
  # ) +
  # geom_point(
  #   data = subset(fun_BOGR_OTU_OA_sign_tax, Genus == "Darksidea"),
  #   aes(y = (RelAbund/100), x = Occurrence, size = "Darksidea"),
  #   color = "black", shape = 17, size = 8, alpha=1
  ) +
  labs(
    y = "Mean relative abundance, if present",
    x = "Frequency of OTU occurrence by sample\n(proportion of samples with OTU present)",
    title = expression("Occupancy-abundance: fungal OTUs from " * italic("Blue grama"))
    # ,subtitle = expression("OTUs belonging to " * italic("Darksidea") * " represented as black triangles (n=2)")
    # ,caption = "Note that both axes are shown on a log scale. Plot contains data from all samples and OTUs. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data."
  ) +
  # scale_y_log10(labels = scales::percent,limits = c(0.00004, 0.6), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.6)) +
  # scale_x_log10(limits = c(0.009, 1), breaks=c(0.01, 0.025, 0.1, 0.2, 1)) +
  scale_y_log10(labels = scales::percent,limits = c(0.00004, 0.61), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.6)) +
  scale_x_log10(limits = c(0.007, 1), breaks=c(0.01, 0.025, 0.1, 0.2, 1)) +
  guides(
    color = guide_legend(title = NULL,override.aes = list(size=3,alpha = 0.8)),
    shape = guide_legend(title = NULL),
    size = guide_legend(title = NULL)
  ) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(color="black", size=13),
    axis.text.y = element_text(color="black", size=13),
    legend.text = element_text(color="black", size=11),
    legend.title = element_blank(),
    legend.position=c(.8,.87),
    legend.key.height = unit(0.4, 'cm'),
    plot.title = element_text(color="black", size=14),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

ggsave(filename = "../figures/darksidea_plots/OccAbund_GvS_ITS_Bluegrama.pdf", plot = last_plot(), device = "pdf",width=10, height = 5, units = "in", dpi=300)
```

### *Black grama* {.unnumbered}

```{r, echo=F,eval=T,fig.width=8, fig.cap='(ref:OA-BOER-allOTUs-fig-caption)', cache=T}
# OTUs - abundance vs occurrence plot 
ggplot(fun_BOER_OTU_OA_sign_tax, aes(y = (RelAbund/100), x = Occurrence, color = Sign, shape = Sign)) +
  geom_jitter(size = 1.8, alpha=0.4,width = 0.02,height = 0.025) +
  scale_shape_manual(
    values = c(
      "Darksidea" = 17,
      "Generalists" = 16,
      "Specialists" = 16,
      "NS" = 16,
      "Singletons" = 1),
    guide = "legend") +
  scale_color_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singletons" = "grey"),
    guide = "legend"
  ) +
  geom_point(
    data = subset(fun_BOER_OTU_OA_sign_tax, Genus == "Darksidea"),
    aes(y = (RelAbund/100), x = Occurrence, size = "Darksidea"),
    color = "black", shape = 17, size = 8, alpha=1
  ) +
  labs(
    y = "Mean relative abundance, if present",
    x = "Frequency of OTU occurrence by sample\n(proportion of samples with OTU present)",
    title = expression("Occupancy-abundance: fungal OTUs from " * italic("Black grama")),
    subtitle = expression("OTUs belonging to " * italic("Darksidea") * " represented as black triangles (n=3)")
  ) +
  #scale_y_log10(labels = scales::percent,limits = c(0.00006, 0.6), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.6)) +
  scale_y_log10(labels = scales::percent,limits = c(0.00006, 0.6), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.6)) +
  scale_x_log10(limits = c(0.01, 1), breaks=c(0.012, 0.025, 0.1, 0.2, 1)) +
  # scale_size_manual(
  #   values = c(italic("Darksidea") = 5),
  #   guide = "legend"
  # ) +
  guides(
    color = guide_legend(title = NULL,override.aes = list(size=3,alpha = 0.8)),
    shape = guide_legend(title = NULL),
    size = guide_legend(title = NULL)
  ) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(color="black", size=13),
    axis.text.y = element_text(color="black", size=13),
    legend.text = element_text(color="black", size=11),
    legend.title = element_blank(),
    legend.position=c(.8,.87),
    legend.key.height = unit(0.4, 'cm'),
    plot.title = element_text(color="black", size=14),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )
```

\

***Monosporascus***\

*Monosporascus* sp. only detected under *Blue grama* and *Buffalograss* hosts. Figures below show data from all samples and sites, or samples associated with specific grass hosts as indicated.

## ***Monosporascus*** {.tabset .tabset-pills .unnumbered}

*Monosporascus* sp. only detected at sites FCT and FMT (OTU4769) and CNF (OTU8515) under *Blue grama* and *Buffalograss* (*B. dactyloides*) hosts. Figures below show data from all samples and sites, or samples associated with specific grass hosts as indicated.

### All Grasses {.unnumbered}

```{r, echo=F,eval=T,fig.width=8, fig.cap='(ref:OA-AllG-allOTUs-fig-caption)', cache=T}
# OTUs - abundance vs occurrence plot 
ggplot(fun_AllG_OTU_OA_sign_tax, aes(y = (RelAbund/100), x = Occurrence, color = Sign, shape = Sign)) +
  geom_jitter(size = 1.8, alpha=0.4,width = 0.02,height = 0.025) +
  scale_shape_manual(
    values = c(
      "Monosporascus" = 15,
      "Generalists" = 16,
      "Specialists" = 16,
      "NS" = 16,
      "Singletons" = 1),
    guide = "legend") +
  scale_color_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singletons" = "grey"),
    guide = "legend"
  ) +
  geom_point(
    data = subset(fun_AllG_OTU_OA_sign_tax, Genus == "Monosporascus"),
    aes(y = (RelAbund/100), x = Occurrence, size = "Monosporascus"),
    color = "black", shape = 15, size = 8, alpha=1
  ) +
  labs(
    y = "Mean relative abundance, if present",
    x = "Frequency of OTU occurrence by sample\n(proportion of samples with OTU present)",
    title = "Occupancy-abundance: fungal OTUs from all grass hosts",
    subtitle = expression("OTUs belonging to " * italic("Monosporascus") * " represented as black squares (n=2)")
    # ,caption = "Note that both axes are shown on a log scale. Plot contains data from all samples and OTUs. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data."
  ) +
  scale_y_log10(labels = scales::percent,limits = c(0.00009, 0.5), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.5)) +
  scale_x_log10(limits = c(0.0015, 1), breaks=c(0.002, 0.01, 0.1, 0.2, 1)) +
  # scale_size_manual(
  #   values = c(italic("Monosporascus") = 5),
  #   guide = "legend"
  # ) +
  guides(
    color = guide_legend(title = NULL,override.aes = list(size=3,alpha = 0.8)),
    shape = guide_legend(title = NULL),
    size = guide_legend(title = NULL)
  ) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(color="black", size=13),
    axis.text.y = element_text(color="black", size=13),
    legend.text = element_text(color="black", size=11),
    legend.title = element_blank(),
    legend.position=c(.8,.87),
    legend.key.height = unit(0.4, 'cm'),
    plot.title = element_text(color="black", size=14),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )
```

### *Blue grama* {.unnumbered}

```{r, echo=F,eval=T,fig.width=8, fig.cap='(ref:OA-BOGR-allOTUs-fig-caption)', cache=T}
# OTUs - abundance vs occurrence plot 
ggplot(fun_BOGR_OTU_OA_sign_tax, aes(y = (RelAbund/100), x = Occurrence, color = Sign, shape = Sign)) +
  geom_jitter(size = 1.8, alpha=0.4,width = 0.02,height = 0.025) +
  scale_shape_manual(
    values = c(
      "Monosporascus" = 15,
      "Generalists" = 16,
      "Specialists" = 16,
      "NS" = 16,
      "Singletons" = 1),
    guide = "legend") +
  scale_color_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singletons" = "grey"),
    guide = "legend"
  ) +
  geom_point(
    data = subset(fun_BOGR_OTU_OA_sign_tax, Genus == "Monosporascus"),
    aes(y = (RelAbund/100), x = Occurrence, size = "Monosporascus"),
    color = "black", shape = 15, size = 8, alpha=1
  ) +
  labs(
    y = "Mean relative abundance, if present",
    x = "Frequency of OTU occurrence by sample\n(proportion of samples with OTU present)",
    title = expression("Occupancy-abundance: fungal OTUs from " * italic("Blue grama")),
    subtitle = expression("OTUs belonging to " * italic("Monosporascus") * " represented as black squares (n=2)")
    # ,caption = "Note that both axes are shown on a log scale. Plot contains data from all samples and OTUs. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data."
  ) +
  scale_y_log10(labels = scales::percent,limits = c(0.00004, 0.6), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.6)) +
  scale_x_log10(limits = c(0.009, 1), breaks=c(0.01, 0.025, 0.1, 0.2, 1)) +
                #,expand = expansion(mult = c(0.055, 0))) +
  # scale_size_manual(
  #   values = c(italic("Monosporascus") = 5),
  #   guide = "legend"
  # ) +
  guides(
    color = guide_legend(title = NULL,override.aes = list(size=3,alpha = 0.8)),
    shape = guide_legend(title = NULL),
    size = guide_legend(title = NULL)
  ) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(color="black", size=13),
    axis.text.y = element_text(color="black", size=13),
    legend.text = element_text(color="black", size=11),
    legend.title = element_blank(),
    legend.position=c(.8,.87),
    legend.key.height = unit(0.4, 'cm'),
    plot.title = element_text(color="black", size=14),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )
```

### *Buffalograss* {.unnumbered}

```{r, echo=F,eval=T,fig.width=8, fig.cap='(ref:OA-BUDA-allOTUs-fig-caption)', cache=T}
# OTUs - abundance vs occurrence plot 
ggplot(fun_BUDA_OTU_OA_sign_tax, aes(y = (RelAbund/100), x = Occurrence, color = Sign, shape = Sign)) +
  geom_jitter(size = 1.8, alpha=0.4,width = 0.02,height = 0.025) +
  scale_shape_manual(
    values = c(
      "Monosporascus" = 15,
      "Generalists" = 16,
      "Specialists" = 16,
      "NS" = 16,
      "Singletons" = 1),
    guide = "legend") +
  scale_color_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singletons" = "grey"),
    guide = "legend"
  ) +
  geom_point(
    data = subset(fun_BUDA_OTU_OA_sign_tax, Genus == "Monosporascus"),
    aes(y = (RelAbund/100), x = Occurrence, size = "Monosporascus"),
    color = "black", shape = 15, size = 8, alpha=1
  ) +
  labs(
    y = "Mean relative abundance, if present",
    x = "Frequency of OTU occurrence by sample\n(proportion of samples with OTU present)",
    title = expression("Occupancy-abundance: fungal OTUs from " * italic("Buffalograss")),
    subtitle = expression("OTUs belonging to " * italic("Monosporascus") * " represented as black squares (n=1)")
  ) +
  #scale_y_log10(labels = scales::percent,limits = c(0.00006, 0.6), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.6)) +
  scale_y_log10(labels = scales::percent,limits = c(0.00006, 0.6), breaks=c(0.0001, 0.001, 0.01, 0.1, 0.6)) +
  scale_x_log10(limits = c(0.01, 1), breaks=c(0.012, 0.025, 0.1, 0.2, 1)) +
  # scale_size_manual(
  #   values = c(italic("Monosporascus") = 5),
  #   guide = "legend"
  # ) +
  guides(
    color = guide_legend(title = NULL,override.aes = list(size=3,alpha = 0.8)),
    shape = guide_legend(title = NULL),
    size = guide_legend(title = NULL)
  ) +
  theme_bw() +
  theme(
    plot.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank(),
    axis.text.x = element_text(color="black", size=13),
    axis.text.y = element_text(color="black", size=13),
    legend.text = element_text(color="black", size=11),
    legend.title = element_blank(),
    legend.position=c(.8,.87),
    legend.key.height = unit(0.4, 'cm'),
    plot.title = element_text(color="black", size=14),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )
```

\

------------------------------------------------------------------------

# Distribution of *Darksidea* OTUs {.unnumbered}

n=3 *Darksidea* OTUs

[At the SEV only:]{.underline}

-   Of the 2 grasses present at the SEV (*Blue* and *Black grama*),

    -   0 OTUs detected in *Blue* *grama* associated soils

    -   1 / 3 OTUs detected in 11.1% of *Black* *grama* associated soils

[Across all sites]{.underline}

-   Mean occurrence frequency of the 3 OTUs across all sites and samples: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]$global_occurrence)*100,digits=2)`%

-   At least 1 OTU detected in all 5 grasses

    -   ANGE (*Andropogon gerardii* / Big bluestem) associated soils

        -   Lowest occurrence frequencies

        -   1 / 3 OTUs detected with an occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$ANGE_allsites >0,]$ANGE_allsites, na.rm = TRUE)*100,digits=2)`%, others not present\

    -   BOER (*Bouteloua eriopoda* / Black grama) associated soils

        -   3 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BOER_allsites >0,]$BOER_allsites, na.rm = TRUE)*100,digits=2)`%\
            \

    -   BOGR (*B. gracilis* / Blue grama) associated soils

        -   2 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BOGR_allsites >0,]$BOGR_allsites, na.rm = TRUE)*100,digits=3)`%, when present\
            \

    -   BUDA (*B. dactyloides* / Buffalograss) associated soils

        -   2 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BUDA_allsites >0,]$BUDA_allsites, na.rm = TRUE)*100,digits=3)`%, when present\
            \

    -   SCSC (*Schizachyrium scoparium* / Little bluestem) associated soils

        -   3 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$SCSC_allsites >0,]$SCSC_allsites, na.rm = TRUE)*100,digits=2)`%\
            \

[By Site]{.underline}

-   23 sites in dataset

-   1 site had all 3 *Darksidea* OTUs detected, with the following mean frequency

    -   BLM: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BLM >0,]$BLM, na.rm = TRUE)*100,digits=3)`%\

-   9 sites had 2 of 3 *Darksidea* OTUs detected, with the following mean frequencies (when present)

    -   OTUs `OTU9749` and `OTU5613`
        -   CPR: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$CPR >0,]$CPR, na.rm = TRUE)*100,digits=3)`%
        -   FCP: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$FCP >0,]$FCP, na.rm = TRUE)*100,digits=3)`%
        -   FMT: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$FMT >0,]$FMT, na.rm = TRUE)*100,digits=2)`%
        -   HAR: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$HAR >0,]$HAR, na.rm = TRUE)*100,digits=3)`%
        -   HPG: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$HPG >0,]$HPG, na.rm = TRUE)*100,digits=3)`%
        -   KNZ: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$KNZ >0,]$KNZ, na.rm = TRUE)*100,digits=3)`%
        -   LBJ: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$LBJ >0,]$LBJ, na.rm = TRUE)*100,digits=2)`%
        -   RNF: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$RNF >0,]$RNF, na.rm = TRUE)*100,digits=3)`%
    -   OTUs `OTU7431` and `OTU5613`
        -   GMT: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$GMT >0,]$GMT, na.rm = TRUE)*100,digits=3)`%

-   5 sites had 1 of 3 *Darksidea* OTUs detected, with the following mean frequencies (when present)

    -   OTU `OTU9749`
        -   CNF: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$CNF >0,]$CNF, na.rm = TRUE)*100,digits=3)`%
        -   DMT: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$DMT >0,]$DMT, na.rm = TRUE)*100,digits=3)`%
        -   KAE: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$KAE >0,]$KAE, na.rm = TRUE)*100,digits=3)`%
        -   SEV: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$SEV >0,]$SEV, na.rm = TRUE)*100,digits=3)`%
    -   OTU `OTU5613`
        -   SCP: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$SCP >0,]$SCP, na.rm = TRUE)*100,digits=4)`%

-   8 sites had no *Darksidea* OTUs detected (BNP, CAD, GNF, LAR, NWP, ONF, SFA, UHC)\
    \

------------------------------------------------------------------------

# Distribution of *Monosporascus* OTUs {.unnumbered}

n=2 *Monosporascus* OTUs

[Not detected at the SEV]{.underline}

-   Of the 2 grasses present at the SEV (*Blue* and *Black grama*),

    -   0 OTUs detected in *Blue grama* associated soils

    -   0 OTUs detected in *Black grama* associated soils

[Across all sites]{.underline}

-   Mean occurrence frequency of the 2 OTUs across all sites and samples: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]$global_occurrence)*100,digits=2)`%

-   No OTUs detected in 3 grasses

    -   ANGE (*Andropogon gerardii* / Big bluestem) associated soils
    -   BOER (*Bouteloua eriopoda* / Black grama) associated soils
    -   SCSC (*Schizachyrium scoparium* / Little bluestem) associated soils\
        \

-   At least 1 OTU detected in 2 grasses

    -   BOGR (*B. gracilis* / Blue grama) associated soils

        -   2 / 2 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$BOGR_allsites >0,]$BOGR_allsites, na.rm = TRUE)*100,digits=2)`%, when present\
            \

    -   BUDA (*B. dactyloides* / Buffalograss) associated soils

        -   1 / 2 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$BUDA_allsites >0,]$BUDA_allsites, na.rm = TRUE)*100,digits=3)`%, when present\
            \

[By Site]{.underline}

-   23 sites in dataset

-   0 sites had neither of the *Monosporascus* OTUs detected

-   3 sites had 1 of 2 *Monosporascus* OTUs detected, with the following mean frequencies (when present)

    -   OTU `OTU8515`
        -   CNF: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$CNF >0,]$CNF, na.rm = TRUE)*100,digits=3)`%
    -   OTU `OTU4769`
        -   FCP: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$FCP >0,]$FCP, na.rm = TRUE)*100,digits=3)`%
        -   FMT: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$FMT >0,]$FMT, na.rm = TRUE)*100,digits=2)`%

-   20 sites had no *Monosporascus* OTUs detected (BLM, BNP, CAD, CPR, DMT, GMT, GNF, HAR, HPG, KAE, KNZ, LAR, LBJ, NWP, ONF, RNF, SCP, SEV, SFA, UHC)\
    \

```{r distribution-stats, include=F, eval=F}
View(t(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]))

# At the SEV only
OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]$BOGR_SEVonly
OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]$BOER_SEVonly

OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]$BOGR_SEVonly
OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]$BOER_SEVonly



# OTU occurrence by site and grass
OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]

OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]


prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]$global_occurrence)*100,digits=2)

prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BOER_allsites >0,]$BOER_allsites, na.rm = TRUE)*100,digits=2)

# idk
View(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",c(1,20:39)])

mean(OTU_freq_occurr_bySite_Grass$ANGE_allsites[Genus=="Darksidea"])

```

------------------------------------------------------------------------

# Metadata variables impacting presence of *Darksidea* OTUs {.unnumbered}

*Methods*: For discrete metadata factors (e.g., Grass Host, Site), calculating occurrence for *Darksidea* OTUs by factor level and assessing significance of Occurrence \~ Factor via Kruskal-Wallis rank sum test ("KW"). For continuous metadata variables, I have converted the *Darksidea* OTU table to presence/absence, and assessed significance via a one-way ANOVA. All statistical tests in this section are single-term only. All continuous metadata variables have been mean standardized with the exceptions of elevation, latitude, and longitude.

-   Furthermore I calculated the variance explained by each variable

    -   Kruskal-Wallis: $\eta^2$ (effect size, essentially equivalent to $R^2$) using $\eta$$p^2 = (F * dfn)/(F * dfn + dfd)$, given $F(dfn, dfd)$=$\eta^2/(k-1)$. Because these are single term-based statistics and consequently, no other variable accounts for variance in the dependent variable, $\eta$$p²$ equals $\eta^2$.
    -   ANOVA: traditional $R^2$ as calculated using sum of squares from the regression / total sum of squares

```{r KW, include=FALSE,eval=TRUE,cache=F}
#KW for factors
occ_otu_site_long <- calc_occurrence(Fun, variable = "Site", justdf = T, long=T)
occ_otu_site_kw <- kruskal.test(Occurrence ~ Site, data = occ_otu_site_long)

occ_otu_grass_long <- calc_occurrence(Fun, variable = "Grass", justdf = T, long=T)
occ_otu_grass_kw <- kruskal.test(Occurrence ~ Grass, data = occ_otu_grass_long)

occ_otu_Bin_long <- calc_occurrence(Fun, variable = "Bin", justdf = T, long=T)
occ_otu_Bin_kw <- kruskal.test(Occurrence ~ Bin, data = occ_otu_Bin_long)

# occ_otu_Gradient_long <- calc_occurrence(Fun, variable = "Gradient", justdf = T, long=T)
# occ_otu_Gradient_kw <- kruskal.test(Occurrence ~ Gradient, data = occ_otu_Gradient_long)

occ_otu_Grassland_long <- calc_occurrence(Fun, variable = "Grassland", justdf = T, long=T)
occ_otu_Grassland_kw <- kruskal.test(Occurrence ~ Grassland, data = occ_otu_Grassland_long)
```

```{r eta-sq, include=FALSE,eval=F}
# Calculate Eta-squared
prettyNum(((occ_otu_Bin_kw$statistic/occ_otu_Bin_kw$parameter)*occ_otu_Bin_kw$parameter)/(((occ_otu_Bin_kw$statistic/occ_otu_Bin_kw$parameter)*occ_otu_Bin_kw$parameter)+(509-occ_otu_Bin_kw$parameter+1))*100, digits=4)

prettyNum(((occ_otu_grass_kw$statistic/occ_otu_grass_kw$parameter)*occ_otu_grass_kw$parameter)/(((occ_otu_grass_kw$statistic/occ_otu_grass_kw$parameter)*occ_otu_grass_kw$parameter)+(509-occ_otu_grass_kw$parameter+1))*100, digits=4)

prettyNum(((occ_otu_site_kw$statistic/occ_otu_site_kw$parameter)*occ_otu_site_kw$parameter)/(((occ_otu_site_kw$statistic/occ_otu_site_kw$parameter)*occ_otu_site_kw$parameter)+(509-occ_otu_site_kw$parameter+1))*100, digits=4)

prettyNum(((occ_otu_Gradient_kw$statistic/occ_otu_Gradient_kw$parameter)*occ_otu_Gradient_kw$parameter)/(((occ_otu_Gradient_kw$statistic/occ_otu_Gradient_kw$parameter)*occ_otu_Gradient_kw$parameter)+(509-occ_otu_Gradient_kw$parameter+1))*100, digits=3)

prettyNum(((occ_otu_Grassland_kw$statistic/occ_otu_Grassland_kw$parameter)*occ_otu_Grassland_kw$parameter)/(((occ_otu_Grassland_kw$statistic/occ_otu_Grassland_kw$parameter)*occ_otu_Grassland_kw$parameter)+(509-occ_otu_Grassland_kw$parameter+1))*100, digits=4)

```

```{r perform-anova-for-taxa-function, include=FALSE,eval=F}
perform_anova_for_taxa <- function(phylo_obj, taxa_of_interest, ContinuousVar) {
  # Subset the  phyloseq object to include only the specified taxa_of_interest 
  subset_phylo_obj <- subset_taxa(phylo_obj, Genus == taxa_of_interest)
  #glom instead as input
  subset_phylo_glom <- tax_glom(subset_phylo_obj, taxrank = "Genus")

  #convert phylo object to presence/absence
  phylo_obj_presence <- convert_phylo_to_presence_absence(subset_phylo_glom)
  
  # Convert phyloseq object to data frames
  otu_df <- phylo2vegan_OTU(phylo_obj_presence)
  sample_data_df <- phylo2vegan_sd(phylo_obj_presence)

    # Subset the sample data to include only the specified taxa_of_interest and the ContinuousVar
  subset_data <- subset(sample_data_df, select = c(ContinuousVar))

  # Perform ANOVA for the specified taxa_of_interest against the ContinuousVar
  anova_result <- aov(otu_df ~ subset_data[[ContinuousVar]])
  tidy_aov <- tidy(anova_result)
  tidy_df <- as.data.frame(tidy_aov)
  
  tidy_df$term[1] <- paste0(ContinuousVar)

  sum_squares_regression <- tidy_df$sumsq[1]
  sum_squares_residuals <- tidy_df$sumsq[2]

  R_squared <- sum_squares_regression /
            (sum_squares_regression + sum_squares_residuals)*100

  tidier_df <- data.frame(statistic = numeric(),
                          R_squared = numeric(),p.value = character(),
                          stringsAsFactors = FALSE)
  
  #tidier_df[1,"predictor"] <- tidy_df$term[1]
  tidier_df[1,"statistic"] <- tidy_df$statistic[1]
  tidier_df[1,"R_squared"] <- R_squared
  tidier_df[1,"p.value"] <- tidy_df$p.value[1]

  # Return ANOVA results for the specified taxa_of_interest
  return(tidier_df)
}
```

```{r perform-anova-for-taxa, include=FALSE,eval=F}
# aov for continuous

# numeric metadata
#coll_GDD,GDD30yr,GDD3yr,GDD2015,ppt30yr,ppt3yr,ppt2015,soil_moisture,GWC,perc_moisture,pH,phos,ammonium,nitrate,SOM,herbivory_perc,avg_SLA,avg_SRL
ContinuousVar.list <- c("Latitude","Longitude","Elevation_m",
                        "GDD3yr_m.std","ppt3yr_m.std","soil_moisture_m.std",
                        "pH_m.std","phos_m.std","ammonium_m.std",
                        "nitrate_m.std","SOM_m.std","herbivory_perc_m.std",
                        "avg_SLA_m.std","avg_SRL_m.std")

tidy_aov_for_taxa_ContinuousVar.list = vector("list", length = length(ContinuousVar.list))

TAXA<-c("Darksidea")

for (ContinuousVar in ContinuousVar.list) {
  
  #current_variable <- ContinuousVar.list[i]
  
  tidy_aov_for_taxa_ContinuousVar <- perform_anova_for_taxa(Fun, "Darksidea", ContinuousVar)
  
  tidy_aov_for_taxa_ContinuousVar$ContinuousVar <- ContinuousVar 
  
  tidy_aov_for_taxa_ContinuousVar.list[[ContinuousVar]] <- (as.data.frame(tidy_aov_for_taxa_ContinuousVar))

  }

tidy_aov_for_taxa_ContinuousVar_out = do.call(rbind, tidy_aov_for_taxa_ContinuousVar.list)
tidy_aov_for_taxa_ContinuousVar_out$p.value <- as.numeric(tidy_aov_for_taxa_ContinuousVar_out$p.value)

tidy_aov_for_taxa_ContinuousVar_out$Significance <- ifelse(tidy_aov_for_taxa_ContinuousVar_out$p.value > 0.05, "NS", as.character(tidy_aov_for_taxa_ContinuousVar_out$p.value))

tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar <- as.factor(tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar)

prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="Latitude",]$R_squared, digits=3)
prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="ppt3yr_m.std",]$p.value, digits=3)

saveRDS(tidy_aov_for_taxa_ContinuousVar_out, file = paste0(OAF_PATHDIR,"/",PROJABBV,"_tidy_aov_for_taxa_ContinuousVar_out.rds"))
```

```{r, include=F,eval=T}
# tidy_aov_for_taxa_ContinuousVar_out <- readRDS(file = paste0(OAF_PATHDIR,"/",PROJABBV,"_tidy_aov_for_taxa_ContinuousVar_out.rds"))

tidy_aov_for_taxa_ContinuousVar_out <- readRDS(file = "../docs/processed_data/OTU_occurrencefrequency/tidy_aov_for_taxa_ContinuousVar_out.rds")
```

*Darksidea* OTU occurrence frequency significantly influenced by

[*Sampling Variables:*]{.underline}

-   **Site**: (KW; $\eta^2$ = `r prettyNum(((occ_otu_site_kw$statistic/occ_otu_site_kw$parameter)*occ_otu_site_kw$parameter)/(((occ_otu_site_kw$statistic/occ_otu_site_kw$parameter)*occ_otu_site_kw$parameter)+(509-occ_otu_site_kw$parameter+1))*100, digits=4)`, p-value \< 2.2e-16, $\chi^2$ = `r prettyNum(occ_otu_site_kw$statistic, digits=4)`)
-   **Grassland type**: (KW; $\eta^2$ = `r prettyNum(((occ_otu_Grassland_kw$statistic/occ_otu_Grassland_kw$parameter)*occ_otu_Grassland_kw$parameter)/(((occ_otu_Grassland_kw$statistic/occ_otu_Grassland_kw$parameter)*occ_otu_Grassland_kw$parameter)+(509-occ_otu_Grassland_kw$parameter+1))*100, digits=4)`, p-value = `r prettyNum(occ_otu_Grassland_kw$p.value, digits=3)`, $\chi^2$ = `r prettyNum(occ_otu_Grassland_kw$statistic, digits=4)`)
-   **Grass host**: (KW; $\eta^2$ = `r prettyNum(((occ_otu_grass_kw$statistic/occ_otu_grass_kw$parameter)*occ_otu_grass_kw$parameter)/(((occ_otu_grass_kw$statistic/occ_otu_grass_kw$parameter)*occ_otu_grass_kw$parameter)+(509-occ_otu_grass_kw$parameter+1))*100, digits=4)`, p-value = `r prettyNum(occ_otu_grass_kw$p.value, digits=3)`, $\chi^2$ = `r prettyNum(occ_otu_grass_kw$statistic, digits=4)`)
-   **Latitudinal Bin**: (KW; $\eta^2$ = `r prettyNum(((occ_otu_Bin_kw$statistic/(length(unique(occ_otu_Bin_long$Bin)) - 1))*(length(unique(occ_otu_Bin_long$Bin)) - 1))/((occ_otu_Bin_kw$statistic/(length(unique(occ_otu_Bin_long$Bin)) - 1))*(length(unique(occ_otu_Bin_long$Bin)) - 1)+(509-length(unique(occ_otu_Bin_long$Bin))))*100, digits=3)`, p-value = `r prettyNum(occ_otu_Bin_kw$p.value, digits=4)`, $\chi^2$ = `r prettyNum(occ_otu_Bin_kw$statistic, digits=4)`)
-   **Elevation** (in meters): (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="Elevation_m",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="Elevation_m",]$p.value, digits=3)`)
-   **Latitude**: (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="Latitude",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="Latitude",]$p.value, digits=3)`)
-   **Longitude:** (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="Longitude",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="Longitude",]$p.value, digits=3)`)

[*Climate Data:*]{.underline}

-   **GDD** = growing degree days: (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="GDD3yr_m.std",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="GDD3yr_m.std",]$p.value, digits=3)`)
-   **ppt** = mean annual precipitation: (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="ppt3yr_m.std",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="ppt3yr_m.std",]$p.value, digits=3)`)

[*Edaphic Variables:*]{.underline}

-   **soil_moisture**: (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="soil_moisture_m.std",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="soil_moisture_m.std",]$p.value, digits=2)`)
-   **ammonium**: (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="ammonium_m.std",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="ammonium_m.std",]$p.value, digits=3)`)
-   **nitrate**: (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="nitrate_m.std",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="nitrate_m.std",]$p.value, digits=3)`)
-   **phosphorous**: (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="phos_m.std",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="phos_m.std",]$p.value, digits=2)`)
-   ~~**SOM**~~ (via LoI): *NS*
-   ~~**pH**~~: *NS*

[*Plant Traits:*]{.underline}

-   **avg_SRL**: (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="avg_SRL_m.std",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="avg_SRL_m.std",]$p.value, digits=3)`)
-   **herbivory_perc**: (AOV; $R^2$ = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="herbivory_perc_m.std",]$R_squared, digits=3)`, p-value = `r prettyNum(tidy_aov_for_taxa_ContinuousVar_out[tidy_aov_for_taxa_ContinuousVar_out$ContinuousVar=="herbivory_perc_m.std",]$p.value, digits=3)`)
-   ~~**avg_SLA**~~: *NS*

# IGNORE/OLD {.hidden .unlisted .unnumbered}

```{r, include=F, eval=F}
# create container array (dim = c(nrow, ncol, n_df))
adonis.out_predGrass.list = vector("list", length = length(comm.names.by.grass[1:2]))

for (comm in comm.names.by.grass[1:2]) {
  adonis.temp_predGrass <- adonis2(phyloseq::distance(get(comm), method="bray") ~ Grass, 
          data = (as(sample_data(get(comm)), "data.frame")))
  
  adonis.temp_predGrass$comm <- comm 
  
  adonis.out_predGrass.list[[comm]] <- (as.data.frame(adonis.temp_predGrass))

  }

adonis.out.array_predGrass = do.call(rbind, adonis.out_predGrass.list)

Fun.OTU.glom_Darksidea <- Fun %>% 
  transform_sample_counts(function(x) {x / sum(x)} ) %>% 
  subset_taxa(Genus=="Darksidea")

Fun.OTU.glom_Darksidea_dat <- Fun %>% 
  transform_sample_counts(function(x) {x / sum(x)} ) %>% 
  subset_taxa(Genus=="Darksidea") %>% 
  psmelt()

Fun.Genus.glom_Darksidea <- subset_taxa(Fun.Genus.r, Genus=="Darksidea")

Fun.OTU.glom_Darksidea2 <- prune_taxa(taxa_sums(Fun.OTU.glom_Darksidea) > 0, Fun.OTU.glom_Darksidea)
otu_table(Fun.OTU.glom_Darksidea2)
Fun.Genus.glom_Darksidea2 <- prune_taxa(taxa_sums(Fun.Genus.glom_Darksidea) > 0, Fun.Genus.glom_Darksidea)
adonis2(phyloseq::distance(Fun.Genus.glom_Darksidea, method="bray") ~ Grass, 
          data = (as(sample_data(Fun.Genus.glom_Darksidea), "data.frame")))


```

```{r, include=FALSE,eval=FALSE}
knitr::kable((as.data.frame(subset(adonis.out.array_predGrass, 
                    comm==comm.names.by.grass[1])))[,1:5], "simple", 
             digits =c (0,1,3,2,3), 
             align = 'c')

knitr::kable((as.data.frame(subset(adonis.out.array_predGrass, 
                    comm==comm.names.by.grass[2])))[,1:5], "simple", 
             digits =c (0,1,3,2,3), 
             align = 'c')

```

```{r, include=FALSE,eval=FALSE}
# Histogram overlaid with kernel density curve
ggplot(Fun.OTU.glom_Darksidea_dat, aes(x=Occurrence, fill=Phylum)) + 
    geom_histogram(aes(y=after_stat(density)),      # Histogram with density instead of count on y-axis
                   binwidth=.001,
                   position="identity") +
    geom_density(alpha=.2) + # Overlay with transparent density plot
  # scale_x_log10()+
  geom_vline(aes(xintercept=mean(Occurrence, na.rm=T)),   # Ignore NA values for mean
               color="red", linetype="dashed", linewidth=1)


```

```{r, include=F,eval=F}
# intersect(fun_AllGrasses_OA$OTU,fun_OTU_spegen_AllG_res$OTU))



fun_AllGrasses_tax_results1<-cbind(fun_AllGrasses_OTU3, fun_AllGrasses_OA)
colnames(fun_AllGrasses_tax_results1)[8] <- "Abundance_count"
fun_AllGrasses_tax_results2<-subset(fun_AllGrasses_tax_results1, select=-c(Abundance_count,Occurrence))
fun_AllGrasses_tax_table.mat<-as.matrix(fun_AllGrasses_tax_results2)
fun_AllGrasses_tax_table<-tax_table(fun_AllGrasses_tax_table.mat)


# Convert to relative abundance
fun_OTU_relabundA <- Fun %>% 
        transform_sample_counts(function(x) {x / sum(x)} ) %>%
        psmelt() %>% select(OTU, Sample, Site, Grass, Latitude, Longitude, 
                            Grassland, Gradient, Bin, Abundance) %>% 
        mutate(Abundance = Abundance*100) 

fun_OTU_relabund <- fun_OTU_relabundA %>% 
        spread(Sample, Abundance) 

# Convert to presence/absence 
fun_OTU_pa <- fun_OTU_relabund
fun_OTU_pa[,-(1:8)]=sapply(fun_OTU_pa[,-(1:8)],function(x) replace(x,x>0,1))

# Calculate frequency of occurrence (presence) for each OTU
fun_OTU_occurrence_frequency <- rowSums(fun_OTU_PA_mat)

# Create a copy of the OTU matrix
fun_OTU_mat <- phylo2vegan_OTU(Fun)





# Add frequency of occurrence as a column to the relative abundance data
fun_relabund_OTU <- fun_relabund_OTU %>% 
  left_join(data.frame(OTU = rownames(fun_OTU_mat), Frequency = fun_OTU_occurrence_frequency),
            by = "OTU")


# Merge relative abundance data with taxonomy from 'Fun'
fun_OTU_RA_PA_freq_psmelt <- merge(
  fun_relabund_OTU,
  as.data.frame(sample_data(Fun)),
  by = "Sample",
  all.x = TRUE
) %>%
  merge(
    as.data.frame(tax_table(Fun)),
    by = "OTU",
    all.x = TRUE
  )


# Cleaning up merged data
rownames(fun_OTU_RA_PA_freq_psmelt) <- fun_OTU_RA_PA_freq_psmelt$OTU
fun_OTU_RA_PA_freq_psmelt$OTU <- NULL
fun_OTU_RA_PA_freq_psmelt$Sample <- NULL
```
