---
title: '**RAM paired amplicons taxa queries:** Distribution and diversity of *Darksidea* and *Monosporascus* sp. with variable input data'
author: "Kelli Feeser"
subtitle: "`r params$INPUT_DATA_DESCR` "
date: "`r Sys.Date()`"
params:
  PROJABBV: "allSamplesincEDGE"
  INPUT_DATA_DESCR: "Input data NOT subset. All samples to include a) include EDGE samples, and b) include fungal samples with [m]{.underline}issing bacterial [p]{.underline}air ('allSamplesincEDGE')"
output:
  bookdown::html_document2:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 3
    fig.caption: yes
    keep_md: yes
  html_notebook:
    code_folding: hide
    df_print: paged
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 3
    fig.caption: yes
  prettydoc::html_pretty:
    toc: yes
    fig.caption: yes
editor_options:
  chunk_output_type: inline
---

\

::: homelink
<a href="https://kellifeeser.github.io/ksu-paired-amplicon-workflow/index.html" target="_blank" style="text-align:right">Back to Home</a>
:::

\

------------------------------------------------------------------------

Document last updated: `r Sys.Date()`

------------------------------------------------------------------------

\

# Set-up project naming from reproducible code {.unlisted .unnumbered}

```{r project-naming, include=F, eval=T}
PROJECT_PAGENAME <- 'RAM paired amplicons'
PROJECT_SUBPAGENAME <- 'Queries of taxa of interest with variable input data'
INPUT_DATA_DESCR <- "Input data subset to a) include EDGE samples, and b) include fungal samples with [m]{.underline}issing bacterial pair ('allSamplesincEDGE')"
PROJABBV <- 'allSamplesincEDGE'

bindingIsLocked("params", env = .GlobalEnv)
unlockBinding("params", env = .GlobalEnv)
params$INPUT_DATA_DESCR <- INPUT_DATA_DESCR

# DATASET1 <- 'Bac'
# DATASET2 <- 'Fun'

# RDS1 <- 'Fun_wholecommunity.rds'
# RDS2 <- 'Fun_min2_rare10116_v2.rds' # n=509 samples, yes MP no EDGE
RDS3 <- 'Fun_min2_rare10116_v1.rds' # n=654 samples, yes MP yes EDGE

#create paths for directories (sans prefix)
FIG_PATHDIR=paste0("../docs/figures/darksidea_plots/",PROJABBV)
CACHE_PATHDIR=paste0("../docs/github_darksidea_cache/github_darksidea_",PROJABBV,"_files/")
# occupancy/abundance/tax frequencies
OAF_PATHDIR=paste0("../docs/processed_data/OTU_occurrencefrequency/",PROJABBV)
OAF_PATHDIR2=paste0("../processed_data/OTU_occurrencefrequency/",PROJABBV)


FIG_PATH=paste0("../docs/figures/darksidea_plots/",PROJABBV,"/RAM_",PROJABBV,"-")
CACHE_PATH=paste0("../docs/github_darksidea_cache/github_darksidea_",PROJABBV,"_files/RAM_",PROJABBV,"_")
# occupancy/abundance/tax frequencies
OAF_PATH=paste0("../docs/processed_data/OTU_occurrencefrequency/",PROJABBV,"/RAM_",PROJABBV,"_Freq_OTU_Occurrence_")
OAF_PATH2=paste0("../processed_data/OTU_occurrencefrequency/",PROJABBV,"/RAM_",PROJABBV,"_Freq_OTU_Occurrence_")


OAF_PATH_GENUS=paste0("../docs/processed_data/OTU_occurrencefrequency/",PROJABBV,"/byGenus","/RAM_",PROJABBV,"_Freq_Genus_Occurrence_")
```

```{r global-options, include = FALSE}
library(knitr)

# set up global R options
  # to list the current values of all options but better to F1
    #options()
    # easier way:
      #?options
  # to access the value of a single option
    #getOption("width")

# figure guidelines
  # save as vector (.ai, .eps, .pdf)
  # resolution: min 300 dpi
  # dimensions: 
    # Nature Research Journals
      # 1 col width = 88mm (3.45 in)
      # 2 col width = 180mm (7.08 in)
    # All Other Journals
      # 1 col width = 85mm (3.34 in)
      # 2 col width = 175mm (6.88 in)
  # file size should be < 50 MB

# text
  # font: sans-serif (helvetica or arial)
  # size: min 5pt, max 7pt

## ~~create directories~~ ##
###########################################
dir.create(FIG_PATHDIR, showWarnings = T,recursive=T)
dir.create(CACHE_PATHDIR, showWarnings = T,recursive=T)
dir.create(OAF_PATHDIR, showWarnings = T,recursive=T)
dir.create(OAF_PATHDIR2, showWarnings = T,recursive=T)
###########################################


# set up knitr global chunk options
  # See str(knitr::opts_chunk$get()) for a list of default chunk options.
#knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(echo = FALSE,
                      fig.path = FIG_PATH,
                      cache.path = CACHE_PATH,
                      eval.after = "fig.cap",
                      fig.pos = '!h'
                      )
```

```{r LANL-proxies, eval=T, include=FALSE}
# LANL proxies
Sys.setenv('HTTP_PROXY'='http://proxyout.lanl.gov:8080')
Sys.setenv('HTTPS_PROXY'='http://proxyout.lanl.gov:8080')
```

# Set-up {.unlisted .unnumbered .hidden}

## load packages

```{r load packages, eval=T, include=F}
library("ggplot2"); packageVersion("ggplot2") #‘3.4.4’
library("vegan"); packageVersion("vegan") # # ‘2.6.4’
library("phyloseq"); packageVersion("phyloseq") #‘1.44.0’
library("scales"); packageVersion("scales") #‘1.2.1’
library("swatches"); packageVersion("swatches") #‘0.5.0’
library("ggsci"); packageVersion("ggsci") #‘3.0.0’
library("ggpubr"); packageVersion("ggpubr") #‘0.6.0’
library("viridis"); packageVersion("viridis") #‘0.6.4’
library("RColorBrewer"); packageVersion("RColorBrewer") #‘1.1.3’
library("cowplot"); packageVersion("cowplot") #‘1.1.1’
library("dplyr"); packageVersion("dplyr") #‘1.1.3’
library("plyr"); packageVersion("plyr") #‘1.8.9’
library("knitr"); packageVersion("knitr") #‘1.44’
#install.packages('kableExtra', dependencies=TRUE, repos='http://cran.rstudio.com/')
#library("kableExtra"); packageVersion("kableExtra") #‘1.3.4’
library("DT"); packageVersion("DT") #‘0.30’
library("tidyverse"); packageVersion("tidyverse") #‘2.0.0’
library("prettydoc"); packageVersion("prettydoc") #‘0.4.1’
library("magrittr"); packageVersion("magrittr") #‘2.0.3’
library("broom"); packageVersion("broom") #‘1.0.5’
library("kableExtra")

# library("devtools")
# devtools::install_github("vmikk/metagMisc")
# library("metagMisc"); packageVersion("metagMisc") #‘2.0.3’



  # if (!require("BiocManager", quietly = TRUE))
  #     install.packages("BiocManager")
  # BiocManager::install(version = "3.11")
  # 
  # BiocManager::install("microbiome")
library("microbiome"); packageVersion("microbiome") #‘1.30.0’

source("ksu_colors.R")
source("../../black-sheep-club/code/sheepish_scripts.R")
```

## load rds

```{r load-rds, include=FALSE,eval=TRUE}
###.###.######.###.###
# phylo #
###.###.###
# Fun<-readRDS(file = paste0("../processed_data/clean_rds_saves/",RDS3))
Fun<-readRDS(file = "../processed_data/clean_rds_saves/ITS_min2_rare10116_v1.rds")
Fun

colnames(sample_data(Fun))[8] <- "Grass"
levels(sample_data(Fun)$Grass)

levels(sample_data(Fun)$Site)
sample_data(Fun)$Site <- 
revalue(sample_data(Fun)$Site, c("unknown" = "SFA","CRP" = "CPR","CDA" = "CAD"))

###.###.######.###.###
```

(ref:OA-AllG-allOTUs-fig-caption) Occupancy-abundance figure overlayed with generalist framework results from OTU down-selection for co-occurrence (top 2.5% of distribution). Plot contains data from all samples and OTUs. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data. Note both axes are shown on a log scale.

(ref:OA-BOGR-allOTUs-fig-caption) Occupancy-abundance figure overlayed with generalist framework results from OTU down-selection for co-occurrence (top 2.5% of distribution). Plot contains data from all OTUs associated with *Blue grama* hosts. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data. Note both axes are shown on a log scale.

(ref:OA-BOER-allOTUs-fig-caption) Occupancy-abundance figure overlayed with generalist framework results from OTU down-selection for co-occurrence (top 2.5% of distribution). Plot contains data from all OTUs associated with *Black grama* hosts. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data. Note both axes are shown on a log scale.

(ref:OA-BUDA-allOTUs-fig-caption) Occupancy-abundance figure overlayed with generalist framework results from OTU down-selection for co-occurrence (top 2.5% of distribution). Plot contains data from all OTUs associated with *Buffalograss* hosts. I have added in the generalist vs. specialist (GvS) OTU annotations. 'Generalist' OTUs have the highest 2.5% of occurrence:abundance distribution, 'specialists' are in the lowest lowest 2.5%, 'NS' are the non-significant OTUs falling in the middle 95%, 'singletons' are OTUs that only appeared in a single sample and thus were excluded from GvS analyses. GvS calculations were run independently for the full dataset (all grass hosts) and for the by-host data. Note both axes are shown on a log scale.

# Data Overview

`r INPUT_DATA_DESCR`

------------------------------------------------------------------------

### OTU tables {.unnumbered}

```{r make-Darksidea-stats-Fun_Genus_Mean_Abund, eval=T, include=F, cache=T}
# dat_Genus.r <- Fun %>%
#   tax_glom(taxrank = "Genus") %>%
#   transform_sample_counts(function(x) {x / sum(x)} ) %>%
#   psmelt()
# dat_Genus.r$Genus <- factor(dat_Genus.r$Genus)
# saveRDS(dat_Genus.r, file="../processed_data/clean_rds_saves/dat_Genus.r_allSamplesincEDGE.rds")

dat_Genus.r <- readRDS(file="../processed_data/clean_rds_saves/dat_Genus.r_allSamplesincEDGE.rds")
dat_Genus.r$Genus <- factor(dat_Genus.r$Genus)

Darksidea_Genus_stats <- dat_Genus.r %>% 
  as_tibble %>%
  group_by(Genus) %>%
  reframe(
  av = mean(Abundance)*100,
  sd = sd(Abundance,na.rm = T)*100) %>% 
  filter(Genus=="Darksidea")

# dat_OTU.r <- Fun %>%
#   transform_sample_counts(function(x) {x / sum(x)} ) %>%
#   psmelt()
# dat_OTU.r$Genus <- factor(dat_OTU.r$Genus)
# dat_OTU.r$OTU <- factor(dat_OTU.r$OTU)

# saveRDS(dat_OTU.r, file="../processed_data/clean_rds_saves/dat_OTU.r_allSamplesincEDGE.rds")

dat_OTU.r <- readRDS(file="../processed_data/clean_rds_saves/dat_OTU.r_allSamplesincEDGE.rds")

Darksidea_OTU_stats <- dat_OTU.r %>% 
  as_tibble %>%
  filter(Genus=="Darksidea") %>% 
  group_by(OTU) %>%
  reframe(
  av = mean(Abundance)*100,
  sd = sd(Abundance,na.rm = T)*100) 

###############
# Monosporascus
Monosporascus_Genus_stats <- dat_Genus.r %>% 
  as_tibble %>%
  group_by(Genus) %>%
  reframe(
  av = mean(Abundance)*100,
  sd = sd(Abundance,na.rm = T)*100) %>% 
  filter(Genus=="Monosporascus")

Monosporascus_OTU_stats <- dat_OTU.r %>% 
  as_tibble %>%
  filter(Genus=="Monosporascus") %>% 
  group_by(OTU) %>%
  reframe(
  av = mean(Abundance)*100,
  sd = sd(Abundance,na.rm = T)*100) 
###############
# sample_data(Fun)$merge_col <- rep("merge",nrow(sample_data(Fun)))
# Fun_merged <- merge_samples(Fun, "merge_col",fun=mean)
# Fun_merged.Genus.A <- tax_glom(Fun_merged, taxrank="Genus")
# Fun_merged.Genus <- prune_taxa(taxa_sums(Fun_merged.Genus.A) > 0, Fun_merged.Genus.A)
# Fun_merged.Genus.r  = transform_sample_counts(Fun_merged.Genus, function(x) x / sum(x) )
  #767

# saveRDS(Fun_merged.Genus.r, file="../processed_data/clean_rds_saves/Fun_merged.Genus.r_allSamplesincEDGE.rds")

Fun_merged.Genus.r <- readRDS(file="../processed_data/clean_rds_saves/Fun_merged.Genus.r_allSamplesincEDGE.rds")

sample_names(Fun_merged.Genus.r)<-"Fun_Genus_Mean"
Fun_Genus_Mean_Abund <- merge(tax_table(Fun_merged.Genus.r)[,c("Domain","Phylum","Class","Order","Family","Genus")], t(otu_table(Fun_merged.Genus.r)), by=0, all=TRUE)
colnames(Fun_Genus_Mean_Abund)[1]<-"Fun_OTU"
Fun_Genus_Mean_Abund$Fun_Genus_Mean<-as.numeric(Fun_Genus_Mean_Abund$Fun_Genus_Mean*100)
Fun_Genus_Mean_Abund <-Fun_Genus_Mean_Abund[order(Fun_Genus_Mean_Abund$Fun_Genus_Mean, decreasing = T),]

Fun_Genus_Mean_Abund2 <- Fun_Genus_Mean_Abund 

colnames(Fun_Genus_Mean_Abund2)[8] <- "Mean rel abund (%)"
# Fun_Genus_Mean_Abund2$Rank_RA <- rank(Fun_Genus_Mean_Abund2$`Abundance (%)`,ties.method	= "max")

Fun_Genus_Mean_Abund2$Rank_RA<-(dplyr::dense_rank(dplyr::desc(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`)))

Fun_Genus_Mean_Abund2$Percentile<-(dplyr::percent_rank(dplyr::desc(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`)))*100


# round mean rel abundance
    # Round values less than 0.0001 to 5 decimal places
    # Round values between 0.0001 and 0.001 to 4 decimal places
    # Round values between 0.001 and 0.01 to 3 decimal places
    # Round values between 0.01 and 0.1 to 2 decimal places
    # Round values between 0.1 and 10 to 1 decimal place
    # Round values greater than 10 to 0 decimal places.


Fun_Genus_Mean_Abund2$`Mean rel abund (%)` <- ifelse(
  Fun_Genus_Mean_Abund2$`Mean rel abund (%)` < 0.0001,
  round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 5),
  ifelse(
    Fun_Genus_Mean_Abund2$`Mean rel abund (%)` < 0.001,
    round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 4),
    ifelse(
      Fun_Genus_Mean_Abund2$`Mean rel abund (%)` < 0.01,
      round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 3),
      ifelse(
        Fun_Genus_Mean_Abund2$`Mean rel abund (%)` < 0.1,
        round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 2),
        ifelse(
          Fun_Genus_Mean_Abund2$`Mean rel abund (%)` > 10,
          round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 0),
          round(Fun_Genus_Mean_Abund2$`Mean rel abund (%)`, 1)
        )
      )
    )
  )
)


#Fun_Genus_Mean_Abund2$Percentile <- round(Fun_Genus_Mean_Abund2$Percentile,1)

MAX_RANK_VALUE<-max(Fun_Genus_Mean_Abund2[,9])

colnames(Fun_Genus_Mean_Abund2)[9] <- paste0("Rank_RA (max ",MAX_RANK_VALUE,")")
```

```{r calc_occurrence-data-manipulation, include=F,eval=T,cache=F}

Fun_Sev <-subset_samples(Fun, Site=="SEV")
Fun_Sev_no0x <- prune_taxa(taxa_sums(Fun_Sev) > 0, Fun_Sev)
Fun_Sev_no1x <- prune_taxa(taxa_sums(Fun_Sev) > 1, Fun_Sev)

# taxa_frequency = 
  # "count" - just counts OTU occurrence (number of samples)
  # "percentage" - (default) Frequency of occurrence 
      # proportion of sampling units that contains the OTU 
      # (sample units in which OTU occurred / total # of sample units)
  # "relfreq" - Relative occurrence frequency
      # aka relative frequency of OTU 
      # (frequency of each OTU [see 'percentage'] / sum of the frequency of all OTUs)


occ_otu_null <- calc_occurrence(Fun, justdf = T)
occ_otu_site <- calc_occurrence(Fun, variable = "Site", justdf = T)
occ_otu_grass_allsites <- calc_occurrence(Fun, "Grass", justdf = T)
occ_otu_grass_SEVonly <- calc_occurrence(Fun_Sev_no0x, variable = "Grass", justdf = T)


tax_table1<-as.data.frame(tax_table(Fun))
tax_table <- tibble::rownames_to_column(tax_table1, "OTU")

colnames(occ_otu_site)[1] <- "OTU"

colnames(occ_otu_null)[1] <- "OTU"
colnames(occ_otu_null)[2] <- "global_occurrence"

colnames(occ_otu_grass_allsites)[1] <- "OTU"
colnames(occ_otu_grass_allsites)[2] <- "ANGE_allsites"
colnames(occ_otu_grass_allsites)[3] <- "BOCU_allsites"
colnames(occ_otu_grass_allsites)[4] <- "BOER_allsites"
colnames(occ_otu_grass_allsites)[5] <- "BOGR_allsites"
colnames(occ_otu_grass_allsites)[6] <- "BUDA_allsites"
colnames(occ_otu_grass_allsites)[7] <- "SCSC_allsites"


colnames(occ_otu_grass_SEVonly)[1] <- "OTU"
colnames(occ_otu_grass_SEVonly)[2] <- "BOER_SEVonly"
colnames(occ_otu_grass_SEVonly)[3] <- "BOGR_SEVonly"



OTU_freq_occurr_bySite_Grass<-
  tax_table %>% 
  left_join(occ_otu_null,by = "OTU") %>% 
  left_join(occ_otu_grass_SEVonly,by = "OTU") %>% 
  left_join(occ_otu_grass_allsites,by = "OTU") %>% 
  left_join(occ_otu_site,by = "OTU")

#head(OTU_freq_occurr_bySite_Grass)

OTU_freq_occurr_bySite_Grass_Monosporascus <- subset(OTU_freq_occurr_bySite_Grass, Genus == "Monosporascus")
OTU_freq_occurr_bySite_Grass_Monosporascus$Domain<-as.factor(OTU_freq_occurr_bySite_Grass_Monosporascus$Domain)
OTU_freq_occurr_bySite_Grass_Monosporascus$OTU<-as.factor(OTU_freq_occurr_bySite_Grass_Monosporascus$OTU)

OTU_freq_occurr_bySite_Grass_Xylariales <- subset(OTU_freq_occurr_bySite_Grass, Order == "Xylariales")
OTU_freq_occurr_bySite_Grass_Darksidea <- subset(OTU_freq_occurr_bySite_Grass, Genus == "Darksidea")

# write to docs/processed_data
write.csv(OTU_freq_occurr_bySite_Grass, file=paste0(OAF_PATH,"bySitexGrass_OTU_alltaxa.csv"))

write.csv(OTU_freq_occurr_bySite_Grass_Monosporascus, file=paste0(OAF_PATH,"bySitexGrass_genus_Monosporascus.csv"))

write.csv(OTU_freq_occurr_bySite_Grass_Xylariales, file=paste0(OAF_PATH,"bySitexGrass_order_Xylariales.csv"))

write.csv(OTU_freq_occurr_bySite_Grass_Darksidea, file=paste0(OAF_PATH,"bySitexGrass_genus_Darksidea.csv"))

# write to ../processed_data
# write.csv(OTU_freq_occurr_bySite_Grass, file="../processed_data/OTU_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_OTU_alltaxa.csv")

# write.csv(OTU_freq_occurr_bySite_Grass_Monosporascus, file="../processed_data/OTU_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_genus_Monosporascus.csv")

# write.csv(OTU_freq_occurr_bySite_Grass_Xylariales, file="../processed_data/OTU_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_order_Xylariales.csv")

# write.csv(OTU_freq_occurr_bySite_Grass_Darksidea, file="../processed_data/OTU_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_genus_Darksidea.csv")

```

```{r calc_occurrence-data-manipulation-GENUS, include=F,eval=T,cache=T}
Fun_Genus <- Fun %>% tax_glom(taxrank = "Genus") #%>% 
  # transform_sample_counts(function(x) {x / sum(x)} ) %>%

Fun_GENUS <- prune_taxa(taxa_sums(Fun_Genus) > 0, Fun_Genus)

levels(sample_data(Fun_GENUS)$Site_Cover)
sample_data(Fun_GENUS)$Site_Cover <- 
revalue(sample_data(Fun_GENUS)$Site_Cover, c("unknown_SCSC" = "SFA_SCSC",
                                       "CRP_BOGR_E" = "CPR_BOGR_E",
                                       "CDA_BOGR" = "CAD_BOGR"))


Fun_GENUS.r <- Fun_GENUS %>% 
  transform_sample_counts(function(x) {x / sum(x)} )

Fun_GENUS_Sev <-subset_samples(Fun_GENUS, Site=="SEV")
Fun_GENUS_Sev_no0x <- prune_taxa(taxa_sums(Fun_GENUS_Sev) > 0, Fun_GENUS_Sev)
Fun_GENUS_Sev_no1x <- prune_taxa(taxa_sums(Fun_GENUS_Sev) > 1, Fun_GENUS_Sev)

# taxa_frequency = 
  # "count" - just counts OTU occurrence (number of samples)
  # "percentage" - (default) Frequency of occurrence 
      # proportion of sampling units that contains the OTU 
      # (sample units in which OTU occurred / total # of sample units)
  # "relfreq" - Relative occurrence frequency
      # aka relative frequency of OTU 
      # (frequency of each OTU [see 'percentage'] / sum of the frequency of all OTUs)


occ_GENUS_null <- calc_occurrence(Fun_GENUS, justdf = T)
occ_GENUS_site <- calc_occurrence(Fun_GENUS, variable = "Site", justdf = T)
occ_GENUS_grass_allsites <- calc_occurrence(Fun_GENUS, "Grass", justdf = T)
occ_GENUS_grass_SEVonly <- calc_occurrence(Fun_GENUS_Sev_no0x, variable = "Grass", justdf = T)
occ_GENUS_grass_bysites <- calc_occurrence(Fun_GENUS, "Site_Cover", justdf = T)


# Count number of samples per Site_Cover level
Site_Cover_counts <- phylo2vegan_sd(Fun_GENUS) %>%
  group_by(Site_Cover) %>%
  tally(name = "Sample_Count")


tax_table_GENUS1<-as.data.frame(tax_table(Fun_GENUS))
tax_table_GENUS2 <- tibble::rownames_to_column(tax_table_GENUS1, "Taxa")
# tax_table_GENUS3 <- cbind(tax_table_GENUS2$Taxa,tax_table_GENUS2$Genus,tax_table_GENUS2[,2:6]


colnames(occ_GENUS_site)[1] <- "Taxa"

colnames(occ_GENUS_null)[1] <- "Taxa"
colnames(occ_GENUS_null)[2] <- "global_occurrence"

colnames(occ_GENUS_grass_allsites)[1] <- "Taxa"
colnames(occ_GENUS_grass_allsites)[2] <- "ANGE_allsites"
colnames(occ_GENUS_grass_allsites)[3] <- "BOCU_allsites"
colnames(occ_GENUS_grass_allsites)[4] <- "BOER_allsites"
colnames(occ_GENUS_grass_allsites)[5] <- "BOGR_allsites"
colnames(occ_GENUS_grass_allsites)[6] <- "BUDA_allsites"
colnames(occ_GENUS_grass_allsites)[7] <- "SCSC_allsites"


colnames(occ_GENUS_grass_SEVonly)[1] <- "Taxa"
colnames(occ_GENUS_grass_SEVonly)[2] <- "BOER_SEVonly"
colnames(occ_GENUS_grass_SEVonly)[3] <- "BOGR_SEVonly"

# colnames(occ_GENUS_grass_bysites)

GENUS_freq_occurr_bySite_Grass1<-
  tax_table_GENUS2 %>% 
  left_join(occ_GENUS_null,by = "Taxa") %>% 
  left_join(occ_GENUS_grass_SEVonly,by = "Taxa") %>% 
  left_join(occ_GENUS_grass_allsites,by = "Taxa") %>% 
  left_join(occ_GENUS_site,by = "Taxa") %>% 
  left_join(occ_GENUS_grass_bysites,by = "Taxa")

#head(GENUS_freq_occurr_bySite_Grass1)
colnames(GENUS_freq_occurr_bySite_Grass1)

GENUS_freq_occurr_bySite_Grass<-cbind("Genus"=GENUS_freq_occurr_bySite_Grass1$Genus,"Family"=GENUS_freq_occurr_bySite_Grass1$Family,"Order"=GENUS_freq_occurr_bySite_Grass1$Order,"Class"=GENUS_freq_occurr_bySite_Grass1$Class,"Phylum"=GENUS_freq_occurr_bySite_Grass1$Phylum,"Domain"=GENUS_freq_occurr_bySite_Grass1$Domain,GENUS_freq_occurr_bySite_Grass1[,9:109])

GENUS_freq_occurr_bySite_Grass_Monosporascus <- subset(GENUS_freq_occurr_bySite_Grass, Genus == "Monosporascus")
# GENUS_freq_occurr_bySite_Grass_Monosporascus$Domain<-as.factor(GENUS_freq_occurr_bySite_Grass_Monosporascus$Domain)
# GENUS_freq_occurr_bySite_Grass_Monosporascus$GENUS<-as.factor(GENUS_freq_occurr_bySite_Grass_Monosporascus$GENUS)

GENUS_freq_occurr_bySite_Grass_Xylariales <- subset(GENUS_freq_occurr_bySite_Grass, Order == "Xylariales")
GENUS_freq_occurr_bySite_Grass_Darksidea <- subset(GENUS_freq_occurr_bySite_Grass, Genus == "Darksidea")

# write to docs/processed_data
write.csv(GENUS_freq_occurr_bySite_Grass, file=paste0(OAF_PATH_GENUS,"bySitexGrass_GENUS_alltaxa.csv"))

write.csv(GENUS_freq_occurr_bySite_Grass_Monosporascus, file=paste0(OAF_PATH_GENUS,"bySitexGrass_genus_Monosporascus.csv"))

write.csv(GENUS_freq_occurr_bySite_Grass_Xylariales, file=paste0(OAF_PATH_GENUS,"bySitexGrass_order_Xylariales.csv"))

write.csv(GENUS_freq_occurr_bySite_Grass_Darksidea, file=paste0(OAF_PATH_GENUS,"bySitexGrass_genus_Darksidea.csv"))

# write to ../processed_data
# write.csv(GENUS_freq_occurr_bySite_Grass, file="../processed_data/GENUS_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_GENUS_Occurrence_bySitexGrass_GENUS_alltaxa.csv")

# write.csv(GENUS_freq_occurr_bySite_Grass_Monosporascus, file="../processed_data/GENUS_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_GENUS_Occurrence_bySitexGrass_genus_Monosporascus.csv")

# write.csv(GENUS_freq_occurr_bySite_Grass_Xylariales, file="../processed_data/GENUS_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_GENUS_Occurrence_bySitexGrass_order_Xylariales.csv")

# write.csv(GENUS_freq_occurr_bySite_Grass_Darksidea, file="../processed_data/GENUS_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_GENUS_Occurrence_bySitexGrass_genus_Darksidea.csv")

```

```{r calc_relabund-data-manipulation-GENUS, include=F,eval=T,cache=T}
RAF_PATH_GENUS <- "../docs/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus/RAM_allSamplesincEDGE_Genus_RelativeAbundance_"

Fun_GENUS.r <- transform_sample_counts(Fun_GENUS, function(x) 100 * x / sum(x))


# Extract relative abundance table
otu_rel <- as.data.frame(otu_table(Fun_GENUS.r))
if (!taxa_are_rows(Fun_GENUS.r)) {
  otu_rel <- t(otu_rel)
}
otu_rel <- tibble::rownames_to_column(otu_rel, "Taxa")

# ---- Global relative abundance ----
otu_df <- otu_rel %>%
  column_to_rownames("Taxa") %>%
  t() %>%
  as.data.frame()
otu_df$Sample <- rownames(otu_df)

meta_df <- as(sample_data(Fun_GENUS.r), "data.frame")
if (!"Sample" %in% colnames(meta_df)) {
  meta_df <- tibble::rownames_to_column(meta_df, "Sample")
}

# Join on Sample
otu_meta <- left_join(otu_df, meta_df, by = "Sample")

# Now get global percent abundance
otu_rel_numeric <- otu_rel %>% column_to_rownames("Taxa")
otu_rel_global <- data.frame(
  Taxa = rownames(otu_rel_numeric),
  global_relative_abundance = rowMeans(otu_rel_numeric, na.rm = TRUE),
  row.names = NULL
)

otu_rel_global[, -1] <- round(otu_rel_global[, -1], 3)



# ---- Relative abundance by Site (across all Grasses) ----
# Compute mean relative abundance per Genus by Site
site_groups <- split(otu_meta, otu_meta$Site)

otu_rel_site <- lapply(site_groups, function(df) {
  df_genus <- df[, !(colnames(df) %in% colnames(meta_df))]  # drop all metadata
  colMeans(df_genus[, setdiff(names(df_genus), "Sample"), drop = FALSE], na.rm = TRUE)
}) %>%
  do.call(cbind, .) %>%
  as.data.frame()

otu_rel_site$Taxa <- rownames(otu_rel_site)
otu_rel_site <- otu_rel_site[, c("Taxa", sort(setdiff(names(otu_rel_site), "Taxa")))]

otu_rel_site[, -1] <- round(otu_rel_site[, -1], 3)



# ---- Relative abundance by Grass (across all sites) ----
grass_groups <- split(otu_meta, otu_meta$Grass)

otu_rel_grass <- lapply(grass_groups, function(df) {
  df_genus <- df[, !(colnames(df) %in% colnames(meta_df))]  # drop all metadata
  colMeans(df_genus[, setdiff(names(df_genus), "Sample"), drop = FALSE], na.rm = TRUE)
}) %>%
  do.call(cbind, .) %>%
  as.data.frame()

otu_rel_grass$Taxa <- rownames(otu_rel_grass)
otu_rel_grass <- otu_rel_grass[, c("Taxa", sort(setdiff(names(otu_rel_grass), "Taxa")))]

otu_rel_grass[, -1] <- round(otu_rel_grass[, -1], 3)


# ---- Relative abundance by Site x Grass ----
Site_Cover_groups <- split(otu_meta, otu_meta$Site_Cover)

otu_rel_Site_Cover <- lapply(Site_Cover_groups, function(df) {
  df_genus <- df[, !(colnames(df) %in% colnames(meta_df))]  # drop all metadata
  colMeans(df_genus[, setdiff(names(df_genus), "Sample"), drop = FALSE], na.rm = TRUE)
}) %>%
  do.call(cbind, .) %>%
  as.data.frame()

otu_rel_Site_Cover$Taxa <- rownames(otu_rel_Site_Cover)
otu_rel_Site_Cover <- otu_rel_Site_Cover[, c("Taxa", sort(setdiff(names(otu_rel_Site_Cover), "Taxa")))]

otu_rel_Site_Cover[, -1] <- round(otu_rel_Site_Cover[, -1], 3)







# ---- Relative abundance by Grass within SEV only ----
# Subset metadata to SEV samples
meta_df_sev <- meta_df[meta_df$Site == "SEV", ]

# Subset abundance table to SEV samples
otu_df_sev <- otu_df[otu_df$Sample %in% meta_df_sev$Sample, ]

# Join SEV-only abundance with SEV metadata
otu_meta_sev <- left_join(otu_df_sev, meta_df_sev, by = "Sample")

# Split by Grass type
grass_groups_sev <- split(otu_meta_sev, otu_meta_sev$Grass)

# Calculate relative abundance (mean) by Grass group in SEV only
otu_rel_grass_sev <- lapply(grass_groups_sev, function(df) {
  df_genus <- df[, !(colnames(df) %in% colnames(meta_df_sev))]  # drop metadata
  colMeans(df_genus[, setdiff(names(df_genus), "Sample"), drop = FALSE], na.rm = TRUE)
}) %>%
  do.call(cbind, .) %>%
  as.data.frame()

otu_rel_grass_sev$Taxa <- rownames(otu_rel_grass_sev)
otu_rel_grass_sev <- otu_rel_grass_sev[, c("Taxa", sort(setdiff(names(otu_rel_grass_sev), "Taxa")))]

# Identify columns (excluding "Taxa") that contain any NaN
cols_with_nan <- sapply(otu_rel_grass_sev[, -1], function(x) any(is.nan(x)))

# Drop those columns
otu_rel_grass_sev <- otu_rel_grass_sev[, c(TRUE, !cols_with_nan)]

otu_rel_grass_sev[, -1] <- round(otu_rel_grass_sev[, -1], 3)



# ---- Merge all into full Genus-level table ----
colnames(otu_rel_grass)[1] <- "Taxa"
colnames(otu_rel_grass)[2] <- "ANGE_allsites"
colnames(otu_rel_grass)[3] <- "BOCU_allsites"
colnames(otu_rel_grass)[4] <- "BOER_allsites"
colnames(otu_rel_grass)[5] <- "BOGR_allsites"
colnames(otu_rel_grass)[6] <- "BUDA_allsites"
colnames(otu_rel_grass)[7] <- "SCSC_allsites"


colnames(otu_rel_grass_sev)[1] <- "Taxa"
colnames(otu_rel_grass_sev)[2] <- "BOER_SEVonly"
colnames(otu_rel_grass_sev)[3] <- "BOGR_SEVonly"



GENUS_relabund_bySite_Grass1<-
  tax_table_GENUS2 %>% 
  left_join(otu_rel_global,by = "Taxa") %>% 
  left_join(otu_rel_grass_sev,by = "Taxa") %>% 
  left_join(otu_rel_grass,by = "Taxa") %>% 
  left_join(otu_rel_site,by = "Taxa") %>% 
  left_join(otu_rel_Site_Cover,by = "Taxa")

#head(GENUS_relabund_bySite_Grass1)
colnames(GENUS_relabund_bySite_Grass1)


# ---- Reorder and subset for export ----
GENUS_relabund_bySite_Grass<-cbind("Genus"=GENUS_relabund_bySite_Grass1$Genus,"Family"=GENUS_relabund_bySite_Grass1$Family,"Order"=GENUS_relabund_bySite_Grass1$Order,"Class"=GENUS_relabund_bySite_Grass1$Class,"Phylum"=GENUS_relabund_bySite_Grass1$Phylum,"Domain"=GENUS_relabund_bySite_Grass1$Domain,GENUS_relabund_bySite_Grass1[,9:109])


# Optional taxon filters
GENUS_rel_abund_Monosporascus <- subset(GENUS_relabund_bySite_Grass, Genus == "Monosporascus")
GENUS_rel_abund_Xylariales <- subset(GENUS_relabund_bySite_Grass, Order == "Xylariales")
GENUS_rel_abund_Darksidea <- subset(GENUS_relabund_bySite_Grass, Genus == "Darksidea")

# ---- Export ----
write.csv(GENUS_relabund_bySite_Grass, file = paste0(RAF_PATH_GENUS, "alltaxa.csv"), row.names = FALSE)
write.csv(GENUS_rel_abund_Monosporascus, file = paste0(RAF_PATH_GENUS, "genus_Monosporascus.csv"), row.names = FALSE)
write.csv(GENUS_rel_abund_Xylariales, file = paste0(RAF_PATH_GENUS, "order_Xylariales.csv"), row.names = FALSE)
write.csv(GENUS_rel_abund_Darksidea, file = paste0(RAF_PATH_GENUS, "genus_Darksidea.csv"), row.names = FALSE)
```



Taxonomy assigned using SINTAX algorithm (USEARCH v11)\
Reference database: most current UNITE UNITE USEARCH/UTAX release for eukaryotes

-   version no 9.0, release date 2023-07-18

-   taxon group: All eukaryotes

-   location: <https://dx.doi.org/10.15156/BIO/2938084>

-   citation: Abarenkov, Kessy; Zirk, Allan; Piirmann, Timo; Pöhönen, Raivo; Ivanov, Filipp; Nilsson, R. Henrik; Kõljalg, Urmas (2023): UNITE USEARCH/UTAX release for eukaryotes. Version 18.07.2023. UNITE Community. <https://doi.org/10.15156/BIO/2938084>

All EDGE samples have been included\
ITS OTU table rarefied to `r prettyNum(max(sample_sums(Fun)),big.mark = ",")` seqs/sample

**Fungal (ITS) Dataset:**

-   `r prettyNum(ntaxa(Fun),big.mark = ",")` OTUs
-   `r prettyNum(nsamples(Fun),big.mark = ",")` samples\
-   `r prettyNum(length(get_taxa_unique(Fun, taxonomic.rank = "Genus")),big.mark = ",")` unique genera detected, inc. *Darksidea*\

------------------------------------------------------------------------

### Available metadata {.unnumbered}

Note that numeric metadata has been mean standardized.

```{r mean-std-metadata, eval=TRUE,include=FALSE}
#######
# Fun #
#######
#Fun_m.std <- perform_mean_standardization(Fun)
# coll_GDD,GDD30yr,GDD3yr,GDD2015,ppt30yr,ppt3yr,ppt2015,soil_moisture,GWC,perc_moisture,pH,phos,ammonium,nitrate,SOM,herbivory_perc,avg_SLA,avg_SRL

#saveRDS(Fun_m.std, file="../processed_data/clean_rds_saves/Fun_m.std.rds")

# Fun_m.std <- readRDS(file="../processed_data/clean_rds_saves/Fun_m.std.rds")


# saveRDS(Fun.Genus, file="../processed_data/clean_rds_saves/Fun.Genus.rds")
# Fun.Genus_m.std <- perform_mean_standardization(Fun.Genus)
# saveRDS(Fun.Genus_m.std, file="../processed_data/clean_rds_saves/Fun.Genus_m.std.rds")
# Fun.Genus_m.std <- readRDS(file="../processed_data/clean_rds_saves/Fun.Genus_m.std.rds")

#######
# Bac #
#######
# Bac_m.std <- perform_mean_standardization(Bac)
# coll_GDD,GDD30yr,GDD3yr,GDD2015,ppt30yr,ppt3yr,ppt2015,soil_moisture,GWC,perc_moisture,pH,phos,ammonium,nitrate,SOM,herbivory_perc,avg_SLA,avg_SRL

# saveRDS(Bac_m.std, file="../processed_data/clean_rds_saves/Bac_m.std.rds")

# Bac_m.std <- readRDS(file="../processed_data/clean_rds_saves/Bac_m.std.rds")


```

```{r, echo=F, eval=F}
sd_Fun<-sample_data(Fun)
colnames(sd_Fun)
```

[*Sampling Variables:*]{.underline}

-   **Grass host** (n=6)
    -   ANGE (*Andropogon gerardii* / Big bluestem)
    -   BOCU (*Bouteloua curtipendula* / Sideoats grama)
    -   BOER (*Bouteloua eriopoda* / Black grama)
    -   BOGR (*B. gracilis* / Blue grama)
    -   BUDA (*B. dactyloides* / Buffalograss)
    -   SCSC (*Schizachyrium scoparium* / Little bluestem)
-   **Grassland type** (n=4)
    -   desert, shortgrass, mixed grass, and tallgrass
-   **Site** (n=21)
-   GPS coordinates - **latitude** and **longitude**
-   '**Longitudinal Gradient**' (n=4)
    -   North, North Central, South Central, South
-   '**Latitudinal Bin**' (n=3)
    -   East, Middle, West
-   **Elevation** (in meters)\

[*Climate Data:*]{.underline}

-   **GDD** = growing degree days (determined over 3- and 30-year windows and in 2015)
-   **ppt** = mean annual precipitation (determined over 3- and 30-year windows and in 2015)\
    \*Data from PRISM Climate Group 2019 (Oregon State University)\

[*Edaphic Variables:*]{.underline}

-   **ammonium** (NH~4~^+^)
-   **nitrate** (NO~3~^--^)
-   **SOM** (via LoI)
-   **phosphorous** (total P)
-   **pH**
-   **soil_moisture** = percentage by weight of water in soil (by sample, Rudgers et al. called this a climate variable)\
    \*soils combined within each Grass Host x Site combination\

[*Plant Traits:*]{.underline}

-   **avg_SLA**: specific leaf area
-   **avg_SRL**: specific root length
    -   Large values of specific leaf area (SLA) and specific root length (SRL) reflect large plant investment in resource acquisition; small values indicate investment in plant tissue longevity
-   **herbivory_perc**: average site level damage estimate for herbivory, averaged over all species and individuals at the site; could be used to indicate herbivory pressure at the site level
    -   ranges from 1:17\

------------------------------------------------------------------------

\

## Links to download csv files of OTU occurrence frequencies by Site-by-Grass host combinations

`colnames` and/or `indices` = explanation

-   `OTU` = OTU ID number (fungal)

-   `Domain` : `Species` `([,c(3:9)])` = taxonomic classifications for each OTU

-   `global_occurrence` `[,10]` = proportion of samples with OTU detected over of total number of samples (max of 1)

-   Sevilleta-site only (`r nsamples(Fun_Sev)` samples total, of which `r nrow(subset(sample_data(Fun_Sev), Grass=="BOER"))` were associated with *Black grama* and `r nrow(subset(sample_data(Fun_Sev), Grass=="BOGR"))` with *Blue grama*)

    -   `BOER_SEVonly` `[,11]`

    -   `BOGR_SEVonly` `[,12]`

-   By grass host `([,c(13:18)])` = splitting samples by grass host, proportion of samples with OTU detected over of total number of samples per host. Denominator ranged from 25 - 185 (samples per grass host).

    -   `ANGE_allsites` (n = 118), `BOCU_allsites` (n = 25), `BOER_allsites` (n = 106), `BOGR_allsites` (n = 130), `BUDA_allsites` (n = 90), `SCSC_allsites` (n = 185)

-   By site `([,c(19:42)])` = splitting samples by site, proportion of samples with OTU detected over of total number of samples per site. Denominator ranged from 16 - 91 (samples per site).\

\

[OTU-level frequency of occurrence]{.underline}

All OTUs

-   [RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_OTU_alltaxa.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/master/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_OTU_alltaxa.csv "RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_OTU_alltaxa.csv")

OTUs assigned to genus *Darksidea* (n = 3)

-   [RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_genus_Darksidea.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/master/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_genus_Darksidea.csv "RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_genus_Darksidea.csv")

OTUs assigned to genus *Monosporascus* (n = 2)

-   [RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_genus_Monosporascus.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/master/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_genus_Monosporascus.csv "RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_genus_Monosporascus.csv")

OTUs assigned to order *Xylariales* (order containing *Monosporascus*)(n = 109)

-   [RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_order_Xylariales.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/master/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_order_Xylariales.csv "RAM_allSamplesincEDGE_Freq_OTU_Occurrence_bySitexGrass_order_Xylariales.csv")

\
\
\

## Links to download csv files of Genus-level occurrence frequencies by Site-by-Grass host combinations

Link to GitHub directory: https://github.com/kellifeeser/ksu-paired-amplicon-workflow/tree/master/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus \

Explanation of column names:

-   `Genus` = Fungal genus (all OTUs within each genus were agglomerated)

-   `Family` : `Domain` `([,c(2:6)])` = taxonomic classifications for each genus

-   `global_occurrence` `[,7]` = proportion of samples with genera detected over of total number of samples (max of 1)

-   Sevilleta-site only (`r nsamples(Fun_GENUS_Sev_no0x)` samples total, of which `r nrow(subset(sample_data(Fun_GENUS_Sev_no0x), Grass=="BOER"))` were associated with *Black grama* and `r nrow(subset(sample_data(Fun_GENUS_Sev_no0x), Grass=="BOGR"))` with *Blue grama*)

    -   `BOER_SEVonly` `[,8]`

    -   `BOGR_SEVonly` `[,9]`

-   By grass host `([,c(10:15)])` = splitting samples by grass host, proportion of samples with genera detected over of total number of samples per host. Denominator ranged from 25 - 185 (samples per grass host).

    -   `ANGE_allsites` (n = 118), `BOCU_allsites` (n = 25), `BOER_allsites` (n = 106), `BOGR_allsites` (n = 130), `BUDA_allsites` (n = 90), `SCSC_allsites` (n = 185)

-   By site `([,c(16:39)])` = splitting samples by site, proportion of samples with genera detected over of total number of samples per site. Denominator ranged from 16 - 91 (samples per site).

-   By unique site x grass host combinations `([,c(40:107)])` = splitting samples by each site and grass, proportion of samples with OTU detected over of total number of samples per site. Denominator ranged from `r min(Site_Cover_counts$Sample_Count)` - `r max(Site_Cover_counts$Sample_Count)` (samples per site x grass host, including EDGE samples). *Note*: EDGE samples were included and calculated separately from their corresponding group, and denoted by \*\_E, \*\_E_A, and \*\_E_B. \

\

[Genus-level frequency of occurrence]{.underline}

All Genera


-   [RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_GENUS_alltaxa.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/60e3718728f0ca14573da2edc89e6e9196649d30/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus/RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_GENUS_alltaxa.csv "RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_GENUS_alltaxa")

*Darksidea* (combined from n = 3 OTUs)

-   [RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_genus_Darksidea.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/60e3718728f0ca14573da2edc89e6e9196649d30/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus/RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_genus_Darksidea.csv "RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_genus_Darksidea")

*Monosporascus* (combined from n = 2 OTUs)

-   [RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_genus_Monosporascus.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/60e3718728f0ca14573da2edc89e6e9196649d30/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus/RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_genus_Monosporascus.csv "RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_genus_Monosporascus")

Genera assigned to order *Xylariales* (order containing *Monosporascus*) (combined from n = 109 OTUs)

-   [RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_order_Xylariales.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/60e3718728f0ca14573da2edc89e6e9196649d30/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus/RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_order_Xylariales.csv "RAM_allSamplesincEDGE_Freq_Genus_Occurrence_bySitexGrass_order_Xylariales")

\
\
\

## Links to download csv files of Genus-level Relative Abundance summaries by Site-by-Grass host combinations

Link to GitHub directory: https://github.com/kellifeeser/ksu-paired-amplicon-workflow/tree/master/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus \

Explanation of column names:

-   `Genus` = Fungal genus (all OTUs within each genus were agglomerated)

-   `Family` : `Domain` `([,c(2:6)])` = taxonomic classifications for each genus

-   `global_relative_abundance` `[,7]` = global relative abundance across all samples (sums to 100%)

-   Sevilleta-site only (`r nsamples(Fun_GENUS_Sev_no0x)` samples total, of which `r nrow(subset(sample_data(Fun_GENUS_Sev_no0x), Grass=="BOER"))` were associated with *Black grama* and `r nrow(subset(sample_data(Fun_GENUS_Sev_no0x), Grass=="BOGR"))` with *Blue grama*)

    -   `BOER_SEVonly` `[,8]`

    -   `BOGR_SEVonly` `[,9]`

-   By grass host `([,c(10:15)])` = grouping samples by grass host

    -   `ANGE_allsites` (n = 118), `BOCU_allsites` (n = 25), `BOER_allsites` (n = 106), `BOGR_allsites` (n = 130), `BUDA_allsites` (n = 90), `SCSC_allsites` (n = 185)

-   By site `([,c(16:39)])` = grouping samples by site

-   By unique site x grass host combinations `([,c(40:107)])` = grouping samples by each site and grass, (samples per site x grass host, including EDGE samples). *Note*: EDGE samples were included and calculated separately from their corresponding group, and denoted by \*\_E, \*\_E_A, and \*\_E_B. \

\

[Genus-level Relative Abundance (sums to 100%)]{.underline}

All Genera


-   [RAM_allSamplesincEDGE_Genus_RelativeAbundance_alltaxa.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/962ff3bbc3c4e94f80a2d51912205f62c4ea6def/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus/RAM_allSamplesincEDGE_Genus_RelativeAbundance_alltaxa.csv "RAM_allSamplesincEDGE_Genus_RelativeAbundance_alltaxa.csv")

*Darksidea* (combined from n = 3 OTUs)

-   [RAM_allSamplesincEDGE_Genus_RelativeAbundance_genus_Darksidea.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/962ff3bbc3c4e94f80a2d51912205f62c4ea6def/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus/RAM_allSamplesincEDGE_Genus_RelativeAbundance_genus_Darksidea.csv "RAM_allSamplesincEDGE_Genus_RelativeAbundance_genus_Darksidea.csv")

*Monosporascus* (combined from n = 2 OTUs)

-   [RAM_allSamplesincEDGE_Genus_RelativeAbundance_genus_Monosporascus.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/962ff3bbc3c4e94f80a2d51912205f62c4ea6def/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus/RAM_allSamplesincEDGE_Genus_RelativeAbundance_genus_Monosporascus.csv "RAM_allSamplesincEDGE_Genus_RelativeAbundance_genus_Monosporascus.csv")

Genera assigned to order *Xylariales* (order containing *Monosporascus*) (combined from n = 109 OTUs)

-   [RAM_allSamplesincEDGE_Genus_RelativeAbundance_order_Xylariales.csv](https://github.com/kellifeeser/ksu-paired-amplicon-workflow/blob/962ff3bbc3c4e94f80a2d51912205f62c4ea6def/processed_data/OTU_occurrencefrequency/allSamplesincEDGE/byGenus/RAM_allSamplesincEDGE_Genus_RelativeAbundance_order_Xylariales.csv "RAM_allSamplesincEDGE_Genus_RelativeAbundance_order_Xylariales.csv")

\
\
\

------------------------------------------------------------------------

## OTU-level relative abundance {.unnumbered}

Relative abundance of each OTU were averaged across all samples for these results.

***Darksidea***

This dataset contains `r ntaxa(subset_taxa(Fun, Genus=="Darksidea"))` OTUs belonging to genus *Darksidea*.

At the OTU/species-level, all *Darksidea* OTUs were classified as "`r sort(get_taxa_unique(subset_taxa(Fun, Genus=="Darksidea"), taxonomic.rank = "Species"))`".\

Mean and sd of *Darksidea* OTUs abundance across all samples:

-   `r Darksidea_OTU_stats %>% dplyr::slice(1) %$% OTU`: `r round(Darksidea_OTU_stats %>% dplyr::slice(1) %$% av,3)` ± `r round(Darksidea_OTU_stats %>% dplyr::slice(1) %$% sd,2)`%
-   `r Darksidea_OTU_stats %>% dplyr::slice(2) %$% OTU`: `r prettyNum(Darksidea_OTU_stats %>% dplyr::slice(2) %$% av,digits=2)` ± `r round(Darksidea_OTU_stats %>% dplyr::slice(2) %$% sd,3)`%
-   `r Darksidea_OTU_stats %>% dplyr::slice(3) %$% OTU`: `r prettyNum(Darksidea_OTU_stats %>% dplyr::slice(3) %$% av,digits=2)` ± `r round(Darksidea_OTU_stats %>% dplyr::slice(3) %$% sd,2)`%
    -   Global sum: `r prettyNum(sum(Darksidea_OTU_stats$av),digits=2)`%

***Monosporascus***

This dataset contains `r ntaxa(subset_taxa(Fun, Genus=="Monosporascus"))` OTUs belonging to genus *Monosporascus*.

At the OTU/species-level, all *Monosporascus* OTUs were classified as "`r sort(get_taxa_unique(subset_taxa(Fun, Genus=="Monosporascus"), taxonomic.rank = "Species"))`".

Mean and sd of *Monosporascus* OTUs abundance across all samples:

-   `r Monosporascus_OTU_stats %>% dplyr::slice(1) %$% OTU`: `r prettyNum(Monosporascus_OTU_stats %>% dplyr::slice(1) %$% av,digits=2)` ± `r prettyNum(Monosporascus_OTU_stats %>% dplyr::slice(1) %$% sd,digits=1)`%
-   `r Monosporascus_OTU_stats %>% dplyr::slice(2) %$% OTU`: `r prettyNum(Monosporascus_OTU_stats %>% dplyr::slice(2) %$% av,digits=2)` ± `r round(Monosporascus_OTU_stats %>% dplyr::slice(2) %$% sd,3)`%
    -   Global sum: `r prettyNum(sum(Monosporascus_OTU_stats$av),digits=2)`%

------------------------------------------------------------------------

## Genus-level relative abundance {.unnumbered}

Relative abundance of each OTU were averaged across all samples for these results.\

***Darksidea***

This dataset contains `r ntaxa(subset_taxa(Fun, Genus=="Darksidea"))` OTUs belonging to genus *Darksidea*.

At the species-level, all *Darksidea* OTUs were classified as "`r sort(get_taxa_unique(subset_taxa(Fun, Genus=="Darksidea"), taxonomic.rank = "Species"))`".

Across all samples, *Darksidea* had an average relative abundance of `r round(Darksidea_Genus_stats$av,3)` ± `r round(Darksidea_Genus_stats$sd,2)`%.\

***Monosporascus***

This dataset contains `r ntaxa(subset_taxa(Fun, Genus=="Monosporascus"))` OTUs belonging to genus *Monosporascus*.

At the species-level, all *Monosporascus* OTUs were classified as "`r sort(get_taxa_unique(subset_taxa(Fun, Genus=="Monosporascus"), taxonomic.rank = "Species"))`".

Across all samples, *Monosporascus* had an average relative abundance of `r prettyNum(Monosporascus_Genus_stats$av,digits=2)` ± `r round(Monosporascus_Genus_stats$sd,3)`% (sum of all relative abundances is 100%).\
\

```{r, echo=F, eval=T, warning=F}
datatable(Fun_Genus_Mean_Abund2, caption = "SEARCHABLE table of genus-level taxonomy relative abundance & abundance rank and percentile, across all non-EDGE samples.    Notes: 'Abundance' is expressed as mean relative abundance percentage of genera across all non-EDGE samples. 'Rank_RA' indicates the rank of each genus by mean global relative abundance (rank of 1 is the most abundant, 416 is the least [gapless variant of ties.method = 'min']). 'Percentile' indicates the proportion of taxa with a higher mean relative abundance (0 is the most abundant).",
          rownames = FALSE,
          filter = 'top',
          #width=680,
          #style = "font-size: 8px",
          options = list(searching = T,
                         order = list(list(7, 'desc')), #sort by decreasing abundance
                         initComplete = JS(
              "function(settings, json) {",
              "  var table = settings.oInstance.api();",
              "  var darksideaIndex = table.column(6).data().indexOf('Darksidea');",
              "  var darksideaRow = table.row(darksideaIndex).node();",
              "  var fusariumIndex = table.column(6).data().indexOf('Fusarium');",
              "  var fusariumRow = table.row(fusariumIndex).node();",
              "  var monosporascusIndex = table.column(6).data().indexOf('Monosporascus');",
              "  var monosporascusRow = table.row(monosporascusIndex).node();",
              "  var tbody = table.table().body();",
              "  $(darksideaRow).prependTo(tbody);",
              "  $(darksideaRow).css('background-color', '#94D747');",
              "  $(fusariumRow).insertAfter(darksideaRow);",
              "  $(fusariumRow).css('background-color', '#EF9F43');",
              "  $(monosporascusRow).insertAfter(fusariumRow);", # Placing Monosporascus after Fusarium
              "  $(monosporascusRow).css('background-color', '#9CD2E8');", # 
              "}"
            ),
            scrollX=TRUE,
                         columnDefs = list(list(visible = FALSE, targets = c('Fun_OTU', 'Domain', 'Order')), #domain
                                           list(width = '16%', targets = c('Class', 'Family', 'Genus')), #family
                                           list(width = '25%', targets = 'Mean rel abund (%)'), #mean RA
                                           list(width = '25%', targets = colnames(Fun_Genus_Mean_Abund2)[9]) 
                                           #list(width = '125%', targets = 8) #rank RA
                                           #,list(width = '110%', targets = 9) #percentile
                                           )
                         )
          ) %>% formatRound('Percentile', 1)

```

\
\

------------------------------------------------------------------------

# Distribution of *Darksidea* OTUs {.unnumbered}

n=3 *Darksidea* OTUs

[At the SEV only:]{.underline}

-   Of the 2 grasses present at the SEV (*Blue* and *Black grama*),

    -   1 / 3 OTUs not detected at the SEV (`OTU7431`)

    -   1 / 3 OTUs detected in 53.6% *Blue grama* associated soils (`OTU5613`)

    -   2 / 3 OTUs detected in *Black grama* associated soils (`OTU9749` in 2.7%, `OTU5613` in 42.3%)

[Across all sites]{.underline}

-   Mean occurrence frequency of the 3 OTUs across all sites and samples: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]$global_occurrence)*100,digits=2)`%

-   At least 2 OTUs detected in all 6 grasses

    -   ANGE (*Andropogon gerardii* / Big bluestem) associated soils

        -   Lowest occurrence frequencies

        -   2 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$ANGE_allsites >0,]$ANGE_allsites, na.rm = TRUE)*100,digits=2)`%, when present\

    -   BOCU (*Bouteloua curtipendula* / Sideoats grama) associated soils

        -   2 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BOCU_allsites >0,]$BOCU_allsites, na.rm = TRUE)*100,digits=2)`%, when present\
            \

    -   BOER (*Bouteloua eriopoda* / Black grama) associated soils

        -   3 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BOER_allsites >0,]$BOER_allsites, na.rm = TRUE)*100,digits=2)`%\
            \

    -   BOGR (*B. gracilis* / Blue grama) associated soils

        -   3 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BOGR_allsites >0,]$BOGR_allsites, na.rm = TRUE)*100,digits=3)`%\
            \

    -   BUDA (*B. dactyloides* / Buffalograss) associated soils

        -   2 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BUDA_allsites >0,]$BUDA_allsites, na.rm = TRUE)*100,digits=3)`%, when present\
            \

    -   SCSC (*Schizachyrium scoparium* / Little bluestem) associated soils

        -   3 / 3 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$SCSC_allsites >0,]$SCSC_allsites, na.rm = TRUE)*100,digits=2)`%\
            \

[By Site]{.underline} ONF RNF SCP SEV UHC SFA - 24 sites in dataset

-   2 sites had all 3 *Darksidea* OTUs detected, with the following mean frequency

    -   BLM: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BLM >0,]$BLM, na.rm = TRUE)*100,digits=3)`%\

    -   GMT: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$GMT >0,]$GMT, na.rm = TRUE)*100,digits=3)`%\

    -   `OTU7431`: only detected in the above 2 sites (with other 2 OTUs)\

-   8 sites had 2 of 3 *Darksidea* OTUs detected, with the following mean frequencies (when present)

    -   OTUs `OTU9749` and `OTU5613`
        -   CPR: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$CPR >0,]$CPR, na.rm = TRUE)*100,digits=3)`%
        -   FCP: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$FCP >0,]$FCP, na.rm = TRUE)*100,digits=3)`%
        -   FMT: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$FMT >0,]$FMT, na.rm = TRUE)*100,digits=2)`%
        -   HAR: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$HAR >0,]$HAR, na.rm = TRUE)*100,digits=3)`%
        -   HPG: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$HPG >0,]$HPG, na.rm = TRUE)*100,digits=3)`%
        -   KNZ: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$KNZ >0,]$KNZ, na.rm = TRUE)*100,digits=3)`%
        -   RNF: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$RNF >0,]$RNF, na.rm = TRUE)*100,digits=3)`%
        -   SEV: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$SEV >0,]$SEV, na.rm = TRUE)*100,digits=3)`%

-   6 sites had 1 of 3 *Darksidea* OTUs detected, with the following mean frequencies (when present)

    -   OTU `OTU9749`
        -   CAD: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$CAD >0,]$CAD, na.rm = TRUE)*100,digits=3)`%
        -   ONF: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$ONF >0,]$ONF, na.rm = TRUE)*100,digits=3)`%
        -   SCP: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$SCP >0,]$SCP, na.rm = TRUE)*100,digits=4)`%
    -   OTU `OTU5613`
        -   CNF: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$CNF >0,]$CNF, na.rm = TRUE)*100,digits=3)`%
        -   DMT: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$DMT >0,]$DMT, na.rm = TRUE)*100,digits=3)`%
        -   LBJ: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$LBJ >0,]$LBJ, na.rm = TRUE)*100,digits=2)`%

-   8 sites had no *Darksidea* OTUs detected (BNP, GNF, KAE, LAR, NWP, ONF, SFA, UHC)\
    \

------------------------------------------------------------------------

# Distribution of *Monosporascus* OTUs {.unnumbered}

n=2 *Monosporascus* OTUs

[Not detected at the SEV]{.underline}

-   Of the 2 grasses present at the SEV (*Blue* and *Black grama*),

    -   0 OTUs detected in *Blue grama* associated soils

    -   1 / 2 OTUs not detected at the SEV (`OTU4769`)

    -   1 / 2 OTUs detected in 2.7% *Black grama* associated SEV soils (`OTU8515`)

[Across all sites]{.underline}

-   Mean occurrence frequency of the 2 OTUs across all sites and samples: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]$global_occurrence)*100,digits=2)`%

-   No OTUs detected in 3 grasses

    -   ANGE (*Andropogon gerardii* / Big bluestem) associated soils
    -   BOCU (*Bouteloua curtipendula* / Sideoats grama) associated soils
    -   SCSC (*Schizachyrium scoparium* / Little bluestem) associated soils\
        \

-   At least 1 OTU detected in 3 grasses

    -   BOER (*B. eriopoda* / Black grama) associated soils

        -   1 / 2 OTUs (`OTU8515`) detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$BOER_allsites >0,]$BOER_allsites, na.rm = TRUE)*100,digits=2)`%, when present\
            \

    -   BOGR (*B. gracilis* / Blue grama) associated soils

        -   2 / 2 OTUs detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$BOGR_allsites >0,]$BOGR_allsites, na.rm = TRUE)*100,digits=2)`%, when present\
            \

    -   BUDA (*B. dactyloides* / Buffalograss) associated soils

        -   1 / 2 OTUs (`OTU4769`) detected with mean occurrence frequency of `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$BUDA_allsites >0,]$BUDA_allsites, na.rm = TRUE)*100,digits=3)`%, when present\
            \

[By Site]{.underline}

-   24 sites in dataset

-   0 sites had both of the *Monosporascus* OTUs detected

-   4 sites had 1 of 2 *Monosporascus* OTUs detected, with the following mean frequencies (when present)

    -   OTU `OTU8515`
        -   CNF: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$CNF >0,]$CNF, na.rm = TRUE)*100,digits=3)`%
        -   SEV: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$SEV >0,]$SEV, na.rm = TRUE)*100,digits=3)`%
    -   OTU `OTU4769`
        -   FCP: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$FCP >0,]$FCP, na.rm = TRUE)*100,digits=3)`%
        -   FMT: `r prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus" & OTU_freq_occurr_bySite_Grass$FMT >0,]$FMT, na.rm = TRUE)*100,digits=2)`%

-   18 sites had no *Monosporascus* OTUs detected (BLM, BNP, CAD, CPR, DMT, GMT, GNF, HAR, HPG, KAE, KNZ, LAR, LBJ, NWP, ONF, RNF, SCP, SFA, UHC)\
    \

```{r distribution-stats, include=F, eval=F}
View(t(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]))

# At the SEV only
OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]$BOGR_SEVonly
OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]$BOER_SEVonly

OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]$BOGR_SEVonly
OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]$BOER_SEVonly



# OTU occurrence by site and grass
OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]

OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Monosporascus",]


prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",]$global_occurrence)*100,digits=2)

prettyNum(mean(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea" & OTU_freq_occurr_bySite_Grass$BOER_allsites >0,]$BOER_allsites, na.rm = TRUE)*100,digits=2)

# idk
View(OTU_freq_occurr_bySite_Grass[OTU_freq_occurr_bySite_Grass$Genus=="Darksidea",c(1,20:39)])

mean(OTU_freq_occurr_bySite_Grass$ANGE_allsites[Genus=="Darksidea"])

```

------------------------------------------------------------------------
