---
title: "Fungal F1 vs. non-F1"
author: "Kelli Feeser"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
    fig.caption: yes
  bookdown::html_document2:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
    fig.caption: yes
    keep_md: yes
  html_notebook:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
    fig.caption: yes
editor_options:
  chunk_output_type: inline
---

\

# Set-up {.unlisted .unnumbered .hidden}

```{r global-options, include=FALSE,eval=T}

FIG_PATH = "../docs/github_F1vnonF1/"
CACHE_PATH = "../docs/github_F1vnonF1_cache/"

## ~~create directories~~ ##
###########################################
dir.create(FIG_PATH, showWarnings = T,recursive = T)
dir.create(CACHE_PATH, showWarnings = T,recursive = T)
###########################################

knitr::opts_chunk$set(fig.path=FIG_PATH,
                      cache.path = CACHE_PATH)
opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(fig.pos = '!h')
options(knitr.kable.NA = '')
```

## load packages

```{r load-packages, eval=T, message=F,results='hide',warning=F, include=FALSE}
library("vegan"); packageVersion("vegan") # ‘2.6.10’
library("phyloseq"); packageVersion("phyloseq") # ‘1.50.0’
library("betapart"); packageVersion("betapart") # ‘1.6’
library("ecodist"); packageVersion("ecodist") # ‘2.1.3’
library("measurements"); packageVersion("measurements") # ‘1.5.1’
library("geosphere"); packageVersion("geosphere") # ‘1.5.20’
  # The legacy packages maptools, rgdal, and rgeos, underpinning the sp package,
  # which was just loaded, were retired in October 2023.
  # Please refer to R-spatial evolution reports for details, especially
  # https://r-spatial.org/r/2023/05/15/evolution4.html.
  # It may be desirable to make the sf package available;
  # package maintainers should consider adding sf to Suggests:.
library("gdata"); packageVersion("gdata") # ‘3.0.1’
# library("spam"); packageVersion("spam") # ‘2.9.1’
library("ggplot2"); packageVersion("ggplot2") # ‘3.5.1’
library("ggpubr"); packageVersion("ggpubr") # ‘0.6.0’
library("cowplot"); packageVersion("cowplot") # ‘1.1.3’
library("geodist"); packageVersion("geodist") # ‘0.1.0’
library("knitr"); packageVersion("knitr") # ‘1.49’
library("kableExtra"); packageVersion("kableExtra") # ‘1.4.0’

source("RAM_bulk_harvester.R")
```

## load cache & rds

```{r load-cache, include=FALSE, eval=T}
# Load RDS
  # OBJECT <- if (file.exists('OBJECT.rds')) {
  #   readRDS('OBJECT.rds')
  # } else {
  #   "The value of `OBJECT` is not available yet"
  # }

env_dist_matrices <- readRDS(file = "../processed_data/clean_rds_saves/env_dist_matrices.rds")
betapart_results <- readRDS(file = "../processed_data/clean_rds_saves/betapart_results.rds")

# Load chunk cache / chunk cache object
  # knitr::load_cache('CHUNK-NAME', 'OBJECT')
  # knitr::load_cache('calc-DD-BF-all')

knitr::load_cache('calc-Fun-clus2-sor-sim-sne-geo')
```


# Pure spatial effects (S)

## Spatial Distance Decay

### Fungal F1 vs. non-F1

Spatial (geographic) distance decay - Fitting distance decay models to pairwise community dissimilarity\

- Spatial distance in kilometers\
    -   We assessed the increase of community dissimilarity within each fungal subassemblage group (F1 vs. non-F1) with geographic distance (equivalent to the distance decay of similarity), using `as.matrix(geodist(fungal_F1.geo, measure = 'geodesic' )/1000)`. We then generated a distance matrix containing the geographic distance, based on longitude and latitude (in kilometers) among the pairwise comparisons for each group.\ \ 

The subassemblage groups span significantly different distributions of geographic distances (Wilcoxon rank sum test; )

```{r wilcox-Fun-clus2-geodist, echo=FALSE}
library(dplyr)
library(e1071)
library(car)
# If not installed: install.packages("nortest")
library(nortest)

dist1 <- env_dist_matrices[["Fun_sor_clus2"]][["F1_geodist"]]
dist2 <- env_dist_matrices[["Fun_sor_clus2"]][["non-F1_geodist"]]

# Convert 'dist' to numeric vectors
vec1 <- as.numeric(dist1)  # F1
vec2 <- as.numeric(dist2)  # non-F1

# Create tidy dataframe
dist_df <- data.frame(
  Distance = c(vec1, vec2),
  Group = factor(c(rep("F1", length(vec1)), rep("non-F1", length(vec2))),
                 levels = c("F1", "non-F1"))
)

# Summary stats
summary_stats <- dist_df %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    Mean = mean(Distance),
    Median = median(Distance),
    Variance = var(Distance),
    SD = sd(Distance),
    Min = min(Distance),
    Max = max(Distance),
    Skewness = skewness(Distance),
    Kurtosis = kurtosis(Distance),
    .groups = "drop"
  )

# Normality
# Anderson-Darling test for large sample normality
ad_f1 <- nortest::ad.test(vec1)
ad_nonf1 <- nortest::ad.test(vec2)

# Optionally also run Kolmogorov–Smirnov against normal
ks_f1 <- ks.test(vec1, "pnorm", mean(vec1), sd(vec1))
ks_nonf1 <- ks.test(vec2, "pnorm", mean(vec2), sd(vec2))

# Updated logic for mean test based on AD or KS
if (ad_f1$p.value > 0.05 & ad_nonf1$p.value > 0.05 &
    abs(var(vec1) - var(vec2)) < 0.5 * max(var(vec1), var(vec2))) {
  mean_test <- t.test(Distance ~ Group, data = dist_df)
  test_label <- "t-test"
} else {
  mean_test <- wilcox.test(Distance ~ Group, data = dist_df)
  test_label <- "Wilcoxon rank-sum test"
}
# Variance test
var_test <- leveneTest(Distance ~ Group, data = dist_df)

# Custom colors
col_f1 <- "#8E0152"
col_f1_alpha <- "#8E015266"
col_nonf1 <- "#7FBC41"
col_nonf1_alpha <- "#276419"

fill_colors <- c("F1" = col_f1, "non-F1" = col_nonf1)

# Violin + Boxplot + Jitter
p_violin <- ggplot(dist_df, aes(x = Group, y = Distance, fill = Group)) +
  geom_violin(alpha = 0.3, trim = FALSE) +
  geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.9) +
  geom_jitter(width = 0.1, size = 0.7, alpha = 0.1) +
  scale_fill_manual(values = fill_colors) +
  labs(
    title = "Comparison of Spatial Distances (within groups)",
    subtitle = paste0(
      "Boxplot of spatial distances between F1 and non-F1 samples. ",
      "F1 spans a maximum distance of ", round(max(vec1)), " km ",
      "(mean ± SD: ", round(mean(vec1), 1), " ± ", round(sd(vec1), 1), "). ",
      "non-F1 spans a maximum distance of ", round(max(vec2)), " km ",
      "(mean ± SD: ", round(mean(vec2), 1), " ± ", round(sd(vec2), 1), "). ",
      "The groups differed in mean (", test_label, 
      "; W = ", round(mean_test$statistic, 2), 
      ", p = ", signif(mean_test$p.value, 3), "). ",
      "Variance comparison (Levene’s test: F = ", round(var_test$`F value`[1], 2), 
      ", p = ", signif(var_test$`Pr(>F)`[1], 3), ")."
    ),
    x = NULL, y = "Distance (km)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

# Q-Q plots
qq_f1 <- ggplot(data.frame(Distance = vec1), aes(sample = Distance)) +
  stat_qq(color = col_f1) +
  stat_qq_line(color = col_f1_alpha) +
  labs(title = "Q-Q Plot: F1") +
  theme_minimal()

qq_nonf1 <- ggplot(data.frame(Distance = vec2), aes(sample = Distance)) +
  stat_qq(color = col_nonf1) +
  stat_qq_line(color = col_nonf1_alpha) +
  labs(title = "Q-Q Plot: non-F1") +
  theme_minimal()

# Combine all plots
final_plot <- plot_grid(
  plot_grid(qq_f1, qq_nonf1, labels = c("A", "B"), ncol = 2),
  p_violin + labs(caption = "Figure 1: Q-Q plots (A, B) and distribution (C) of spatial distances within F1 and non-F1 groups."),
  labels = c("", "C"),
  ncol = 1,
  rel_heights = c(0.4, 0.6)
)

# Show plot
print(final_plot)

# Show summary stats table
kable(summary_stats, digits = 2, caption = "Summary statistics for within-group spatial distances") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

```



```{r fig-spatial-dist, fig.cap = "Figure 1. Q–Q plots and distribution of spatial distances within F1 and non-F1 groups.", fig.width=10, fig.height=6, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(e1071)
library(car)
library(nortest)
library(cowplot)
library(knitr)
library(kableExtra)
library(glue)

dist1 <- env_dist_matrices[["Fun_sor_clus2"]][["F1_geodist"]]
dist2 <- env_dist_matrices[["Fun_sor_clus2"]][["non-F1_geodist"]]

# Data prep
vec1 <- as.numeric(dist1)
vec2 <- as.numeric(dist2)

dist_df <- data.frame(
  Distance = c(vec1, vec2),
  Group = factor(c(rep("F1", length(vec1)), rep("non-F1", length(vec2))),
                 levels = c("F1", "non-F1"))
)

# Summary stats
summary_stats <- dist_df %>%
  group_by(Group) %>%
  summarise(
    N = n(),
    Mean = mean(Distance),
    SD = sd(Distance),
    Median = median(Distance),
    Min = min(Distance),
    Max = max(Distance),
    Variance = var(Distance),
    Skewness = skewness(Distance),
    Kurtosis = kurtosis(Distance),
    .groups = "drop"
  )

# Normality tests
ad_f1 <- ad.test(vec1)
ad_nonf1 <- ad.test(vec2)

# Variance test
var_test <- leveneTest(Distance ~ Group, data = dist_df)

# Mean/median test
if (ad_f1$p.value > 0.05 & ad_nonf1$p.value > 0.05 &
    abs(var(vec1) - var(vec2)) < 0.5 * max(var(vec1), var(vec2))) {
  mean_test <- t.test(Distance ~ Group, data = dist_df)
  test_label <- "t-test"
} else {
  mean_test <- wilcox.test(Distance ~ Group, data = dist_df)
  test_label <- "Wilcoxon rank-sum test"
}

# Add flags to summary table
summary_stats <- summary_stats %>%
  mutate(
    Normality = c(ifelse(ad_f1$p.value < 0.05, "No", "Yes"),
                  ifelse(ad_nonf1$p.value < 0.05, "No", "Yes"))
  )

var_p <- var_test$`Pr(>F)`[1]
mean_p <- mean_test$p.value

summary_stats <- summary_stats %>%
  mutate(
    Var_Differs = ifelse(var_p < 0.05, "Yes", "No"),
    Mean_Differs = ifelse(mean_p < 0.05, "Yes", "No")
  )

# Custom colors
col_f1 <- "#8E0152"; col_f1_line <- "#8E015266"
col_nonf1 <- "#7FBC41"; col_nonf1_line <- "#276419"
fill_colors <- c("F1" = col_f1, "non-F1" = col_nonf1)

# Minimal black theme with min font size 14
black_theme <- theme_minimal(base_size = 14) +
  theme(
    text = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    plot.title = element_text(color = "black", face = "bold"),
    strip.text = element_text(color = "black")
  )

# Q–Q plots
qq_f1 <- ggplot(data.frame(Distance = vec1), aes(sample = Distance)) +
  stat_qq(color = col_f1) + stat_qq_line(color = col_f1_line) +
  labs(title = "Q–Q Plot: F1") +
  black_theme

qq_nonf1 <- ggplot(data.frame(Distance = vec2), aes(sample = Distance)) +
  stat_qq(color = col_nonf1) + stat_qq_line(color = col_nonf1_line) +
  labs(title = "Q–Q Plot: non-F1") +
  black_theme

# Violin + boxplot
p_dist <- ggplot(dist_df, aes(x = Group, y = Distance, fill = Group)) +
  geom_violin(trim = FALSE, alpha = 0.3) +
  geom_boxplot(width = 0.15, outlier.shape = NA, alpha = 0.9) +
  geom_jitter(width = 0.1, alpha = 0.1, size = 0.4) +
  scale_fill_manual(values = fill_colors) +
  labs(title = "C. Distance Distributions", y = "Distance (km)", x = NULL) +
  black_theme +
  theme(legend.position = "none")

# Combine plots
plot_grid(
  plot_grid(qq_f1, qq_nonf1, labels = c("A", "B"), ncol = 2),
  p_dist,
  ncol = 1, rel_heights = c(0.4, 0.6)
)
```


```{r fig1-summary-table-ext-stats, results='asis'}
# Extended stats
f1_ext <- data.frame(
  Group = "F1",
  IQR = IQR(vec1),
  Range = diff(range(vec1)),
  CV = sd(vec1) / mean(vec1)
)

nf1_ext <- data.frame(
  Group = "non-F1",
  IQR = IQR(vec2),
  Range = diff(range(vec2)),
  CV = sd(vec2) / mean(vec2)
)

# Merge with base stats
group_stats <- summary_stats %>%
  left_join(bind_rows(f1_ext, nf1_ext), by = "Group")

# Create comparison row with test results embedded in columns
comparison_row <- data.frame(
  Group = "Comparison",
  N = NA,
  Mean = glue::glue("{ifelse(mean_p < 0.05, 'different', 'not different')} (p = {signif(mean_p, 3)})"),
  SD = NA,
  Median = NA,
  Min = NA,
  Max = NA,
  Variance = glue::glue("{ifelse(var_p < 0.05, 'different', 'not different')} (p = {signif(var_p, 3)})"),
  Skewness = NA,
  Kurtosis = NA,
  Normality = NA,
  IQR = NA,
  Range = NA,
  CV = NA,
  Cohens_d = round(cohen_d, 2),
  Log_Response_Ratio = round(log_ratio, 2)
)

# Final table
final_stats <- bind_rows(group_stats, comparison_row)

# Render it
kable(final_stats, digits = 2, caption = "Table 1. Summary statistics for within-group spatial distances (km), with comparison-level test results embedded in 'Comparison' row.") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), font_size = 10, full_width = FALSE)
```



```{r fig1-summary-text, results='asis'}
f1 <- final_stats %>% filter(Group == "F1")
nf1 <- final_stats %>% filter(Group == "non-F1")

# Compare values
greater_var <- ifelse(f1$Variance > nf1$Variance, "F1", "non-F1")
greater_mean <- ifelse(f1$Mean > nf1$Mean, "F1", "non-F1")
greater_skew <- ifelse(f1$Skewness > nf1$Skewness, "F1", "non-F1")
greater_kurt <- ifelse(f1$Kurtosis > nf1$Kurtosis, "F1", "non-F1")
greater_cv <- ifelse(f1$CV > nf1$CV, "F1", "non-F1")
greater_iqr <- ifelse(f1$IQR > nf1$IQR, "F1", "non-F1")

cat(glue::glue("
**Figure 1 Summary:**  

F1 distances were {ifelse(f1$Normality == 'No', 'not ', '')}normally distributed (Anderson–Darling *p* = {signif(ad_f1$p.value, 3)}).\  

Mean ± SD = {round(f1$Mean, 1)} ± {round(f1$SD, 1)} km,  
Skewness = {round(f1$Skewness, 2)}, Kurtosis = {round(f1$Kurtosis, 2)}.  

non-F1 distances were {ifelse(nf1$Normality == 'No', 'not ', '')}normally distributed (p = {signif(ad_nonf1$p.value, 3)}).  

Mean ± SD = {round(nf1$Mean, 1)} ± {round(nf1$SD, 1)} km,  
Skewness = {round(nf1$Skewness, 2)}, Kurtosis = {round(nf1$Kurtosis, 2)}.  

Group variances differed {ifelse(var_p < 0.05, '**significantly**', 'non-significantly')}  
(Levene’s *F* = {round(var_test$`F value`[1], 2)}, *p* = {signif(var_p, 3)}),  
with **{greater_var}** showing greater variance.  

Central tendency differed {ifelse(mean_p < 0.05, '**significantly**', 'non-significantly')}  
({test_label}, *p* = {signif(mean_p, 3)}),  
with **{greater_mean}** having the greater mean distance.  

F1 had a {ifelse(f1$CV > nf1$CV, 'higher', 'lower')} coefficient of variation (CV = {round(f1$CV, 2)}) than non-F1 (CV = {round(nf1$CV, 2)}).  

The groups differed in interquartile range, with **{greater_iqr}** having greater spread (IQR).  

Cohen's *d* = {round(cohen_d, 2)}; log-response ratio = {round(log_ratio, 2)}.
  
---

**Ecological Interpretation:**  
F1 appears to be more spatially clustered than non-F1.  
This is supported by its lower variance, smaller interquartile range, and lower coefficient of variation,  
indicating less spread in pairwise spatial distances.  
The distribution also exhibits lower skewness and kurtosis, suggesting fewer extreme distances and a more central tendency.  
In contrast, non-F1 distances are more dispersed, possibly reflecting broader habitat use or sampling across a wider spatial domain.
"))
```

```{r fig1-summary-table, results='asis'}
kable(summary_stats, digits = 2, caption = "Table 1. Summary statistics for spatial distances within F1 and non-F1 groups") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```


    -   The F1 subassemblage spans a maximum distance of `r round(max(env_dist_matrices[["Fun_sor_clus2"]][["F1_geodist"]]))` km between samples\

    -   The non-F1 subassemblage group spans a maximum distance of `r round(max(env_dist_matrices[["Fun_sor_clus2"]][["non-F1_geodist"]]))` km\

Distance-decay relationships between geographic distance (in km) among fungal communities in F1 vs. non-F1 subassemblages and i) total community pairwise dissimilarity (β~sor~), ii) spatial turnover-resultant dissimilarity (β~sim~), and iii) nestedness-resultant dissimilarity (β~sne~) using 999 permutations. Differences between F1 and non-F1 model parameters were assessed via block-site bootstrap whether the parameters of both models are equal (n = 999 bootstrap resamples). Both tests were conducted independently, one for the equality of the first parameters (intercepts) of both models, and for the equality of the second parameters (slopes) of both models of each dissimilarity measure.\

```{r calc-Fun-clus2-sor-sim-sne-geo, warning=T,eval=TRUE,cache=T}
#sor
# F1
decay.Fun_sor_clus2F1.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["F1_Fun"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clus2"]][["F1_geodist"]], y.type="dissim", model.type="exp", perm = 999)

# non-F1
decay.Fun_sor_clus2nonF1.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["non-F1_Fun"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clus2"]][["non-F1_geodist"]], y.type="dissim", model.type="exp", perm = 999)

#sor
# F1
decay.Fun_sor_clus2F1.exp.sim<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["F1_Fun"]]$beta.sim, x=env_dist_matrices[["Fun_sor_clus2"]][["F1_geodist"]], y.type="dissim", model.type="exp", perm = 999)

# non-F1
decay.Fun_sor_clus2nonF1.exp.sim<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["non-F1_Fun"]]$beta.sim, x=env_dist_matrices[["Fun_sor_clus2"]][["non-F1_geodist"]], y.type="dissim", model.type="exp", perm = 999)

#sne
# F1
decay.Fun_sor_clus2F1.exp.sne<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["F1_Fun"]]$beta.sne, x=env_dist_matrices[["Fun_sor_clus2"]][["F1_geodist"]], y.type="dissim", model.type="exp", perm = 999)

# non-F1
decay.Fun_sor_clus2nonF1.exp.sne<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["non-F1_Fun"]]$beta.sne, x=env_dist_matrices[["Fun_sor_clus2"]][["non-F1_geodist"]], y.type="dissim", model.type="exp", perm = 999)



# Site-block bootstrap

# Assessing the differences between parameters of two distance-decay models computed with decay.model().
zdepDDBF_F1nonF1_sor<-zdep(decay.Fun_sor_clus2F1.exp.sor, decay.Fun_sor_clus2nonF1.exp.sor, resamples=999, st.val = c(1, 0))

zdepDDBF_F1nonF1_sim<-zdep(decay.Fun_sor_clus2F1.exp.sim, decay.Fun_sor_clus2nonF1.exp.sim, resamples=999, st.val = c(1, 0))

zdepDDBF_F1nonF1_sne<-zdep(decay.Fun_sor_clus2F1.exp.sne, decay.Fun_sor_clus2nonF1.exp.sne, resamples=999, st.val = c(1, 0))
```


```{r plot-Fun-clus-sor-sim-geo, warning=T,eval=TRUE,cache=F, fig.dim=c(8,4),fig.dpi=250}
par(mfrow = c(1, 3),#c(nr, nc) 
    mar = c(2, 2, 2.5, 2),# c(bottom, left, top, right) defaults mar  c(5.1, 4.1, 4.1, 2.1)
    oma = (c(3, 0, 2, 0) + 0.1))

# sor
############################################################
plot.decay.pretty(decay.Fun_sor_clus2F1.exp.sor, col=paste(col_n10[1],"66",sep = ""), add=F, main="Total Dissimilarity (βsor)\nFungi",cex.main=0.9,
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[1],xaxp = c(0,1500,3),
           ylab="",xlab="")
mtext("\nDistance between Communities (in km)", side = 1, line = 1, outer = TRUE) # side 1=bottom, 2=left, 3=top, 4=right

plot.decay.pretty(decay.Fun_sor_clus2nonF1.exp.sor, col="#7FBC41", add=T, main="Total Dissimilarity (βsor)\nFungi: non-F1", cex.main=0.9,
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line="#276419",xaxp = c(0,1500,3),
           ylab="",xlab="")
mtext((paste("F1 - ",round(decay.Fun_sor_clus2F1.exp.sor$pseudo.r.squared,3)," (",round(decay.Fun_sor_clus2F1.exp.sor$p.value,3),")",sep = "")),side=1,line=-1,adj=0.05,outer = F,padj = 0,col=col_n10[1])
mtext((paste("non-F1 - ",round(decay.Fun_sor_clus2nonF1.exp.sor$pseudo.r.squared,3)," (",round(decay.Fun_sor_clus2nonF1.exp.sor$p.value,3),")",sep = "")),side=1,line=-2.2,adj=0.05,outer = F,padj = 0, col="#7FBC41")
mtext((paste("pseudo R2 (p-value)",sep = "")),side=1,line=-3.4,adj=0.05,outer = F,padj = 0, col="black")
############################################################



# sim
############################################################
plot.decay.pretty(decay.Fun_sor_clus2F1.exp.sim, col=paste(col_n10[1],"66",sep = ""), add=F, main="Turnover-Resultant Dissimilarity (βsim)\nFungi",cex.main=0.9,
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[1],xaxp = c(0,1500,3),
           ylab="",xlab="")
mtext("\nDistance between Communities (in km)", side = 1, line = 1, outer = TRUE) # side 1=bottom, 2=left, 3=top, 4=right

plot.decay.pretty(decay.Fun_sor_clus2nonF1.exp.sim, col="#7FBC41", add=T, main="Turnover-Resultant Dissimilarity (βsim)\nFungi: non-F1", cex.main=0.9,
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line="#276419",xaxp = c(0,1500,3),
           ylab="",xlab="")
mtext((paste("F1 - ",round(decay.Fun_sor_clus2F1.exp.sim$pseudo.r.squared,3)," (",round(decay.Fun_sor_clus2F1.exp.sim$p.value,3),")",sep = "")),side=1,line=-1,adj=0.05,outer = F,padj = 0,col=col_n10[1])
mtext((paste("non-F1 - ",round(decay.Fun_sor_clus2nonF1.exp.sim$pseudo.r.squared,3)," (",round(decay.Fun_sor_clus2nonF1.exp.sim$p.value,3),")",sep = "")),side=1,line=-2.2,adj=0.05,outer = F,padj = 0, col="#7FBC41")
mtext((paste("pseudo R2 (p-value)",sep = "")),side=1,line=-3.4,adj=0.05,outer = F,padj = 0, col="black")
############################################################


# sne
############################################################
plot.decay.pretty(decay.Fun_sor_clus2F1.exp.sne, col=paste(col_n10[1],"66",sep = ""), add=F, main="Nestedness-Resultant Dissimilarity (βsne)\nFungi",cex.main=0.9,
           xlim = c(0,1600),ylim = c(0,0.5),las=1, lty=1,lwd=5,
           col.line=col_n10[1],xaxp = c(0,1500,3),
           ylab="",xlab="")
mtext("\nDistance between Communities (in km)", side = 1, line = 1, outer = TRUE) # side 1=bottom, 2=left, 3=top, 4=right

plot.decay.pretty(decay.Fun_sor_clus2nonF1.exp.sne, col="#7FBC41", add=T, main="Nestedness-Resultant (βsne)\nFungi: non-F1", cex.main=0.9,
           xlim = c(0,1600),ylim = c(0,0.5),las=1, lty=1,lwd=5,
           col.line="#276419",xaxp = c(0,1500,3),
           ylab="",xlab="")
mtext((paste("F1 - ",round(decay.Fun_sor_clus2F1.exp.sne$pseudo.r.squared,3)," (",round(decay.Fun_sor_clus2F1.exp.sne$p.value,3),")",sep = "")),side=3,line=-3.4,adj=0.05,outer = F,padj = 0,col=col_n10[1])
mtext((paste("non-F1 - ",round(decay.Fun_sor_clus2nonF1.exp.sne$pseudo.r.squared,3)," (",round(decay.Fun_sor_clus2nonF1.exp.sne$p.value,3),")",sep = "")),side=3,line=-2.2,adj=0.05,outer = F,padj = 0, col="#7FBC41")
mtext((paste("pseudo R2 (p-value)",sep = "")),side=3,line=-1,adj=0.05,outer = F,padj = 0, col="black")

# plot.decay.pretty(decay.Fun_sor_clus2nonF1.exp.sor, col="#7FBC41", add=T, main="non-F1",
#            xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
#            col.line="#276419",xaxp = c(0,1500,3),
#            ylab="",xlab="")

```



