---
title: '**RAM paired amplicons: Organizing Manuscript Analyses'
author: "Kelli Feeser"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
    fig.caption: yes
    keep_md: yes
  html_notebook:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
    fig.caption: yes
  html_document:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
    fig.caption: yes
editor_options:
  chunk_output_type: inline
---

\

::: homelink
<a href="https://kellifeeser.github.io/ksu-paired-amplicon-workflow/index.html" target="_blank" style="text-align:right">Back to Home</a>
:::

\

------------------------------------------------------------------------

Document last updated: `r Sys.Date()`

------------------------------------------------------------------------

\

# Set-up project naming from reproducible code {.unlisted .unnumbered .hidden}

```{r project-naming, include=F, eval=T}
PROJECT_PAGENAME <- 'HFD Microbiome - 12 week experiment'
PROJABBV <- 'HFD12wk'

DATASET1 <- 'Bac'
# DATASET2 <- 'Fun'

RDS1 <- 'hfd_12wk_min2_rare13006_metadataupdated_onlyCONvHFD'
RDS2 <- 'hfd_12wk_min2_rare13006'

PROJABBV_fig_path=paste0("../docs/figures/DA_plots/",PROJABBV)
```

# Set-up {.unlisted .unnumbered .hidden}

```{r global_options, include=FALSE}
# packages
library(knitr)
# library(kableExtra)

# set up global R options
  # to list the current values of all options but better to F1
    #options()
    # easier way:
      #?options
  # to access the value of a single option
    #getOption("width")

# set up knitr global chunk options
  # See str(knitr::opts_chunk$get()) for a list of default chunk options.
#knitr::opts_chunk$set(fig.width = 8)
# knitr::opts_chunk$set(fig.path = "../figures/darksidea_plots/")
#knitr::opts_chunk$set(fig.path = "figures/taxa_plots/prefix-")

# How to use an object before it is created
  # see https://bookdown.org/yihui/rmarkdown-cookbook/load-cache.html

# changing from cache.path="../docs/code/BacterialFungalClusterInteractions_cache/html/") to cache.path="../docs/github_BacterialFungal_Cluster_Interactions_files/figure-html/")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      cache.path="../docs/github_ManuscriptOrganizing_files/figure-html/")
opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(fig.pos = '!h')
# options(knitr.kable.NA = '')

## preload your R variables here ##
# variable_1 <- c("a", "b", "c")
####
```

```{r LANL-proxies, eval=T, include=FALSE}
# LANL proxies
Sys.setenv('HTTP_PROXY'='http://proxyout.lanl.gov:8080')
Sys.setenv('HTTPS_PROXY'='http://proxyout.lanl.gov:8080')
```

## load packages

```{r load-packages, eval=T, include=F}
library("ggplot2"); packageVersion("ggplot2") #‘3.4.4’
library("vegan"); packageVersion("vegan") # # ‘2.6.4’
library("phyloseq"); packageVersion("phyloseq") #‘1.44.0’
library("scales"); packageVersion("scales") #‘1.2.1’
library("swatches"); packageVersion("swatches") #‘0.5.0’
library("ggsci"); packageVersion("ggsci") #‘3.0.0’
library("ggpubr"); packageVersion("ggpubr") #‘0.6.0’
library("viridis"); packageVersion("viridis") #‘0.6.4’
library("RColorBrewer"); packageVersion("RColorBrewer") #‘1.1.3’
library("cowplot"); packageVersion("cowplot") #‘1.1.1’
library("dplyr"); packageVersion("dplyr") #‘1.1.3’
library("plyr"); packageVersion("plyr") #‘1.8.9’
library("knitr"); packageVersion("knitr") #‘1.44’
#library("kableExtra"); packageVersion("kableExtra") #‘1.3.4’
library("DT"); packageVersion("DT") #‘0.30’
library("tidyverse"); packageVersion("tidyverse") #‘2.0.0’
library("prettydoc"); packageVersion("prettydoc") #‘0.4.1’
library("magrittr"); packageVersion("magrittr") #‘2.0.3’
library("broom"); packageVersion("broom") #‘1.0.5’
library("betapart"); packageVersion("betapart") #‘1.6’
# patch see: https://stackoverflow.com/questions/76118194/error-when-loading-kableextra-in-markdown-file-after-updating-to-r-4-3-0
# devtools::install_github("kupietz/kableExtra",force = TRUE)
library("kableExtra")

library("microbiome"); packageVersion("microbiome") #‘1.22.0’
library("supernova"); packageVersion("supernova") #‘3.0.0’
library("lme4"); packageVersion("lme4") #‘1.1.34’
library("circlize"); packageVersion("circlize") #‘0.4.16’
library("ggordiplots"); packageVersion("ggordiplots") #‘0.4.3’

source("ksu_colors.R", local = knitr::knit_global())
source("../../black-sheep-club/code/sheepish_scripts.R", local = knitr::knit_global())
```

## load cache

```{r load-cache, include=FALSE, eval=T}
# OBJECT <- if (file.exists('OBJECT.rds')) {
#   readRDS('OBJECT.rds')
# } else {
#   "The value of `OBJECT` is not available yet"
# }

# knitr::load_cache('CHUNK-NAME', 'OBJECT')



# knitr::load_cache('adonis-predClus-Fun-calc', 'adonis_predClus_Fun')
# knitr::load_cache('adonis-predClus-Bac-calc', 'adonis_predClus_Bac')
# knitr::load_cache('calc-alpha', 'fwc_alpha')
# knitr::load_cache('calc-alpha', 'bwc_alpha')
# knitr::load_cache('gamma-fun-calc', 'Fun_gamma2')
# knitr::load_cache('gamma-bac-calc', 'Bac_gamma2')

# knitr::load_cache('mean-diss-bycluster-Fun', 'md_Fwc_self')
# knitr::load_cache('mean-diss-bycluster-Fun', 'md_Fwc_clus_sor_k10.df.noself')
# knitr::load_cache('mean-diss-bycluster-Bac', 'md_Bwc_self')
# knitr::load_cache('mean-diss-bycluster-Bac', 'md_Bwc_clus_sor_k14.df.noself')

knitr::load_cache('subset-Fun-wc-byClus', 'sor.dist.list')
knitr::load_cache('subset-Bac-wc-byClus')

knitr::load_cache('calc-betadisp')
knitr::load_cache('betadisp-boxplot-Fun', 'betadisp_clus_Fun.df')
# knitr::load_cache('', '')
# knitr::load_cache('', '')
# knitr::load_cache('', '')

```

## load rds

```{r load-rds, include=FALSE,eval=TRUE}
##########
# phylo #
Bac_wholecommunity<-readRDS(file = "../processed_data/clean_rds_saves/Bac_wholecommunity.rds")
Fun_wholecommunity <- readRDS(file="../processed_data/clean_rds_saves/Fun_wholecommunity.rds")
##########

##########
# nmds #
Bac_wc_sor_k5.ord <- readRDS(file = "../processed_data/clean_rds_saves/Bac_wc_sor_k5.ord.rds")
Fun_wc_sor_k4.ord <- readRDS(file = "../processed_data/clean_rds_saves/Fun_wc_sor_k4.ord.rds")
##########


##########
# misc #
Bac_wholecommunity.r<-readRDS(file = "../processed_data/clean_rds_saves/Bac_wholecommunity.r.rds")
Fun_wholecommunity.r<-readRDS(file = "../processed_data/clean_rds_saves/Fun_wholecommunity.r.rds")
sd<-as(sample_data(Fun_wholecommunity), "data.frame") # metadata for phyloseq
mv<-phylo2vegan_sd(Fun_wholecommunity) # metadata for vegan
##########

# metadata check
# unique(subset(sd,Site=="ONF")$Latitude) should be 34.80954


##########
# colors #
col_n14 <- c("B1"="#730025","B2"="#AF0039","B3"="#E8004C","B4"="#FF65A3","B5"="#9E4D97",
             "B6"="#4D42B3","B7"="#7678ed","B8"="#5B7898","B9"="#17A398","B10"="#74A57F",
             "B11"="#B6AF40","B12"="#f7b801", "B13"="#f18701", "B14"="#f35b04")
# creation: https://coolors.co/730025-AF0039-E8004C-FF65A3-9E4D97-4D42B3-7678ed-5B7898-17A398-74A57F-B6AF40-f7b801-f18701-f35b04


col_n10 <- c("F1"="#8E0152","F2"="#C51B7D","F3"="#DE77AE","F4"="#F1B6DA","F5"="#C5C2EA",
             "F6"="#8DC9BB","F7"="#86C37E","F8"="#7FBC41","F9"="#4D9221","F10"="#276419")
             
                 # "#DBBCE2","F6"="#E6F5D0","F7"="#B8E186","F8"="#7FBC41","F9"="#4D9221","F10"="#276419")

             # "#B77EBE","#B0AFDF","#92CDC8","#7BB990","#EFBC9B","#DBA8AC"

# creation: https://coolors.co/8e0152-c51b7d-de77ae-f1b6da-c5c2ea-a8e0ff-8dc9bb-7fbc41-4d9221-276419
col_clusn10 <- c("F1"="#8E0152","F2"="#C51B7D","F3"="#DE77AE","F4"="#F1B6DA","F5"="#FDE0EF",
                 "F6"="#E6F5D0","F7"="#B8E186","F8"="#7FBC41","F9"="#4D9221","F10"="#276419")

# change "F5"="#FDE0EF" to "#C5C2EA",
  # "F6"="#E6F5D0" to "#8DC9BB", 
  # "F7"="#B8E186" to "#86C37E"

latlong_cols<-readRDS(file = "../processed_data/clean_rds_saves/latlong_cols.rds")
##########
```

## Read external scripts into a chunk

```{r, file=c('one.R', 'two.R')}
```

## Read multiple code chunks from an external script (*)


```{r, include=FALSE,eval=FALSE}
# https://bookdown.org/yihui/rmarkdown-cookbook/read-chunk.html

# in source.R file:
# To label a block of code in a script, you write the label after ## ---- (optionally, you can add a series of dashes to the end of this line).

## ---- test-a --------
1 + 1

## ---- test-b --------
if (TRUE) {
  plot(cars)
}


# We assume that the filename of the above script is test.R. In the R Markdown document, we can read it via knitr::read_chunk(), and use the code in code chunks with the labels, e.g.,

# Read an external script:

  # ```{r, include=FALSE, cache=FALSE}
  # knitr::read_chunk('test.R')

# Now we can use the code, e.g.,

  # ```{r, test-a, echo=FALSE}

  # ```{r, test-b, fig.height=4}


# Note that we use knitr::read_chunk() mainly for its side effect, so please make sure the code chunk in which you call this function is not cached
```

