---
title: "R Notebook"
output: html_notebook
---


```{r auto-populate-rank-names-to-plot, results='asis'}
Bac.rank.namkes<-as.vector(rank_names(Bac))
min_nname_Bac_Phylum<-length(unique(Bac_Phylum_Mean_Abund$Phylum[Bac_Phylum_Mean_Abund$Bac_Phylum_Mean > 0.5]))

Bac.rank.names<-rank_names(Bac)[2:6]
Fun.rank.names<-rank_names(Fun)[2:7]


for(rank in Bac.rank.names) {
  
  cat('\n # Set min plotting threshold to 0.5% relative abundance (all-sample a.k.a global mean) \n')
  
  cat('\n')
  
   cat('min_nname_Bac_', rank, '<-length(unique(Bac_', rank, '_Mean_Abund$', 
       rank, '[Bac_', rank, '_Mean_Abund$Bac_', rank, '_Mean > 0.5]))',
       sep = '')
  
  cat('\n \n ')

    cat('if (min_nname_Bac_', rank, ' < 10){
  ) (Bac.', rank, '.Phylum.names.to.plot<-top_n(Bac_', rank, '_Mean_Abund, min_nname_Bac_', rank, ', Bac.', rank, '_Mean)$',rank, ' 
} else {
  Bac.', rank, '.names.to.plot<-top_n(Bac_', rank, '_Mean_Abund, 10, Bac.', rank, '_Mean)$', rank, '
}', sep = '')


    cat('\n \n')

}

```


## AllG (All Grasses) {.unnumbered}
## ANGE (*Andropogon gerardii* / Big bluestem) {.unnumbered}
## BOER (*B. eriopoda* / Black grama) {.unnumbered}
## BOGR (*B. gracilis* / Blue grama) {.unnumbered}
## BUDA (*Bouteloua dactyloides* / Buffalograss) {.unnumbered}
## SCSC (*Schizachyrium scoparium* / Little bluestem) {.unnumbered}


```{r, eval=T}
Fun.AllG_otu <- as.data.frame(t(otu_table(Fun.AllG)))

# Read in the OTU table for all of the data and transpose with R	
	#survey.all = read.csv('../processed_data/AL_Survey.all.Fun_97_nosing.csv', header=T, row.names=1)

# Create presence absence matrix
Fun.AllG_otu.pres.abs = decostand(x=Fun.AllG_otu, method="pa")

# Create distance matrix for beta diversity metrics, index.family="sorensen"
Fun.AllG_otu.dist = beta.pair(Fun.AllG_otu.pres.abs, index.family="sor")

# Total dissimilarity: pair-wise Sorensen dissimilarity (monotonic transformation of beta div)
	Fun.AllG_otu.dist.sor.matrix <- matrix(Fun.AllG_otu.dist$beta.sor)
	# length(Fun.AllG_otu.dist.sor.matrix)
	if (file.exists('../processed_data/nestedness/Fun.AllG_otu.dist.sor.matrix.txt')){
	  # 
	} else {
    write.table(Fun.AllG_otu.dist.sor.matrix, file = '../processed_data/nestedness/Fun.AllG_otu.dist.sor.matrix.txt')
  }
  
# Spatial turnover / species replacement: Simpson 
	Fun.AllG_otu.dist.sim.matrix <- matrix(Fun.AllG_otu.dist$beta.sim)
	# length(Fun.AllG_otu.dist.sim.matrix)
		if (file.exists('../processed_data/nestedness/survey.dist.sim.matrix.txt')){
	  # 
	} else {
    write.table(Fun.AllG_otu.dist.sim.matrix, file = '../processed_data/nestedness/Fun.AllG_otu.dist.sim.matrix.txt')
  }

# Nestedness-resultant dissimilarity
	Fun.AllG_otu.dist.sne.matrix <- matrix(Fun.AllG_otu.dist$beta.sne)
	# length(Fun.AllG_otu.dist.sne.matrix)
	if (file.exists('../processed_data/nestedness/survey.dist.sne.matrix.txt')){
	  # 
	} else {
    write.table(Fun.AllG_otu.dist.sne.matrix, file = '../processed_data/nestedness/Fun.AllG_otu.dist.sne.matrix.txt')
  }

Fun.AllG_soil_moisture.sor.mrm <- MRM(Fun.AllG_otu.dist$beta.sor ~ Fun.soil_moisture.matrix, nperm=1000)
Fun.AllG_soil_moisture.sim.mrm <- MRM(Fun.AllG_otu.dist$beta.sim ~ Fun.soil_moisture.matrix, nperm=1000)
Fun.AllG_soil_moisture.sne.mrm <- MRM(Fun.AllG_otu.dist$beta.sne ~ Fun.soil_moisture.matrix, nperm=1000)

Fun.AllG_soil_moisture.sor.b.m <- c(Fun.AllG_soil_moisture.sor.mrm$coef[1,1], Fun.AllG_soil_moisture.sor.mrm$coef[2,1])
Fun.AllG_soil_moisture.sim.b.m <- c(Fun.AllG_soil_moisture.sim.mrm$coef[1,1], Fun.AllG_soil_moisture.sim.mrm$coef[2,1])
Fun.AllG_soil_moisture.sne.b.m <- c(Fun.AllG_soil_moisture.sne.mrm$coef[1,1], Fun.AllG_soil_moisture.sne.mrm$coef[2,1])

```


# cut occurrence code

# occurrence - scripts dont work, go back to Darksidea to find one that does

  #r occurrence-calc-function-notworking, include=F,eval=F}
    #calc_occurrence <- function(phylo_obj, metadata_factor, tax_rank, taxa_of_interest = NULL) {
    #   if (!is.null(taxa_of_interest)) {
    #     # Filter OTU table based on taxa_of_interest at OTU level if provided
    #     if (tax_rank == "OTU") {
    #       taxa_indices <- taxa_of_interest
    #       filtered_phylo_obj <- prune_taxa(phylo_obj, taxa_indices)
    #     } else {
    #       # Filter OTU table based on taxa_of_interest at tax_rank level if provided
    #       taxa_indices <- taxa_names(phylo_obj)[tax_table(phylo_obj)[, tax_rank] %in% taxa_of_interest]
    #       filtered_phylo_obj <- prune_taxa(phylo_obj, taxa_indices)
    #     }
    #   } else {
    #     filtered_phylo_obj <- phylo_obj
    #   }
    #   
    #   # Calculate occurrence matrix
    #   occurrence_matrix <- table(sample_data(filtered_phylo_obj)[, metadata_factor], tax_table(filtered_phylo_obj)[, tax_rank]) > 0
    #   
    #   # Calculate frequency of occurrence by factor level
    #   frequency_by_factor <- apply(occurrence_matrix, 1, function(x) sum(x) / length(x))
    #   
    #   return(frequency_by_factor)
    # }


  # #r occurrence-calc-function-notworking2, include=F,eval=F}
  # calc_occurrence.test <- function(phylo_obj, metadata_factor, tax_rank, taxa_of_interest = NULL) {
  #   if (!is.null(taxa_of_interest)) {
  #     # Filter OTU table based on taxa_of_interest at OTU level if provided
  #     if (tax_rank == "OTU") {
  #       taxa_indices <- taxa_of_interest
  #       filtered_phylo_obj <- prune_taxa(phylo_obj, taxa_indices)
  #     } else {
  #       # Filter OTU table based on taxa_of_interest at tax_rank level if provided
  #       taxa_indices <- taxa_names(phylo_obj)[tax_table(phylo_obj)[, tax_rank] %in% taxa_of_interest]
  #       filtered_phylo_obj <- prune_taxa(phylo_obj, taxa_indices)
  #     }
  #   } else {
  #     filtered_phylo_obj <- phylo_obj
  #   }
  #   
  #   # Calculate occurrence matrix
  #   occurrence_matrix <- table(sample_data(filtered_phylo_obj)[, metadata_factor], tax_table(filtered_phylo_obj)[, tax_rank]) > 0
  #   
  #   # Calculate frequency of occurrence by factor level
  #   frequency_by_factor <- apply(occurrence_matrix, 1, function(x) sum(x) / length(x))
  #   
  #   return(frequency_by_factor)
  # }
  # 
  # calc_occurrence.test <- function(phylo_obj, metadata_factor, tax_rank, taxa_of_interest = NULL) {
  #   if (!is.null(taxa_of_interest)) {
  #     # Filter OTU table based on taxa_of_interest at OTU level if provided
  #     if (tax_rank == "OTU") {
  #       taxa_indices <- taxa_of_interest
  #       filtered_phylo_obj <- prune_taxa(phylo_obj, taxa_indices)
  #     } else {
  #       # Filter OTU table based on taxa_of_interest at tax_rank level if provided
  #       taxa_indices <- taxa_names(phylo_obj)[tax_table(phylo_obj)[, tax_rank] %in% taxa_of_interest]
  #       #filtered_phylo_obj <- prune_taxa(phylo_obj, taxa_indices)
  #     }
  #   } else {
  #     filtered_phylo_obj <- phylo_obj
  #   }
  #   
  #   # Calculate occurrence matrix
  #   occurrence_matrix <- table(sample_data(filtered_phylo_obj)[, metadata_factor], tax_table(filtered_phylo_obj)[, tax_rank]) > 0
  #   
  #   # Calculate frequency of occurrence by factor level
  #   frequency_by_factor <- apply(occurrence_matrix, 1, function(x) sum(x) / length(x))
  #   
  #   return(frequency_by_factor)
  # }
  # taxa_at_rank <- taxa_names(phylo_obj)[, tax_rank]
  # taxa_indices <- taxa_at_rank[taxa_at_rank %in% taxa_of_interest]
  # filtered_phylo_obj <- prune_taxa(c(taxa_indices),Fun)
  # 
  # #working
  # #input (Fun, "SampleID", "Genus", c("Darksidea"))
  # taxa_at_rank <- tax_table(Fun)[, c("Genus")]
  # taxa_indices <- as.vector(unique(as.data.frame(taxa_at_rank[taxa_at_rank %in% c("Darksidea")])[,1]))
  # filtered_phylo_obj <- prune_taxa(taxa_indices,Fun)
  # taxa_names(Fun)
  # head(tax_table(Fun)[, c("Genus","Species")])
  # 
  # 
  # 
  # #if "genus" and output is otu
  # taxa_at_rank <- tax_table(Fun)[, c("Genus")]
  # taxa_indices <- row.names(as.data.frame(taxa_at_rank[taxa_at_rank %in% c("Darksidea")]))
  # filtered_phylo_obj <- prune_taxa(taxa_indices,Fun)
  # head(otu_table(filtered_phylo_obj))
  # sd_f<-sample_data(filtered_phylo_obj)[, "SampleID"]
  # head(sample_data(filtered_phylo_obj))
  # occurrence_matrix <- table(sample_data(filtered_phylo_obj)[, "SampleID"], tax_table(filtered_phylo_obj)[, "Genus"]) > 0
  # 
  # taxa_at_rank <- tax_table(Fun)[, c("Genus")]
  # taxa_indices <- as.vector(unique(as.data.frame(taxa_at_rank[taxa_at_rank %in% c("Darksidea")])[,1]))
  # filtered_phylo_obj <- prune_taxa(taxa_indices,Fun)
  # 
  # #working
  # #input (Fun, "SampleID", "OTU", c("Darksidea"))
  # 
  # 
  # 
  # 
  # taxa_indices<-taxa_names(phylo_obj)[tax_table(phylo_obj)[, "Genus"] %in% c("Darksidea")]
  # filtered_phylo_obj <- prune_taxa(c(taxa_indices),Fun)
  # 
  # occ_genus_Darksidea_sample <- calc_occurrence.test(Fun, "SampleID", "Genus", c("Darksidea"))
  # occ_genus_Darksidea_sample
  # 
  # 
  # 
  # #if tax_rank == rank of taxa_of_interest
  # #if tax_rank != rank of taxa_of_interest
  # #rank of taxa_of_interest is genus and tax_rank = OTU
  # #rank of taxa_of_interest is genus and tax_rank = Species
  # #if is.null(taxa_of_interest)
  # 
  # phylo_obj<-Fun
  # tax_rank<-"Genus"
  # taxa_of_interest<-c("Darksidea")
  # metadata_factor<-"Grass"

  #this works best rn - still ignore
  # calc_occurrence <- function(phylo_obj, metadata_factor, tax_rank, taxa_of_interest = NULL) {
  #   
  #   print(phylo_obj)
  #   tax_table_cols <- colnames(tax_table(phylo_obj))
  #   
  #   if (metadata_factor=="SampleID") {
  #     stop("Don't use 'SampleID', try 'sample_name' instead")
  #   } else if (!is.null(taxa_of_interest)) {
  #     
  #     if (tax_rank %in% tax_table_cols) {
  #       taxa_col_indices <- which(apply(tax_table(phylo_obj), 2, function(x) any(x %in% taxa_of_interest)))
  #       
  #       if (length(taxa_col_indices) > 0) {
  #         rank_interesting_taxa <- tax_table_cols[max(taxa_col_indices)]
  #         interesting_otu_names<-taxa_names(phylo_obj)[tax_table(phylo_obj)[, tax_rank] %in% taxa_of_interest]
  #         if (rank_interesting_taxa == rank_interesting_taxa && tax_rank == "OTU") {
  #           filtered_phylo_obj <- prune_taxa(interesting_otu_names, phylo_obj)
  #           print("Computing occurrence at OTU-level")
  #         } else if (tax_rank == rank_interesting_taxa && tax_rank == "Species") {
  #           filtered_phylo_obj <- prune_taxa(interesting_otu_names, phylo_obj)
  #           filtered_phylo_obj <- tax_glom(filtered_phylo_obj, taxrank = "Species")
  #           print("Agglomerating to rank 'Species'")
  #         } else if (tax_rank == rank_interesting_taxa && tax_rank == "Genus") {
  #           filtered_phylo_obj <- prune_taxa(interesting_otu_names, phylo_obj)
  #           filtered_phylo_obj <- tax_glom(filtered_phylo_obj, taxrank = "Genus")
  #           print("Agglomerating to rank 'Genus'")
  #         } else {
  #           stop("Invalid combination of taxonomic ranks and taxa_of_interest provided.")
  #         }
  #       } else {
  #         stop("Taxa of interest not found in the taxonomic table.")
  #       }
  #     } else {
  #       stop("Taxonomic ranks required for filtering not found in the phylo_obj.")
  #     }
  #   } else {
  #     filtered_phylo_obj <- phylo_obj
  #     print("No taxa of interest input, computing occurrence for entire OTU table")
  #   }
  #   
  #   # print(filtered_phylo_obj)
  #   
  #   #subset metadata
  #   occ_meta<-sample_data(filtered_phylo_obj)
  #   #occ_meta
  #   #print("Full metadata")
  #   #print(head(occ_meta))
  #   occ_meta_sub <- subset(occ_meta, select = c("SampleID",metadata_factor))
  #   #print(head(occ_meta_sub))
  #   
  #   # Calculate occurrence matrix, convert to presence/absence
  #   occ_matrix <- otu_table(filtered_phylo_obj) %>%
  #     as.data.frame() %>%
  #     t() %>%
  #     {ifelse(. > 1, 1, .)} %>%
  #     as.data.frame() %>%
  #     rownames_to_column(var = "SampleID") %>%
  #     inner_join(occ_meta_sub, by = "SampleID") %>%
  #     select(SampleID,last_col(),c(everything(),-last_col(),SampleID))
  #   print("Sanity check: preview of occurrence matrix")
  #   print(head(occ_matrix))
  #   
  #   # Calculate frequency of occurrence by factor level
  #   frequency_by_factor <- occ_matrix %>%
  #     group_by(.data[[metadata_factor]]) %>%
  #     summarise_at(vars(starts_with("OTU")), 
  #                  ~ sum(. > 0) / n(), 
  #                  .names = "{.col}_frequency",
  #                  .preserve = T) %>%
  #     ungroup()
  #   print("Sanity check: preview of occurrence by factor")
  #   print(head(frequency_by_factor))
  #   
  #   return(frequency_by_factor)
  # }
  # 
  # # Example usage with taxa_of_interest specified at the genus level
  # # Replace 'phylo_obj', 'SampleType' (metadata factor), and 'Genus' (taxonomic rank)
  # result_genus <- calc_occurrence(Fun, "Grass", "Genus", c("Darksidea"))
  # 
  # # Example usage with taxa_of_interest specified at the OTU level
  # # Replace 'phylo_obj', 'SampleType' (metadata factor), and 'otus_of_interest' (list of OTUs)
  # result_otu <- calc_occurrence(Fun, "Gr", "OTU", otus_of_interest)





  #r occurrence-calc-ignore, include=F,eval=F}
    # str(sample_data(Fun))
    # Fun_sd <- sample_data(Fun)
    # head(Fun_sd)
    # 
    # metadata_factor<-"Grass"
    # 
    # meta_sub <- subset(Fun_sd, select = c("SampleID",metadata_factor))
    # phylo_obj<-Fun
    # tax_rank<-"Genus"
    # taxa_of_interest<-c("Darksidea")
    # 
    # # Example usage with taxa_of_interest specified at the genus level
    # # Replace 'phylo_obj', 'SampleType' (metadata factor), and 'Genus' (taxonomic rank)
    # occ_genus_Darksidea_sample <- calc_occurrence(Fun, "SampleID", "Genus", c("Darksidea"))
    # head(sample_data())
    # # Example usage with taxa_of_interest specified at the OTU level
    # # Replace 'phylo_obj', 'SampleType' (metadata factor), and 'otus_of_interest' (list of OTUs)
    # result_otu <- calc_occurrence(phylo_obj, "SampleType", "OTU", otus_of_interest)
