---
title: "R Notebook"
output: html_notebook
---

Latitudinal Bin
Longitudinal Gradient

colorBin
colorGradient

bin_cols_n5
grad_cols_n3

color = "Bin"
color = "Gradient"

Bin_whole word
Grad_whole word

colGrad
colGrad

# ordinations

```{r plot nmds AllG col Grad, include=F, eval=F}

nmds.Bac.AllG.colGrad_wleg<-plot_ordination(Bac.AllG, Bac.AllG.k5.ord, color = "Gradient") +
  scale_color_manual(values = bac_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9)+
  stat_ellipse(type = "t", level = 0.9) +
  ggtitle("NMDS by Longitudinal Gradient (90% confidence level for multivariate t-distribution)", subtitle = "16S: All Grass Hosts") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="bottom",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

leg_nmds.colGrad<-get_legend(nmds.Bac.AllG.colGrad_wleg)
nmds.Bac.AllG.colGrad<-nmds.Bac.AllG.colGrad_wleg+theme(legend.position="none")

nmds.Fun.AllG.colGrad<-plot_ordination(Fun.AllG, Fun.AllG.k5.ord, color = "Gradient") +
  scale_color_manual(values = fun_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9) +
  stat_ellipse(type = "t", level = 0.9) +
  labs(y="") +
  ggtitle("", subtitle = "ITS: All Grass Hosts") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

nmds.AllG.colGrad_upper<-plot_grid(nmds.Bac.AllG.colGrad,nmds.Fun.AllG.colGrad, ncol=2)
plot_grid(nmds.AllG.colGrad_upper,leg_nmds.colGrad, nrow=2, rel_heights = c(0.85,0.15))

ggsave(filename = "../analyses/ordinations/NMDS_AllG_colorGradient.png", plot = last_plot(), 
       device = "png", height = 4, width = 7, units = "in")
```

```{r plot nmds ANGE col Grad, include=F, eval=F}

  Bac.ANGE.k4.ord
  Fun.ANGE.k5.ord # 4-7

nmds.Bac.ANGE.colGrad_wleg<-plot_ordination(Bac.ANGE, Bac.ANGE.k5.ord, color = "Gradient") +
  scale_color_manual(values = bac_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9)+
  stat_ellipse(type = "t", level = 0.9) +
  ggtitle("NMDS by Longitudinal Gradient (90% confidence level for multivariate t-distribution)", subtitle = "16S: Grass Host = ANGE") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="bottom",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

leg_nmds.colGrad<-get_legend(nmds.Bac.ANGE.colGrad_wleg)
nmds.Bac.ANGE.colGrad<-nmds.Bac.ANGE.colGrad_wleg+theme(legend.position="none")

nmds.Fun.ANGE.colGrad<-plot_ordination(Fun.ANGE, Fun.ANGE.k5.ord, color = "Gradient") +
  scale_color_manual(values = fun_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9) +
  stat_ellipse(type = "t", level = 0.9) +
  labs(y="") +
  ggtitle("", subtitle = "ITS: Grass Host = ANGE") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

nmds.ANGE.colGrad_upper<-plot_grid(nmds.Bac.ANGE.colGrad,nmds.Fun.ANGE.colGrad, ncol=2)
plot_grid(nmds.ANGE.colGrad_upper,leg_nmds.colGrad, nrow=2, rel_heights = c(0.85,0.15))

ggsave(filename = "../analyses/ordinations/NMDS_ANGE_colorGradient.png", plot = last_plot(), 
       device = "png", height = 4, width = 7, units = "in")
```

```{r plot nmds BOER col Grad, include=F, eval=F}
subset(ord_checkfit_df,ord_checkfit_df$input_data==comm.names.by.grass[5])
subset(ord_checkfit_df,ord_checkfit_df$input_data==comm.names.by.grass[6])

nmds.Bac.BOER.colGrad_wleg<-plot_ordination(Bac.BOER, Bac.BOER.k5.ord, color = "Gradient") +
  scale_color_manual(values = bac_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9)+
  stat_ellipse(type = "t", level = 0.9) +
  ggtitle("NMDS by Longitudinal Gradient (90% confidence level for multivariate t-distribution)", subtitle = "16S: Grass Host = BOER") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="bottom",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

leg_nmds.colGrad<-get_legend(nmds.Bac.BOER.colGrad_wleg)
nmds.Bac.BOER.colGrad<-nmds.Bac.BOER.colGrad_wleg+theme(legend.position="none")

nmds.Fun.BOER.colGrad<-plot_ordination(Fun.BOER, Fun.BOER.k5.ord, color = "Gradient") +
  scale_color_manual(values = fun_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9) +
  stat_ellipse(type = "t", level = 0.9) +
  labs(y="") +
  ggtitle("", subtitle = "ITS: Grass Host = BOER") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

nmds.BOER.colGrad_upper<-plot_grid(nmds.Bac.BOER.colGrad,nmds.Fun.BOER.colGrad, ncol=2)
plot_grid(nmds.BOER.colGrad_upper,leg_nmds.colGrad, nrow=2, rel_heights = c(0.85,0.15))

ggsave(filename = "../analyses/ordinations/NMDS_BOER_colorGradient.png", plot = last_plot(), 
       device = "png", height = 4, width = 7, units = "in")
```

```{r plot nmds BOGR col Grad, include=F, eval=F}
subset(ord_checkfit_df,ord_checkfit_df$input_data==comm.names.by.grass[7])
subset(ord_checkfit_df,ord_checkfit_df$input_data==comm.names.by.grass[8])

nmds.Bac.BOGR.colGrad_wleg<-plot_ordination(Bac.BOGR, Bac.BOGR.k5.ord, color = "Gradient") +
  scale_color_manual(values = bac_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9)+
  stat_ellipse(type = "t", level = 0.9) +
  ggtitle("NMDS by Longitudinal Gradient (90% confidence level for multivariate t-distribution)", subtitle = "16S: Grass Host = BOGR") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="bottom",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

leg_nmds.colGrad<-get_legend(nmds.Bac.BOGR.colGrad_wleg)
nmds.Bac.BOGR.colGrad<-nmds.Bac.BOGR.colGrad_wleg+theme(legend.position="none")

nmds.Fun.BOGR.colGrad<-plot_ordination(Fun.BOGR, Fun.BOGR.k5.ord, color = "Gradient") +
  scale_color_manual(values = fun_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9) +
  stat_ellipse(type = "t", level = 0.9) +
  labs(y="") +
  ggtitle("", subtitle = "ITS: Grass Host = BOGR") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

nmds.BOGR.colGrad_upper<-plot_grid(nmds.Bac.BOGR.colGrad,nmds.Fun.BOGR.colGrad, ncol=2)
plot_grid(nmds.BOGR.colGrad_upper,leg_nmds.colGrad, nrow=2, rel_heights = c(0.85,0.15))

ggsave(filename = "../analyses/ordinations/NMDS_BOGR_colorGradient.png", plot = last_plot(), 
       device = "png", height = 4, width = 7, units = "in")
```

```{r plot nmds BUDA col Grad, include=F, eval=F}
subset(ord_checkfit_df,ord_checkfit_df$input_data==comm.names.by.grass[9])
subset(ord_checkfit_df,ord_checkfit_df$input_data==comm.names.by.grass[10])

#flip y coord of ITS Fun.BUDA.k6.ord

nmds.Bac.BUDA.colGrad_wleg<-plot_ordination(Bac.BUDA, Bac.BUDA.k6.ord, color = "Gradient") +
  scale_color_manual(values = bac_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9)+
  stat_ellipse(type = "t", level = 0.9) +
  ggtitle("NMDS by Longitudinal Gradient (90% confidence level for multivariate t-distribution)", subtitle = "16S: Grass Host = BUDA") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="bottom",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

leg_nmds.colGrad<-get_legend(nmds.Bac.BUDA.colGrad_wleg)
nmds.Bac.BUDA.colGrad<-nmds.Bac.BUDA.colGrad_wleg+theme(legend.position="none")

nmds.Fun.BUDA.colGrad<-plot_ordination(Fun.BUDA, Fun.BUDA.k6.ord, color = "Gradient") +
  scale_color_manual(values = fun_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9) +
  stat_ellipse(type = "t", level = 0.9) +
  labs(y="") +
  ggtitle("", subtitle = "ITS: Grass Host = BUDA") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

nmds.BUDA.colGrad_upper<-plot_grid(nmds.Bac.BUDA.colGrad,nmds.Fun.BUDA.colGrad, ncol=2)
plot_grid(nmds.BUDA.colGrad_upper,leg_nmds.colGrad, nrow=2, rel_heights = c(0.85,0.15))

ggsave(filename = "../analyses/ordinations/NMDS_BUDA_colorGradient.png", plot = last_plot(), 
       device = "png", height = 4, width = 7, units = "in")
```

```{r plot nmds SCSC col Grad, include=F, eval=F}
subset(ord_checkfit_df,ord_checkfit_df$input_data==comm.names.by.grass[11])
subset(ord_checkfit_df,ord_checkfit_df$input_data==comm.names.by.grass[12])

nmds.Bac.SCSC.colGrad_wleg<-plot_ordination(Bac.SCSC, Bac.SCSC.k5.ord, color = "Gradient") +
  scale_color_manual(values = bac_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9)+
  stat_ellipse(type = "t", level = 0.9) +
  ggtitle("NMDS by Longitudinal Gradient (90% confidence level for multivariate t-distribution)", subtitle = "16S: Grass Host = SCSC") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="bottom",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    legend.title  = element_text(color="black", size=14),
    legend.text  = element_text(color="black", size=12),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

leg_nmds.colGrad<-get_legend(nmds.Bac.SCSC.colGrad_wleg)
nmds.Bac.SCSC.colGrad<-nmds.Bac.SCSC.colGrad_wleg+theme(legend.position="none")

nmds.Fun.SCSC.colGrad<-plot_ordination(Fun.SCSC, Fun.SCSC.k5.ord, color = "Gradient") +
  scale_color_manual(values = fun_grad_cols_n3,
                     name="Longitudinal Gradient") +
  geom_point(size = 3, alpha = 0.9) +
  stat_ellipse(type = "t", level = 0.9) +
  labs(y="") +
  ggtitle("", subtitle = "ITS: Grass Host = SCSC") +
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x = element_text(color="black", size=12),
    axis.text.y = element_text(color="black", size=12),
    legend.position="none",
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14),
    panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
  )

nmds.SCSC.colGrad_upper<-plot_grid(nmds.Bac.SCSC.colGrad,nmds.Fun.SCSC.colGrad, ncol=2)
plot_grid(nmds.SCSC.colGrad_upper,leg_nmds.colGrad, nrow=2, rel_heights = c(0.85,0.15))

ggsave(filename = "../analyses/ordinations/NMDS_SCSC_colorGradient.png", plot = last_plot(), 
       device = "png", height = 4, width = 7, units = "in")
```

############################################################################


############################################################################
# taxonomy

## full dataset: grouped by Cluster

### Fun 

```{r Fun-Phylum-Cluster, message=F, warning=F, error=F, fig.dim = c(12, 9), out.width = "100%", eval=F, echo=F}
# facet site

ggplot(dat_Fun.Phylum.r, aes(x = sample_name, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Phylum)) + 

  facet_wrap("Fun_sor_clusters", nrow=2, ncol=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nsample name', y="Relative Abundance", fill="Phylum")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = c("≤ 0.0002% relative abundance" = '#28A48D', 
            "Ascomycota"="#8569D5",
           "Basidiomycota"="#D1A33D",
           "Mucoromycota"="#5E738F")) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Phylum_Full_sample_cluster.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```

```{r}
ggplot(dat_Fun.Phylum.r, aes(x = sample_name, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Phylum)) + 

  facet_wrap("Fun_sor_clusters", nrow=2, ncol=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nsample name', y="Relative Abundance", fill="Phylum")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = c("≤ 0.0002% relative abundance" = '#28A48D', 
            "Ascomycota"="#8569D5",
           "Basidiomycota"="#D1A33D",
           "Mucoromycota"="#5E738F")) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Phylum_Full_sample_cluster.png", plot = last_plot(), width=12, height=9, units = "in", device="png")


ggplot(dat_Fun.Class.r, aes(x = sample_name,y = as.numeric(Abundance*100), fill = Class)) + 
  facet_wrap("Fun_sor_clusters", nrow=2, ncol=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nsample_name', y="Relative Abundance", fill="Class")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Class_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Class_Full_sample_cluster.png", plot = last_plot(), width=12, height=9, units = "in", device="png")

ggplot(dat_Fun.Order.r, aes(x = sample_name,y = as.numeric(Abundance*100),fill = Order)) + 
  facet_wrap("Fun_sor_clusters", nrow=2, ncol=5, scales = "free_x")+
  labs(x= '\nsample_name', y="Relative Abundance", fill="Order")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Order_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Order_Full_sample_cluster.png", plot = last_plot(), width=12, height=9, units = "in", device="png")


ggplot(dat_Fun.Family.r, aes(x = sample_name, y = as.numeric(Abundance*100),  fill = Family)) + 
  facet_wrap("Fun_sor_clusters", nrow=2, ncol=5, scales = "free_x")+
  labs(x= '\nsample_name', y="Relative Abundance", fill="Family")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Family_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Family_Full_sample_cluster.png", plot = last_plot(), width=12, height=9, units = "in", device="png")


ggplot(dat_Fun.Genus.r, aes(x = sample_name, y = as.numeric(Abundance),  fill = Genus)) + 
  facet_wrap("Fun_sor_clusters", nrow=2, ncol=5, scales = "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nsample_name', y="Relative Abundance", fill="Genus")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Genus_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Genus_Full_sample_cluster.png", plot = last_plot(), width=12, height=9, units = "in", device="png")

GvS_cols<-c("#d01c8b","#4dac26","grey")

ggplot(dat_Fun.Sign.r, aes(x = sample_name, y = as.numeric(Abundance*100),fill = Sign)) + 
  facet_wrap("Fun_sor_clusters", nrow=2, ncol=5, scales = "free_x")+
  labs(x= '\nsample_name', y="Relative Abundance", fill="Sign")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singleton" = "grey"),
    guide = "legend"
  ) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Sign_Full_sample_cluster.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```

### Bac

```{r}
ggplot(dat_Bac.Phylum.r, aes(x = sample_name,y = as.numeric(Abundance),fill = Phylum)) + 
  facet_wrap("Bac_sor_clusters", nrow=3, ncol=5, scales = "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nsample name', y="Relative Abundance", fill="Phylum")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Bac_Phylum_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Bac_Phylum_Full_sample_cluster.png", plot = last_plot(), width=9, height=7, units = "in", device="png")


ggplot(dat_Bac.Class.r, aes(x = sample_name,y = as.numeric(Abundance), fill = Class)) + 
  facet_wrap("Bac_sor_clusters", nrow=3, ncol=5, scales = "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nsample_name', y="Relative Abundance", fill="Class")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Bac_Class_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Bac_Class_Full_sample_cluster.png", plot = last_plot(), width=9, height=7, units = "in", device="png")

ggplot(dat_Bac.Order.r, aes(x = sample_name,y = as.numeric(Abundance),fill = Order)) + 
  facet_wrap("Bac_sor_clusters", nrow=3, ncol=5, scales = "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nsample_name', y="Relative Abundance", fill="Order")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Bac_Order_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Bac_Order_Full_sample_cluster.png", plot = last_plot(), width=9, height=7, units = "in", device="png")


ggplot(dat_Bac.Family.r, aes(x = sample_name, y = as.numeric(Abundance),  fill = Family)) + 
  facet_wrap("Bac_sor_clusters", nrow=3, ncol=5, scales = "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nsample_name', y="Relative Abundance", fill="Family")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Bac_Family_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Bac_Family_Full_sample_cluster.png", plot = last_plot(), width=9, height=7, units = "in", device="png")


ggplot(dat_Bac.Genus.r, aes(x = sample_name, y = as.numeric(Abundance),  fill = Genus)) + 
  facet_wrap("Bac_sor_clusters", nrow=3, ncol=5, scales = "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nsample_name', y="Relative Abundance", fill="Genus")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Bac_Genus_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Bac_Genus_Full_sample_cluster.png", plot = last_plot(), width=9, height=7, units = "in", device="png")
```

## full dataset: grouped by Cluster2 (f1 vs non-f1)

### Fun 

```{r}
ggplot(dat_Fun.Phylum.r, aes(x = Site_Grass, y = as.numeric(Abundance),
                             fill = Phylum)) + 
  # facet_wrap("Fun_sor_clus2", ncol=2, scales = "free_x")+
  facet_grid(.~ Fun_sor_clus2, scales = "free_x", space= "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nSite_Grass', y="Relative Abundance", fill="Phylum")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = c("≤ 0.0002% relative abundance" = '#28A48D', 
            "Ascomycota"="#8569D5",
           "Basidiomycota"="#D1A33D",
           "Mucoromycota"="#5E738F")) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Phylum_Full_sitegrass_cluster2.png", plot = last_plot(), width=12, height=9, units = "in", device="png")


ggplot(dat_Fun.Class.r, aes(x = Site_Grass,y = as.numeric(Abundance), fill = Class)) + 
  facet_grid(.~ Fun_sor_clus2, scales = "free_x", space= "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nSite_Grass', y="Relative Abundance", fill="Class")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Class_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Class_Full_sitegrass_cluster2.png", plot = last_plot(), width=12, height=9, units = "in", device="png")

#### Order #### 
ggplot(dat_Fun.Order.r, aes(x = Site_Grass, y = as.numeric(Abundance*100), 
                                                    fill = Order)) + 
  facet_grid(.~ Fun_sor_clus2, scales = "free_x", space= "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nSite_Grass', y="Relative Abundance", fill="Order")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Order_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Order_Full_sitegrass_cluster2.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
#### Order #### 


#### Family #### 
ggplot(dat_Fun.Family.r, aes(x = Site_Grass,y = as.numeric(Abundance*100), 
                                                    fill = Family)) + 
  facet_grid(.~ Fun_sor_clus2, scales = "free_x", space= "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nSite_Grass', y="Relative Abundance", fill="Family")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Family_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Family_Full_sitegrass_cluster2.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
#### Family #### 


#### Genus #### 
ggplot(dat_Fun.Genus.r, aes(x = Site_Grass,y = as.numeric(Abundance*100), 
                                                    fill = Genus)) + 
  facet_grid(.~ Fun_sor_clus2, scales = "free_x", space= "free_x")+
  scale_y_continuous(labels = scales::percent)+
  labs(x= '\nSite_Grass', y="Relative Abundance", fill="Genus")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Genus_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Genus_Full_sitegrass_cluster2.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
#### Genus #### 

#### GvS #### 
GvS_cols<-c("#d01c8b","#4dac26","grey")

ggplot(dat_Fun.Sign.r, aes(x = Site_Grass, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Sign)) + 

  facet_wrap("Fun_sor_clus2", nrow=2, ncol=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nSite_Grass', y="Relative Abundance", fill="Sign")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(
    values = c(
      "Generalists" = "#4dac26",
      "Specialists" = "#d01c8b",
      "NS" = "grey40",
      "Singleton" = "grey"),
    guide = "legend"
  ) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Sign_Full_sitegrass_cluster2.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
#### GvS #### 

```

##### 



## full dataset: Grass replicates, grouped by Site

### Bac - Phylum

```{r Bac-Phylum-Full-GrassRep-Site, message=F, warning=F, error=F, fig.dim = c(12, 9), out.width = "100%", eval=F, echo=F}
# facet site

ggplot(dat_Bac.Phylum.r, aes(x = Grass_Rep, 
                             y = as.numeric(Abundance*100), 
                             fill = Phylum)) + 

  facet_wrap("Site", nrow=4, ncol=6, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nGrass x Replicate', y="Relative Abundance", fill="Phylum")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  # scale_fill_jco() + # max 10 colors
  # scale_fill_igv() +
  scale_fill_manual(values = Bac_Phylum_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/BA_Phylum_Full_GrassRep_Site.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```

### Fun - Phylum

```{r Fun-Phylum-Full-GrassRep-Site, message=F, warning=F, error=F, fig.dim = c(12, 9), out.width = "100%", eval=F, echo=F}
# facet site

ggplot(dat_Fun.Phylum.r, aes(x = Grass_Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Phylum)) + 

  facet_wrap("Site", nrow=4, ncol=6, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nGrass x Replicate', y="Relative Abundance", fill="Phylum")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = c("≤ 0.0002% relative abundance" = '#28A48D', 
            "Ascomycota"="#8569D5",
           "Basidiomycota"="#D1A33D",
           "Mucoromycota"="#5E738F")) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Phylum_Full_GrassRep_Site.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```

##### 

### Bac - Class

```{r Bac-Class-Full-GrassRep-Site, message=F, warning=F, error=F, fig.dim = c(12, 9), out.width = "100%", eval=F, echo=F}
# facet site

ggplot(dat_Bac.Class.r, aes(x = Grass_Rep, 
                             y = as.numeric(Abundance*100), 
                             fill = Class)) + 

  facet_wrap("Site", nrow=4, ncol=6, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nGrass x Replicate', y="Relative Abundance", fill="Class")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  # scale_fill_jco() + # max 10 colors
  # scale_fill_igv() +
  scale_fill_manual(values = Bac_Class_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/BA_Class_Full_GrassRep_Site.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```

### Fun - Class

```{r Fun-Class-Full-GrassRep-Site, message=F, warning=F, error=F, fig.dim = c(12, 9), out.width = "100%", eval=F, echo=F}
# facet site

ggplot(dat_Fun.Class.r, aes(x = Grass_Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Class)) + 

  facet_wrap("Site", nrow=4, ncol=6, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nGrass x Replicate', y="Relative Abundance", fill="Class")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Class_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Class_Full_GrassRep_Site.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```

##### 

### Bac - Order

```{r Bac-Order-Full-GrassRep-Site, message=F, warning=F, error=F, fig.dim = c(12, 9), out.width = "100%", eval=F, echo=F}
# facet site

ggplot(dat_Bac.Order.r, aes(x = Grass_Rep, 
                             y = as.numeric(Abundance*100), 
                             fill = Order)) + 

  facet_wrap("Site", nrow=4, ncol=6, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nGrass x Replicate', y="Relative Abundance", fill="Order")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  # scale_fill_jco() + # max 10 colors
  # scale_fill_igv() +
  scale_fill_manual(values = Bac_Order_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/BA_Order_Full_GrassRep_Site.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```

### Fun - Order

```{r Fun-Order-Full-GrassRep-Site, message=F, warning=F, error=F, fig.dim = c(12, 9), out.width = "100%", eval=F, echo=F}
# facet site

ggplot(dat_Fun.Order.r, aes(x = Grass_Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Order)) + 

  facet_wrap("Site", nrow=4, ncol=6, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nGrass x Replicate', y="Relative Abundance", fill="Order")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Order_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Order_Full_GrassRep_Site.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```

##### 

### Bac - Family

```{r Bac-Family-Full-GrassRep-Site, message=F, warning=F, error=F, fig.dim = c(12, 9), out.width = "100%", eval=F, echo=F}
# facet site

ggplot(dat_Bac.Family.r, aes(x = Grass_Rep, 
                             y = as.numeric(Abundance*100), 
                             fill = Family)) + 

  facet_wrap("Site", nrow=4, ncol=6, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nGrass x Replicate', y="Relative Abundance", fill="Family")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  # scale_fill_jco() + # max 10 colors
  # scale_fill_igv() +
  scale_fill_manual(values = Bac_Family_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/BA_Family_Full_GrassRep_Site.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```

### Fun - Family

```{r Fun-Family-Full-GrassRep-Site, message=F, warning=F, error=F, fig.dim = c(12, 9), out.width = "100%", eval=F, echo=F}
# facet site

ggplot(dat_Fun.Family.r, aes(x = Grass_Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Family)) + 

  facet_wrap("Site", nrow=4, ncol=6, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nGrass x Replicate', y="Relative Abundance", fill="Family")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Family_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Family_Full_GrassRep_Site.png", plot = last_plot(), width=12, height=9, units = "in", device="png")
```



##### 






####################################################
####################################################
## full dataset: Site replicates, grouped by Grass

### Bac - Phylum

```{r Bac-Phylum-Full-SiteRep-Grass, message=F, warning=F, error=F, fig.dim = c(10, 10), out.width = "100%", eval=F, echo=F}
# facet grass

ggplot(dat_Bac.Phylum.r, aes(x = Site.Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Phylum)) + 
  facet_wrap("Grass", nrow=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nSite x Replicate', y="Relative Abundance", fill="Phylum")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  # scale_fill_jco() + # max 10 colors
  # scale_fill_igv() +
  scale_fill_manual(values = Bac_Phylum_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/BA_Phylum_Full_SiteRep_Grass.png", plot = last_plot(), width=10, height=10, units = "in", device="png")
```

### Fun - Phylum

```{r Fun-Phylum-Full-SiteRep-Grass, message=F, warning=F, error=F, fig.dim = c(10, 10), out.width = "100%", eval=F, echo=F}
# facet grass

ggplot(dat_Fun.Phylum.r, aes(x = Site.Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Phylum)) + 
  facet_wrap("Grass", nrow=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nSite x Replicate', y="Relative Abundance", fill="Phylum")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = c("≤ 0.0002% relative abundance" = '#28A48D', 
            "Ascomycota"="#8569D5",
           "Basidiomycota"="#D1A33D",
           "Mucoromycota"="#5E738F")) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Phylum_Full_SiteRep_Grass.png", plot = last_plot(), width=10, height=10, units = "in", device="png")
```


##### 

### Bac - Class

```{r Bac-Class-Full-SiteRep-Grass, message=F, warning=F, error=F, fig.dim = c(10, 10), out.width = "100%", eval=F, echo=F}
# facet grass

ggplot(dat_Bac.Class.r, aes(x = Site.Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Class)) + 
  facet_wrap("Grass", nrow=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nSite x Replicate', y="Relative Abundance", fill="Class")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  # scale_fill_jco() + # max 10 colors
  # scale_fill_igv() +
  scale_fill_manual(values = Bac_Class_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/BA_Class_Full_SiteRep_Grass.png", plot = last_plot(), width=10, height=10, units = "in", device="png")
```

### Fun - Class

```{r Fun-Class-Full-SiteRep-Grass, message=F, warning=F, error=F, fig.dim = c(10, 10), out.width = "100%", eval=F, echo=F}
# facet grass

ggplot(dat_Fun.Class.r, aes(x = Site.Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Class)) + 
  facet_wrap("Grass", nrow=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nSite x Replicate', y="Relative Abundance", fill="Class")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Class_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Class_Full_SiteRep_Grass.png", plot = last_plot(), width=10, height=10, units = "in", device="png")
```


##### 

### Bac - Order

```{r Bac-Order-Full-SiteRep-Grass, message=F, warning=F, error=F, fig.dim = c(10, 10), out.width = "100%", eval=F, echo=F}
# facet grass

ggplot(dat_Bac.Order.r, aes(x = Site.Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Order)) + 
  facet_wrap("Grass", nrow=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nSite x Replicate', y="Relative Abundance", fill="Order")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  # scale_fill_jco() + # max 10 colors
  # scale_fill_igv() +
  scale_fill_manual(values = Bac_Order_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/BA_Order_Full_SiteRep_Grass.png", plot = last_plot(), width=10, height=10, units = "in", device="png")
```

### Fun - Order

```{r Fun-Order-Full-SiteRep-Grass, message=F, warning=F, error=F, fig.dim = c(10, 10), out.width = "100%", eval=F, echo=F}
# facet grass

ggplot(dat_Fun.Order.r, aes(x = Site.Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Order)) + 
  facet_wrap("Grass", nrow=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nSite x Replicate', y="Relative Abundance", fill="Order")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Order_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Order_Full_SiteRep_Grass.png", plot = last_plot(), width=10, height=10, units = "in", device="png")
```




##### 

### Bac - Family

```{r Bac-Family-Full-SiteRep-Grass, message=F, warning=F, error=F, fig.dim = c(10, 10), out.width = "100%", eval=F, echo=F}
# facet grass

ggplot(dat_Bac.Family.r, aes(x = Site.Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Family)) + 
  facet_wrap("Grass", nrow=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nSite x Replicate', y="Relative Abundance", fill="Family")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  # scale_fill_jco() + # max 10 colors
  # scale_fill_igv() +
  scale_fill_manual(values = Bac_Family_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/BA_Family_Full_SiteRep_Grass.png", plot = last_plot(), width=10, height=10, units = "in", device="png")
```

### Fun - Family

```{r Fun-Family-Full-SiteRep-Grass, message=F, warning=F, error=F, fig.dim = c(10, 10), out.width = "100%", eval=F, echo=F}
# facet grass

ggplot(dat_Fun.Family.r, aes(x = Site.Rep, 
                                                    y = as.numeric(Abundance*100), 
                                                    fill = Family)) + 
  facet_wrap("Grass", nrow=5, scales = "free_x")+
  # geom_vline(xintercept = 2, color="black")+
  labs(x= '\nSite x Replicate', y="Relative Abundance", fill="Family")+
  geom_bar(stat = "identity", position="fill")+
  guides(fill=guide_legend(ncol = 1))+
  scale_fill_manual(values = Fun_Family_cols) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(color="black", size=7, angle = 90, hjust=1,vjust=.5),
    axis.text.y = element_text(color="black", size=12),
    strip.text = element_text(color="black", size=12),
    strip.background = element_blank(),
    legend.title = element_text(color="black", size=12),
    legend.text = element_text(color="black", size=10),
    legend.key.height = unit(0.5, 'cm'),
    axis.title.x = element_text(color="black", size=14),
    axis.title.y = element_text(color="black", size=14, angle=90),
    plot.title = element_text(size=14),
    panel.grid.major = element_blank(), # switch off major gridlines
    panel.grid.minor = element_blank(), # switch off minor gridlines
    panel.border = element_rect(colour = "black", fill=NA, size=.75)
  )

#tax fig naming: Dataset_Taxlevel_subset_xaxis_facet
ggsave(file="../figures/taxa_plots/Fun_Family_Full_SiteRep_Grass.png", plot = last_plot(), width=10, height=10, units = "in", device="png")
```




##### 




####################################################
####################################################
## Across Different Sites by Grass Hosts

### Bac - Phylum

autogen code:

```{r autogen-SitebyGrass-Bac-Phylum, echo=F, results='asis'}
#cut=Site
#color=Grass
#carat=sample_name
#clarity=Phylum
#price=Abundance
#col=fill

BA_tabs <- sort(unique(dat_Bac.Phylum.r2$Site))



for(tab in BA_tabs) {

  cat('\n')

  cat('### ', tab, '{.unnumbered .tabset .tabset-pills}   \n')

  cat('\n')


  sub_values <- dat_Bac.Phylum.r2 %>% filter(Site == tab)

  subtabs <- sort(unique(sub_values$Grass))


  for(subtab in subtabs) {

    cat('\n')

    cat('#### ', subtab, '{.unnumbered}   \n')

    cat('\n')

    plot_values <- sub_values %>% filter(Grass == subtab)


    print(ggplot(plot_values, aes(x = sample_name,y = as.numeric(Abundance*100), fill = Phylum)) +
          geom_bar(stat = "identity", position = "fill") +
          scale_fill_manual(values = c( "≤ 0.4% relative abundance" = '#CBD588', 
                     "Acidobacteria" = '#5F7FC7', 
                     "Actinobacteria" = '#FFA401', 
                     "Armatimonadetes" = '#DA5724',
                     "Bacteroidetes" = '#508578',
                     "Candidatus_Saccharibacteria" = '#CD9BCD',
                     "Firmicutes" = '#673770', 
                     "Planctomycetes" = '#D14285',
                     "Proteobacteria" = '#652926', 
                     "Verrucomicrobia" = '#C84248')) +
          labs(x = '\n Site x Grass x Replicate',
               y = 'Relative Abundance \n',
               fill = 'Phylum') +
          theme_bw()+
          theme(
            axis.text.x = element_text(color="black", size=9, angle = 90, hjust=1,vjust=.5),
            axis.text.y = element_text(color="black", size=12),
            strip.text = element_text(color="black", size=12),
            strip.background = element_blank(),
            legend.title = element_text(color="black", size=12),
            legend.text = element_text(color="black", size=10),
            legend.key.height = unit(0.5, 'cm'),
            axis.title.x = element_text(color="black", size=14),
            axis.title.y = element_text(color="black", size=14, angle=90),
            plot.title = element_text(size=14),
            panel.grid.major = element_blank(), # switch off major gridlines
            panel.grid.minor = element_blank(), # switch off minor gridlines
            panel.border = element_rect(colour = "black", fill=NA, linewidth=.75)
            )
          )


    cat('\n')

  }

}
```
