
CCTACGGGNGGCWGCAG
GACTACHVGGGTATCTAATCC
GATGAAGAACGYAGYRAA
TCCTCCGCTTATTGATATGC
### Predictor: Fungi - whole community, p/a of OTUs (Sørensen) {.tabset}

\

#### Response: Bacteria - abundance-weighted {.tabset .tabset-pills .unnumbered}

##### Whole community (OTUs) {.unnumbered}

-   A: Yes, but barely

```{r, include=T,eval=T}
round_mrm(MRM(Bac_bray_wholecommunity ~ Fun_sor_wholecommunity$beta.sor, nperm = 1000, method = "linear"), digits = 3)
```

\

##### Generalists (OTUs) {.unnumbered}

-   A: Yes, but barely

```{r, include=T,eval=T}
round_mrm(MRM(Bac_bray_generalists ~ Fun_sor_wholecommunity$beta.sor, nperm = 1000, method = "linear"), digits = 3)
```

\

##### Whole community (Genus-level) {.unnumbered}

OTU table agglomerated to genus-level\
-   A: Yes, but barely

```{r, include=T,eval=T}
round_mrm(MRM(Bac_bray_wholecommunity_genus ~ Fun_sor_wholecommunity$beta.sor, nperm = 1000, method = "linear"), digits = 3)
```

\

#### Response: Bacteria - p/a (Sørensen) {.tabset .tabset-pills .unnumbered}

##### Whole community (OTUs) {.unnumbered}

-   A: Slightly more

```{r, include=T,eval=T}
round_mrm(MRM(Bac_sor_wholecommunity$beta.sor ~ Fun_sor_wholecommunity$beta.sor, nperm = 1000, method = "linear"), digits = 3)
```

\

##### Generalists (OTUs) {.unnumbered}

-   A: Slightly more

```{r, include=T,eval=T}
round_mrm(MRM(Bac_sor_generalists$beta.sor ~ Fun_sor_wholecommunity$beta.sor, nperm = 1000, method = "linear"), digits = 3)
```

\

##### Whole community (Genus-level) {.unnumbered}

OTU table agglomerated to genus-level\
-   A: Yes, but barely

```{r, include=T,eval=T}
round_mrm(MRM(Bac_sor_wholecommunity_genus$beta.sor ~ Fun_sor_wholecommunity$beta.sor, nperm = 1000, method = "linear"), digits = 3)
```

\
\
