---
title: "Results: Spatial Structure in Paired Bacterial and Fungal Communities"
output:
  bookdown::html_document2:
    toc: true
    number_sections: true
bibliography: references.bib
csl: apa.csl
---

# Results

## Pure Spatial Effects {.tabset}

### Custom Functions

```{r load-functions, include=FALSE}
library(dplyr)
library(tibble)
library(geosphere)
library(vegan)
library(glue)
library(purrr)
library(tidyr)

# Function to generate decay df from dissimilarity and spatial matrices
pairwise_decay_df <- function(comm_dist, geo_dist) {
  comm_mat <- as.matrix(comm_dist)
  geo_mat <- as.matrix(geo_dist)
  stopifnot(identical(rownames(comm_mat), rownames(geo_mat)))
  idx <- which(lower.tri(comm_mat), arr.ind = TRUE)
  tibble(
    Sample1 = rownames(comm_mat)[idx[, 1]],
    Sample2 = rownames(comm_mat)[idx[, 2]],
    CommDist = comm_mat[idx],
    GeoDist = geo_mat[idx]
  )
}

# Mantel test summary with glue output
mantel_summary <- function(comm_dist, geo_dist, label) {
  m <- mantel(comm_dist, geo_dist, permutations = 999)
  glue("{label}: Mantel r = {round(m$statistic, 2)}, p = {signif(m$signif, 3)}")
}
ENDCHUNK

### Bacterial {.tabset .tabset-pills}

#### Global

```{r bac-global}
decay_bac_global <- pairwise_decay_df(comm_dist_all$Bac, dist1)
mantel_bac_global <- mantel(comm_dist_all$Bac, dist1, permutations = 999)
bac_r <- round(mantel_bac_global$statistic, 2)
bac_p <- signif(mantel_bac_global$signif, 3)
ENDCHUNK

**Figure 1A.**  
Global spatial decay in bacterial community dissimilarity.  
`r glue("Mantel r = {bac_r}, p = {bac_p}")`  
A positive spatial signal indicates geographic separation contributes to bacterial compositional differences across all samples.

#### F1 vs. non-F1

```{r bac-f1-nf1}
decay_bac_f1 <- pairwise_decay_df(comm_dist_F1$Bac, dist1)
decay_bac_nf1 <- pairwise_decay_df(comm_dist_non_F1$Bac, dist2)

mantel_bac_f1 <- mantel(comm_dist_F1$Bac, dist1, permutations = 999)
mantel_bac_nf1 <- mantel(comm_dist_non_F1$Bac, dist2, permutations = 999)

bac_f1_r <- round(mantel_bac_f1$statistic, 2)
bac_f1_p <- signif(mantel_bac_f1$signif, 3)
bac_nf1_r <- round(mantel_bac_nf1$statistic, 2)
bac_nf1_p <- signif(mantel_bac_nf1$signif, 3)
ENDCHUNK

**Figure 1B.**  
Comparison of bacterial spatial decay across groups:
- **F1** group: `r glue("Mantel r = {bac_f1_r}, p = {bac_f1_p}")`
- **non-F1** group: `r glue("Mantel r = {bac_nf1_r}, p = {bac_nf1_p}")`

Bacterial community dissimilarity increases with spatial distance in non-F1, suggesting distance–decay; F1 shows relatively weak spatial dependence.

### Fungal {.tabset .tabset-pills}

#### Global

```{r fun-global}
decay_fun_global <- pairwise_decay_df(comm_dist_all$Fun, dist1)
mantel_fun_global <- mantel(comm_dist_all$Fun, dist1, permutations = 999)
fun_r <- round(mantel_fun_global$statistic, 2)
fun_p <- signif(mantel_fun_global$signif, 3)
ENDCHUNK

**Figure 2A.**  
Global spatial decay in fungal community dissimilarity.  
`r glue("Mantel r = {fun_r}, p = {fun_p}")`  
Fungal dissimilarity does not strongly increase with distance globally, consistent with scale-invariant spatial structure across F1-dominated samples.

#### F1 vs. non-F1

```{r fun-f1-nf1}
decay_fun_f1 <- pairwise_decay_df(comm_dist_F1$Fun, dist1)
decay_fun_nf1 <- pairwise_decay_df(comm_dist_non_F1$Fun, dist2)

mantel_fun_f1 <- mantel(comm_dist_F1$Fun, dist1, permutations = 999)
mantel_fun_nf1 <- mantel(comm_dist_non_F1$Fun, dist2, permutations = 999)

fun_f1_r <- round(mantel_fun_f1$statistic, 2)
fun_f1_p <- signif(mantel_fun_f1$signif, 3)
fun_nf1_r <- round(mantel_fun_nf1$statistic, 2)
fun_nf1_p <- signif(mantel_fun_nf1$signif, 3)
ENDCHUNK

**Figure 2B.**  
Comparison of fungal spatial decay across groups:
- **F1** group: `r glue("Mantel r = {fun_f1_r}, p = {fun_f1_p}")`
- **non-F1** group: `r glue("Mantel r = {fun_nf1_r}, p = {fun_nf1_p}")`

Non-F1 fungal communities show clear spatial structuring, while F1 fungal communities exhibit minimal decay, supporting scale-invariant compositional structure.

