---
title: '**RAM paired amplicons: SFAF Poster Figures**'
author: "Kelli Feeser"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
    fig.caption: yes
    keep_md: yes
  html_notebook:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
    fig.caption: yes
  html_document:
    code_folding: hide
    css: styles.css
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: true
    fig.caption: yes
editor_options:
  chunk_output_type: inline
---

\

::: homelink
<a href="https://kellifeeser.github.io/ksu-paired-amplicon-workflow/index.html" target="_blank" style="text-align:right">Back to Home</a>
:::

\

------------------------------------------------------------------------

Document last updated: `r Sys.Date()`

------------------------------------------------------------------------

\

# Set-up {.unlisted .unnumbered .hidden}

```{r global_options, include=FALSE}
# packages
library(knitr)
# library(kableExtra)

# set up global R options
  # to list the current values of all options but better to F1
    #options()
    # easier way:
      #?options
  # to access the value of a single option
    #getOption("width")

# set up knitr global chunk options
  # See str(knitr::opts_chunk$get()) for a list of default chunk options.
#knitr::opts_chunk$set(fig.width = 8)
# knitr::opts_chunk$set(fig.path = "../figures/darksidea_plots/")
#knitr::opts_chunk$set(fig.path = "figures/taxa_plots/prefix-")

# How to use an object before it is created
  # see https://bookdown.org/yihui/rmarkdown-cookbook/load-cache.html

# changing from cache.path="../docs/code/BacterialFungalClusterInteractions_cache/html/") to cache.path="../docs/github_BacterialFungal_Cluster_Interactions_files/figure-html/")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      cache.path="../docs/github_SFAFposter_files/figure-html/")
opts_knit$set(eval.after = "fig.cap")
knitr::opts_chunk$set(fig.pos = '!h')
options(knitr.kable.NA = '')

## preload your R variables here ##
# variable_1 <- c("a", "b", "c")
####
```

```{r LANL-proxies, eval=T, include=FALSE}
# LANL proxies
Sys.setenv('HTTP_PROXY'='http://proxyout.lanl.gov:8080')
Sys.setenv('HTTPS_PROXY'='http://proxyout.lanl.gov:8080')
```

## load packages

```{r load-packages, eval=T, include=F}
library("ggplot2"); packageVersion("ggplot2") #‘3.4.4’
library("vegan"); packageVersion("vegan") # # ‘2.6.4’
library("phyloseq"); packageVersion("phyloseq") #‘1.44.0’
library("scales"); packageVersion("scales") #‘1.2.1’
library("swatches"); packageVersion("swatches") #‘0.5.0’
library("ggsci"); packageVersion("ggsci") #‘3.0.0’
library("ggpubr"); packageVersion("ggpubr") #‘0.6.0’
library("viridis"); packageVersion("viridis") #‘0.6.4’
library("RColorBrewer"); packageVersion("RColorBrewer") #‘1.1.3’
library("cowplot"); packageVersion("cowplot") #‘1.1.1’
library("dplyr"); packageVersion("dplyr") #‘1.1.3’
library("plyr"); packageVersion("plyr") #‘1.8.9’
library("knitr"); packageVersion("knitr") #‘1.44’
#library("kableExtra"); packageVersion("kableExtra") #‘1.3.4’
library("DT"); packageVersion("DT") #‘0.30’
library("tidyverse"); packageVersion("tidyverse") #‘2.0.0’
library("prettydoc"); packageVersion("prettydoc") #‘0.4.1’
library("magrittr"); packageVersion("magrittr") #‘2.0.3’
library("broom"); packageVersion("broom") #‘1.0.5’
library("betapart"); packageVersion("betapart") #‘1.6’
# patch see: https://stackoverflow.com/questions/76118194/error-when-loading-kableextra-in-markdown-file-after-updating-to-r-4-3-0
# devtools::install_github("kupietz/kableExtra",force = TRUE)
library("kableExtra")

library("microbiome"); packageVersion("microbiome") #‘1.22.0’
library("supernova"); packageVersion("supernova") #‘3.0.0’
library("lme4"); packageVersion("lme4") #‘1.1.34’
library("circlize"); packageVersion("circlize") #‘0.4.16’
library("ggordiplots"); packageVersion("ggordiplots") #‘0.4.3’


# source("ksu_colors.R")
source("RAM_bulk_harvester.R")
source("../../black-sheep-club/code/sheepish_scripts.R")
```

## load cache

```{r load-cache, include=FALSE, eval=F}
# OBJECT <- if (file.exists('OBJECT.rds')) {
#   readRDS('OBJECT.rds')
# } else {
#   "The value of `OBJECT` is not available yet"
# }

# knitr::load_cache('CHUNK-NAME', 'OBJECT')



# knitr::load_cache('adonis-predClus-Fun-calc', 'adonis_predClus_Fun')
# knitr::load_cache('adonis-predClus-Bac-calc', 'adonis_predClus_Bac')
# knitr::load_cache('calc-alpha', 'fwc_alpha')
# knitr::load_cache('calc-alpha', 'bwc_alpha')
# knitr::load_cache('gamma-fun-calc', 'Fun_gamma2')
# knitr::load_cache('gamma-bac-calc', 'Bac_gamma2')

# knitr::load_cache('mean-diss-bycluster-Fun', 'md_Fwc_self')
# knitr::load_cache('mean-diss-bycluster-Fun', 'md_Fwc_clus_sor_k10.df.noself')
# knitr::load_cache('mean-diss-bycluster-Bac', 'md_Bwc_self')
# knitr::load_cache('mean-diss-bycluster-Bac', 'md_Bwc_clus_sor_k14.df.noself')

# knitr::load_cache('subset-Fun-wc-byClus', 'sor.dist.list')
# knitr::load_cache('subset-Bac-wc-byClus')

# knitr::load_cache('calc-betadisp')
# knitr::load_cache('betadisp-boxplot-Fun', 'betadisp_clus_Fun.df')
# knitr::load_cache('', '')
# knitr::load_cache('', '')
# knitr::load_cache('', '')

```

## load rds

```{r load-rds, include=FALSE,eval=TRUE}
##########
# phylo #
Bac_wholecommunity<-readRDS(file = "../processed_data/clean_rds_saves/Bac_wholecommunity.rds")
Fun_wholecommunity <- readRDS(file="../processed_data/clean_rds_saves/Fun_wholecommunity.rds")
##########

##########
# nmds #
Bac_wc_sor_k5.ord <- readRDS(file = "../processed_data/clean_rds_saves/Bac_wc_sor_k5.ord.rds")
Fun_wc_sor_k4.ord <- readRDS(file = "../processed_data/clean_rds_saves/Fun_wc_sor_k4.ord.rds")
##########


##########
# misc #
Bac_wholecommunity.r<-readRDS(file = "../processed_data/clean_rds_saves/Bac_wholecommunity.r.rds")
Fun_wholecommunity.r<-readRDS(file = "../processed_data/clean_rds_saves/Fun_wholecommunity.r.rds")
sd<-as(sample_data(Fun_wholecommunity), "data.frame") # metadata for phyloseq
##########




##########
# colors #
col_n14 <- c("B1"="#730025","B2"="#AF0039","B3"="#E8004C","B4"="#FF65A3","B5"="#9E4D97",
             "B6"="#4D42B3","B7"="#7678ed","B8"="#5B7898","B9"="#17A398","B10"="#74A57F",
             "B11"="#B6AF40","B12"="#f7b801", "B13"="#f18701", "B14"="#f35b04")
# creation: https://coolors.co/730025-AF0039-E8004C-FF65A3-9E4D97-4D42B3-7678ed-5B7898-17A398-74A57F-B6AF40-f7b801-f18701-f35b04


col_n10 <- c("F1"="#8E0152","F2"="#C51B7D","F3"="#DE77AE","F4"="#F1B6DA","F5"="#C5C2EA",
             "F6"="#8DC9BB","F7"="#86C37E","F8"="#7FBC41","F9"="#4D9221","F10"="#276419")
             
                 # "#DBBCE2","F6"="#E6F5D0","F7"="#B8E186","F8"="#7FBC41","F9"="#4D9221","F10"="#276419")

             # "#B77EBE","#B0AFDF","#92CDC8","#7BB990","#EFBC9B","#DBA8AC"

# creation: https://coolors.co/8e0152-c51b7d-de77ae-f1b6da-c5c2ea-a8e0ff-8dc9bb-7fbc41-4d9221-276419
col_clusn10 <- c("F1"="#8E0152","F2"="#C51B7D","F3"="#DE77AE","F4"="#F1B6DA","F5"="#FDE0EF",
                 "F6"="#E6F5D0","F7"="#B8E186","F8"="#7FBC41","F9"="#4D9221","F10"="#276419")

# change "F5"="#FDE0EF" to "#C5C2EA",
  # "F6"="#E6F5D0" to "#8DC9BB", 
  # "F7"="#B8E186" to "#86C37E"

latlong_cols<-readRDS(file = "../processed_data/clean_rds_saves/latlong_cols.rds")
##########
```

## check metadata

```{r sampledata-check, include=FALSE,eval=FALSE}
# Bin should be north
unique(subset(sample_data(Fun_wholecommunity), Site=="ONF")$Bin)
unique(subset(sample_data(Bac_wholecommunity), Site=="ONF")$Bin)

tail(colnames(sample_data(Fun_wholecommunity)))
tail(colnames(sample_data(Bac_wholecommunity)))

# colnames
colnames(sample_data(Fun_wholecommunity))

```

\

# nestedness


```{r}
MRM(betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F10_Fun"]]$beta.sor ~ env_dist_matrices[["Fun_sor_clusters"]][["F10_pH_m.std"]] + env_dist_matrices[["Fun_sor_clusters"]][["F10_soil_moisture_m.std"]] + env_dist_matrices[["Fun_sor_clusters"]][["F10_SOM_m.std"]] + env_dist_matrices[["Fun_sor_clusters"]][["F10_ammonium_m.std"]] + env_dist_matrices[["Fun_sor_clusters"]][["F10_phos_m.std"]] + env_dist_matrices[["Fun_sor_clusters"]][["F10_GDD3yr_m.std"]] + env_dist_matrices[["Fun_sor_clusters"]][["F10_ppt3yr_m.std"]], nperm=1001)

decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F10"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F10_pH_m.std"]] + env_dist_matrices[["Fun_sor_clusters"]][["F10_soil_moisture_m.std"]] +
              env_dist_matrices[["Fun_sor_clusters"]][["F10_SOM_m.std"]], y.type="dissim", model.type="exp")
```


## F1 vs non-F1


### sor

```{r}
# F1
decay.Fun_F1.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["F1"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clus2"]][["F1_geodist"]], y.type="dissim", model.type="exp")
# non-F1
decay.Fun_nonF1.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["non-F1"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clus2"]][["non-F1_geodist"]], y.type="dissim", model.type="exp")

plot.decay.pretty(decay.Fun_F1.exp.sor, col=paste(col_n10[1],"cc",sep = ""), add=F, main="F1", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[1],xaxp = c(0,1500,3),
           ylab="",xlab="")

plot.decay.pretty(decay.Fun_nonF1.exp.sor, col="#7FBC41", add=T, main="non-F1",
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line="#276419",xaxp = c(0,1500,3),
           ylab="",xlab="")

```

### sim

```{r}
# F1
decay.Fun_F1.exp.sim<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["F1"]]$beta.sim, x=env_dist_matrices[["Fun_sor_clus2"]][["F1_geodist"]], y.type="dissim", model.type="exp")
# non-F1
decay.Fun_nonF1.exp.sim<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["non-F1"]]$beta.sim, x=env_dist_matrices[["Fun_sor_clus2"]][["non-F1_geodist"]], y.type="dissim", model.type="exp")

plot.decay.pretty(decay.Fun_F1.exp.sim, col=paste(col_n10[1],"cc",sep = ""), add=F, main="Turnover-resultant dissimilarity", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[1],xaxp = c(0,1500,3),
           ylab="",xlab="")

plot.decay.pretty(decay.Fun_nonF1.exp.sim, col="#7FBC41", add=T, main="non-F1",
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line="#276419",xaxp = c(0,1500,3),
           ylab="",xlab="")
mtext((paste("F1 - ",round(decay.Fun_F1.exp.sim$pseudo.r.squared,3)," (",round(decay.Fun_nonF1.exp.sim$p.value,3),")",sep = "")),side=1,line=-2,adj=0.05,outer = F,padj = 0,col=col_n10[1])
mtext((paste("non-F1 - ",round(decay.Fun_nonF1.exp.sim$pseudo.r.squared,3)," (",round(decay.Fun_nonF1.exp.sim$p.value,3),")",sep = "")),side=1,line=-1,adj=0.05,outer = F,padj = 0, col="#7FBC41")
mtext((paste("pseudo R2 (p-value)",sep = "")),side=1,line=-3,adj=0.05,outer = F,padj = 0, col="black")

```

### sne

```{r}
# F1
decay.Fun_F1.exp.sne<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["F1"]]$beta.sne, x=env_dist_matrices[["Fun_sor_clus2"]][["F1_geodist"]], y.type="dissim", model.type="exp")
# non-F1
decay.Fun_nonF1.exp.sne<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clus2"]][["pairwise"]][["non-F1"]]$beta.sne, x=env_dist_matrices[["Fun_sor_clus2"]][["non-F1_geodist"]], y.type="dissim", model.type="exp")

plot.decay.pretty(decay.Fun_F1.exp.sne, col=paste(col_n10[1],"cc",sep = ""), add=F, main="Nestedness-resultant dissimilarity", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[1],xaxp = c(0,1500,3),
           ylab="",xlab="")

plot.decay.pretty(decay.Fun_nonF1.exp.sne, col="#7FBC41", add=T, main="non-F1",
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line="#276419",xaxp = c(0,1500,3),
           ylab="",xlab="")
mtext((paste("F1 - ",round(decay.Fun_F1.exp.sne$pseudo.r.squared,3)," (",round(decay.Fun_nonF1.exp.sne$p.value,3),")",sep = "")),side=3,line=-3,adj=0.05,outer = F,padj = 0,col=col_n10[1])
mtext((paste("non-F1 - ",round(decay.Fun_nonF1.exp.sne$pseudo.r.squared,3)," (",round(decay.Fun_nonF1.exp.sne$p.value,3),")",sep = "")),side=3,line=-2,adj=0.05,outer = F,padj = 0, col="#7FBC41")
mtext((paste("pseudo R2 (p-value)",sep = "")),side=3,line=-1,adj=0.05,outer = F,padj = 0, col="black")

```

## Fun_sor_clusters

```{r}
betapart_results <- readRDS(file = "../processed_data/clean_rds_saves/betapart_results.rds")

env_dist_matrices <- readRDS(file = "../processed_data/clean_rds_saves/env_dist_matrices.rds")

env_dist_matrices <- readRDS(file = "../processed_data/clean_rds_saves/env_dist_matrices2.rds")
```



```{r}
# col_n10 <- c("F1"="#8E0152","F2"="#C51B7D","F3"="#DE77AE","F4"="#F1B6DA","F5"="#C5C2EA",
#              "F6"="#8DC9BB","F7"="#86C37E","F8"="#7FBC41","F9"="#4D9221","F10"="#276419")

col_n10[1]
paste(col_n10[1],"cc",sep = "")


F1sor_dist <- betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F1_Fun"]]$beta.sor
F1geo_dist <- env_dist_matrices$Fun_sor_clusters$F1_geodist
as.data.frame(F1geo_dist, rownames = names(F1sor_dist), rownames.force = T)

Fun_all.geo<-as(subset(sd, select = c(Longitude,Latitude)), "data.frame")
  Fun_all.geodist_mat <- as.dist(geodist(Fun_all.geo, measure = 'geodesic' )/1000)
Fun_F1.geo<-as(subset(subset(sd, sd$Fun_sor_clusters=="F1"), select = c(Longitude,Latitude)), "data.frame")
  Fun_F1.geodist_mat <- stats::as.dist(geodist(Fun_F1.geo, measure = 'geodesic' )/1000)




MRM(dist(F1sor_dist)~dist(Fun_F1.geodist_mat))
# F1
decay.Fun_F1.exp.sor2<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F1_Fun"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]]$F1_geodist, y.type="dissim", model.type="exp")
# F2 - singular gradient matrix at initial parameter estimates
decay.Fun_F2.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F2"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F2_geodist"]], y.type="dissim", model.type="exp")
# F3
decay.Fun_F3.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F3"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F3_geodist"]], y.type="dissim", model.type="exp")
# F4
decay.Fun_F4.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F4"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F4_geodist"]], y.type="dissim", model.type="exp")
# F5
decay.Fun_F5.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F5"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F5_geodist"]], y.type="dissim", model.type="exp")
# F6
decay.Fun_F6.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F6"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F6_geodist"]], y.type="dissim", model.type="exp")
# F7
decay.Fun_F7.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F7"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F7_geodist"]], y.type="dissim", model.type="exp")
# F8
decay.Fun_F8.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F8"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F8_geodist"]], y.type="dissim", model.type="exp")
# F9
decay.Fun_F9.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F9"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F9_geodist"]], y.type="dissim", model.type="exp")
# F10
decay.Fun_F10.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["F10"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["F10_geodist"]], y.type="dissim", model.type="exp")

# non-F1
decay.Fun_nonF1.exp.sor<-decay.model(y=betapart_results[["Fun"]][["Fun_sor_clusters"]][["pairwise"]][["non-F1"]]$beta.sor, x=env_dist_matrices[["Fun_sor_clusters"]][["non-F1_geodist"]], y.type="dissim", model.type="exp")

plot.decay.pretty(decay.Fun_F1.exp.sor2, col=paste(col_n10[1],"cc",sep = ""), add=F, main="F1", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[1],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Fun_F3.exp.sor, col=paste(col_n10[3],"cc",sep = ""), add=T, main="F3", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[3],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Fun_F4.exp.sor, col=paste(col_n10[4],"cc",sep = ""), add=T, main="F4", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[4],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Fun_F5.exp.sor, col=paste(col_n10[5],"cc",sep = ""), add=T, main="F5", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[5],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Fun_F6.exp.sor, col=paste(col_n10[6],"cc",sep = ""), add=T, main="F6", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[6],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Fun_F7.exp.sor, col=paste(col_n10[7],"cc",sep = ""), add=T, main="F7", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[7],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Fun_F8.exp.sor, col=paste(col_n10[8],"cc",sep = ""), add=T, main="F8", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[8],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Fun_F9.exp.sor, col=paste(col_n10[9],"cc",sep = ""), add=T, main="F9", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[9],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Fun_F10.exp.sor, col=paste(col_n10[10],"cc",sep = ""), add=T, main="F10", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n10[10],xaxp = c(0,1500,3),
           ylab="",xlab="")

plot.decay.pretty(decay.Fun_nonF1.exp.sor, col="#7FBC41", add=T, main="non-F1",
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line="#276419",xaxp = c(0,1500,3),
           ylab="",xlab="")

```


## Bac_sor_clusters

```{r}
col_n10 <- c("B1"="#8E0152","B2"="#C51B7D","B3"="#DE77AE","B4"="#B1B6DA","B5"="#C5C2EA",
             "B6"="#8DC9BB","B7"="#86C37E","B8"="#7BBC41","B9"="#4D9221","B10"="#276419")

col_n10[1]
paste(col_n10[1],"cc",sep = "")

# B1
decay.Bac_B1.exp.sor2<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B1"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B1_geodist"]], y.type="dissim", model.type="exp")
# B2 
decay.Bac_B2.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B2"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B2_geodist"]], y.type="dissim", model.type="exp")
# B3
decay.Bac_B3.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B3"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B3_geodist"]], y.type="dissim", model.type="exp")
# B4
decay.Bac_B4.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B4"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B4_geodist"]], y.type="dissim", model.type="exp")
# B5
decay.Bac_B5.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B5"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B5_geodist"]], y.type="dissim", model.type="exp")
# B6 - singular gradient matrix at initial parameter estimates
decay.Bac_B6.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B6"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B6_geodist"]], y.type="dissim", model.type="exp")
# B7
decay.Bac_B7.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B7"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B7_geodist"]], y.type="dissim", model.type="exp")
# B8
decay.Bac_B8.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B8"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B8_geodist"]], y.type="dissim", model.type="exp")
# B9
decay.Bac_B9.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B9"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B9_geodist"]], y.type="dissim", model.type="exp")
# B10
decay.Bac_B10.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B10"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B10_geodist"]], y.type="dissim", model.type="exp")
# B11
decay.Bac_B11.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B11"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B11_geodist"]], y.type="dissim", model.type="exp")
# B12
decay.Bac_B12.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B12"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B12_geodist"]], y.type="dissim", model.type="exp")
# B13
decay.Bac_B13.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B13"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B13_geodist"]], y.type="dissim", model.type="exp")
# B14 - singular gradient matrix at initial parameter estimates
decay.Bac_B14.exp.sor<-decay.model(y=betapart_results[["Bac"]][["Bac_sor_clusters"]][["pairwise"]][["B14"]]$beta.sor, x=env_dist_matrices[["Bac_sor_clusters"]][["B14_geodist"]], y.type="dissim", model.type="exp")


plot.decay.pretty(decay.Bac_B1.exp.sor2, col=paste(col_n14[1],"cc",sep = ""), main="B1", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[1],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B2.exp.sor, col=paste(col_n14[2],"cc",sep = ""), add=T, main="B2", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[2],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B3.exp.sor, col=paste(col_n14[3],"cc",sep = ""), add=T, main="B3", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[3],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B4.exp.sor, col=paste(col_n14[4],"cc",sep = ""), add=T, main="B4", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[4],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B5.exp.sor, col=paste(col_n14[5],"cc",sep = ""), add=T, main="B5", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[5],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B7.exp.sor, col=paste(col_n14[7],"cc",sep = ""), add=T, main="B7", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[7],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B8.exp.sor, col=paste(col_n14[8],"cc",sep = ""), add=T, main="B8", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[8],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B9.exp.sor, col=paste(col_n14[9],"cc",sep = ""), add=T, main="B9", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[9],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B10.exp.sor, col=paste(col_n14[10],"cc",sep = ""), add=T, main="B10", 
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[10],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B11.exp.sor, col=paste(col_n14[11],"cc",sep = ""), add=T,
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[11],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B12.exp.sor, col=paste(col_n14[12],"cc",sep = ""), add=T,
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[12],xaxp = c(0,1500,3),
           ylab="",xlab="")
plot.decay.pretty(decay.Bac_B13.exp.sor, col=paste(col_n14[13],"cc",sep = ""), add=T,
           xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
           col.line=col_n14[13],xaxp = c(0,1500,3),
           ylab="",xlab="")
# plot.decay.pretty(decay.Bac_B14.exp.sor, col=paste(col_n10[14],"cc",sep = ""), add=T,
#            xlim = c(0,1600),ylim = c(0,1),las=1, lty=1,lwd=5,
#            col.line=col_n10[14],xaxp = c(0,1500,3),
#            ylab="",xlab="")


```



# nmds

(ref:combined-NMDSspider-Sor-PAM-all-fig-caption) NMDS plots showing variation in the presence/absence-based composition of root-associated A) bacteria and B) fungi in relation to PAM cluster designations (labeled) with ellipses showing 90% CI for the standard errors about each cluster centroid. Fungal NMDS plot based on Sørensen dissimilarities: k = 4, 2-dimensional stress = 0.122, non-metric fit R^2^ = 0.985, linear fit R^2^ = 0.908. Bacterial NMDS plot based on Sørensen dissimilarities: k = 5, 2-dimensional stress = 0.055, non-metric fit R^2^ = 0.997, linear fit R^2^ = 0.981.



```{r, echo=FALSE,eval=F, out.height="90%",fig.cap='(ref:combined-NMDSspider-Sor-PAM-all-fig-caption)', fig.show='hold', dpi=300}
library("ggordiplots")

# mean-diss-bycluster-Fun
library("betapart")
Fun_wc_sorensen_dist<-as.dist(as.matrix(beta.pair(decostand(phylo2vegan_OTU(Fun_wholecommunity), method = "pa"), index.family = "sorensen")$beta.sor))

md_Fwc_clus_sor_k10<-meandist(Fun_wc_sorensen_dist, phylo2vegan_sd(Fun_wholecommunity)$Fun_sor_clusters)

md_Fwc_clus_sor_k10.r <- round(md_Fwc_clus_sor_k10,digits = 3)

# Diagonal elements = mean within-group dissimilarity
  # extract the diagonals directly from matrix, preserving names in 
  # vectorized output
  md_Fwc_clus_sor_k10.df.self <- diag(md_Fwc_clus_sor_k10.r)

# vector reflects inverted beta dispersion scaled to max of 1
  (1/md_Fwc_clus_sor_k10.df.self)/max(1/md_Fwc_clus_sor_k10.df.self)
  
  alpha_scale <- 1
  
  Fun_alpha_vec <- data.frame("Group" = names(md_Fwc_clus_sor_k10.df.self),
                             "alpha_val" = as.numeric((1/md_Fwc_clus_sor_k10.df.self)*alpha_scale/max(1/md_Fwc_clus_sor_k10.df.self)))
  
library("forcats")
# forcats::fct_rev()
Fun_ordiplot_spiders1$df_mean.ord2 <- Fun_ordiplot_spiders1$df_mean.ord

Fun_ordiplot_spiders1$df_mean.ord2$Group
Fun_ordiplot_spiders1$df_mean.ord2$Group <- forcats::fct_rev(Fun_ordiplot_spiders1$df_mean.ord2$Group)
Fun_ordiplot_spiders1$df_mean.ord2$x <- rev(Fun_ordiplot_spiders1$df_mean.ord2$x)
Fun_ordiplot_spiders1$df_mean.ord2$y <- rev(Fun_ordiplot_spiders1$df_mean.ord2$y)

  # Fungal ----

Fun_ordiplot_spiders1<-gg_ordiplot(Fun_wc_sor_k4.ord, groups = sd$Fun_sor_clusters, 
            pt.size = 1,kind = "se",conf = 0.90,ellipse = TRUE,
            label = T, spiders = T, hull = F, plot = F)
Fun_ordiplot_spiders <- Fun_ordiplot_spiders1$plot

# Join density alpha values to hull data
Fun_hull_df <- Fun_ordiplot_spiders1$df_hull %>%
  left_join(Fun_alpha_vec, by = "Group")


Fun_ordiplot_spiders2<-Fun_ordiplot_spiders+
  scale_color_manual(values = col_n10, name = "Sørensen clusters", 
                     guide = guide_legend(ncol = 1)) +
  scale_fill_manual(values = col_n10, name = "Sørensen clusters", 
                    guide = guide_legend(ncol = 1)) +
  geom_polygon(
    data = Fun_hull_df,
    aes(x = x, y = y, group = Group, fill = Group, alpha = alpha_val*1.2),
    color = NA
  ) +
  guides(alpha = "none") + # hides the alpha legend
  geom_label(data = Fun_ordiplot_spiders1$df_mean.ord, 
             aes(x = x, y = y, label = Group, color = Group), 
             show.legend = FALSE, size=4.72, label.size = 1.4, 
               label.padding = unit(0.12, "lines"), label.r = unit(0.1, "lines"),
             fontface="bold")+
  ggtitle("Fungal Sub-Assemblages") +
  theme_bw(base_family = "Arial") +
  theme(legend.position="none",
        aspect.ratio = 0.7488341,
        plot.title = element_text(color="#E58908", size=28, face="bold"),
        axis.text = element_text(color="black", size=18),
        axis.title = element_text(color="black", size=20),
        plot.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1.5), #fill="#E5A60C1a"
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) 

# aspect.ratio = (ggplot_build(Fun_ordiplot_spiders2)$layout$panel_scales_y[[1]]$range$range[1]-ggplot_build(Fun_ordiplot_spiders2)$layout$panel_scales_y[[1]]$range$range[2])/(ggplot_build(Fun_ordiplot_spiders2)$layout$panel_scales_x[[1]]$range$range[1]-ggplot_build(Fun_ordiplot_spiders2)$layout$panel_scales_x[[1]]$range$range[2]), # 0.7488341

Fun_ordiplot_spiders2
# 6x8"
# 4.5x6"
ggsave(filename="../figures/sfaf/sfaf2025/nmds_Fun_all.png", plot = Fun_ordiplot_spiders2,height = 6, width=8,units = "in", dpi = 300, device = "png")
```


```{r, echo=FALSE,eval=F, out.height="90%",fig.cap='(ref:combined-NMDSspider-Sor-PAM-all-fig-caption)', fig.show='hold', dpi=300}
# Bacterial ----
Bac_ordiplot_spiders1<-gg_ordiplot(Bac_wc_sor_k5.ord, groups = sd$Bac_sor_clusters, 
            pt.size = 1,kind = "se",conf = 0.90,ellipse = TRUE,
            label = T, spiders = T, hull = F, plot = F)
Bac_ordiplot_spiders <- Bac_ordiplot_spiders1$plot

Bac_ordiplot_spiders2<-Bac_ordiplot_spiders+
  scale_color_manual(values = col_n14, name = "Sørensen clusters",guide = guide_legend(ncol = 2))+
  scale_fill_manual(values = col_n14, name = "Sørensen clusters",guide = guide_legend(ncol = 2))+
  geom_label(data = Bac_ordiplot_spiders1$df_mean.ord, aes(x = x, y = y, label = Group, color = Group), show.legend = FALSE, size=5, label.size = 2, fontface="bold")+
  # geom_point(size = 0.8, alpha = 0.2) +
  ggtitle("Bacterial Sub-Assemblages") +
          # , subtitle = "ellipses showing 90% CI for the standard errors about each cluster centroid") +
  theme_bw() +
  theme(legend.position="none",
        aspect.ratio = 0.8382428,
        plot.title = element_text(color="#006e90", size=28, face="bold"),
        plot.subtitle = element_text(color="black", size=16),
        axis.text = element_text(color="black", size=20),
        axis.title = element_text(color="black", size=24),
        legend.text = element_text(color="black", size=8),legend.title = element_text(color="black", size=10),
        plot.background = element_blank(),panel.border = element_rect(colour = "black", fill=NA, linewidth=1.5), #fill="#006e901a"
        panel.grid.major = element_blank(),panel.grid.minor = element_blank(),panel.background = element_blank())

ggplot_build(Fun_ordiplot_spiders2)$layout$panel_scales_x[[1]]$range$range


#  
ggsave(filename="../figures/sfaf/nmds_Bac_all.png", plot = last_plot(),height = 8, width=9.543774, units = "in", dpi = 300, device = "png")
```


```{r}

Fun_ordiplot_spidersF1a<-gg_ordiplot(Fun_wc_sor_k4.ord, groups = sd$Fun_sor_clus2, 
            pt.size = 1,kind = "se",conf = 0.90,ellipse = TRUE,show.groups = "all",
            label = T, spiders = T, hull = T, plot = F)
Fun_ordiplot_spidersF1 <- Fun_ordiplot_spidersF1a$plot 

Fun_ordiplot_spidersF1+
  scale_color_manual(values = c("#C51B7D","#7FBC41e6"), name = "Sørensen clusters",guide = guide_legend(ncol = 2))+
  scale_fill_manual(values = c("#C51B7D","#7FBC41e6"), name = "Sørensen clusters",guide = guide_legend(ncol = 2))+
  geom_label(data = Fun_ordiplot_spidersF1a$df_mean.ord, aes(x = x, y = y, label = Group, color = Group), show.legend = FALSE, size=3.5, label.size = 1.25, fontface="bold",
             label.padding = unit(0.2, "lines"), label.r = unit(0.3, "lines"))+
  # geom_point(size = 0.8, alpha = 0.2) +
  ggtitle("Fungal Sub-Assemblages") +
  theme_bw(base_family = "Arial") +
  theme(legend.position="none",
        aspect.ratio = 0.7488341,
        plot.title = element_text(color="#E58908", size=14, face="bold"),
        axis.text = element_text(color="black", size=9),
        axis.title = element_text(color="black", size=10),
        plot.background = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=1.5), #fill="#E5A60C1a"
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) 

ggsave(filename="../figures/sfaf/sfaf2025/nmds_Fun_F1vnonF1.png", plot = last_plot(),height = 3, width=4,units = "in", dpi = 300, device = "png")

```

# alpha div

```{r calc-alpha, include=FALSE,eval=TRUE}
fwc_alpha <- data.frame(sample_data(Fun_wholecommunity),
                        estimate_richness(Fun_wholecommunity))


# renaming for consistency
  # simpson_inv 
  # fwc_alpha <- subset(fwc_alpha, select = -c(InvSimpson))
  # names(fwc_alpha)[names(fwc_alpha) == "Simpson"] <- "simpson_inv"
  
  # inv_simpson
  names(fwc_alpha)[names(fwc_alpha) == "InvSimpson"] <- "simpson_inv"
  
  # simpson
  names(fwc_alpha)[names(fwc_alpha) == "Simpson"] <- "simpson"
  
  # shannon
  names(fwc_alpha)[names(fwc_alpha) == "Shannon"] <- "shannonH"


# calculations needed
  # simpsonD
  fwc_alpha$simpsonD <- 1-fwc_alpha$simpson
  
  # simpsoneven
  fwc_alpha$simpsoneven <- 1/(fwc_alpha$simpsonD*fwc_alpha$Observed)
  
  # shannoneven
  fwc_alpha$shannoneven <- fwc_alpha$shannonH/log(fwc_alpha$Observed)

  # fwc_alpha$simpson<- 1-(fwc_alpha$simpsonD)

  
  bwc_alpha <- data.frame(sample_data(Bac_wholecommunity),
                        estimate_richness(Bac_wholecommunity))
  # inv_simpson
  names(bwc_alpha)[names(bwc_alpha) == "InvSimpson"] <- "simpson_inv"
  # simpson
  names(bwc_alpha)[names(bwc_alpha) == "Simpson"] <- "simpson"
  # shannon
  names(bwc_alpha)[names(bwc_alpha) == "Shannon"] <- "shannonH"
  # simpsonD
  bwc_alpha$simpsonD <- 1-bwc_alpha$simpson
  # simpsoneven
  bwc_alpha$simpsoneven <- 1/(bwc_alpha$simpsonD*bwc_alpha$Observed)
  # shannoneven
  bwc_alpha$shannoneven <- bwc_alpha$shannonH/log(bwc_alpha$Observed)

```

```{r}
# test ----
f <- data.frame(
  category = factor(rep(c("A", "B", "C"), each = 5)),
  value = runif(15)
)

# Create a dummy variable to expand category "A"
df$expand_width <- ifelse(df$category == "A", 2, 1) # Expand "A" by a factor of 2

# Create the plot
ggplot(df, aes(x = category, y = value, fill = category)) +
  geom_bar(stat = "identity", position = "dodge", width = df$expand_width) +
  scale_fill_manual(values = c("A" = "red", "B" = "blue", "C" = "green")) +
  labs(title = "Expanded Bar for Category A") +
  theme_minimal()
# ENDtest ----

```


```{r plot-Fun-alpha, echo=FALSE, fig.height=3.5, fig.width=5, dpi=300, out.height="70%",out.width="70%", eval=FALSE}
fwc_alpha2 <- fwc_alpha
levels(fwc_alpha2$Fun_sor_clusters) <- c('',levels(fwc_alpha2$Fun_sor_clusters),'')


ggviolin(fwc_alpha, x = "Fun_sor_clusters", y = "Observed",
          fill = "Fun_sor_clusters", palette = col_n10,
         alpha = 0.85, trim = FALSE, add = c("dotplot",
                                             "boxplot"
                                           ,"mean_sd","mean"
                                           ),#size=1,
         add.params = list(alpha = 0.6, binwidth = 9.25, size=0.8),
         error.plot = "errorbar" ) +
  # stat_compare_means(aes(label = after_stat(p.signif)),
  #                    method = "wilcox.test", ref.group = "F1", # Add pairwise p-value
  #                    label = "p.signif",hide.ns=T, paired = F, size=3
  #                    ,label.y = 300)+ 
  # stat_compare_means(label.y = 7,label.x = 1.15)+     # Add global p-value
  geom_hline(yintercept = mean(fwc_alpha$Observed), linetype = 2)+ # add mean hline
  scale_y_continuous(limits = c(0,350))+
  theme(legend.position = "none",
        axis.text.y = element_text(size=18, color = "black"),
        axis.text.x = element_text(size=20, color = col_n10, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=20, color = "black"),
        plot.subtitle = element_text(size=22, color = "black"),
        plot.title = element_text(color="#E58908", size=28, face="bold"))+
  labs(x = "Fungal Sub-Assemblage",title = "Fungal Species Richness",
       y = "Observed OTUs",
       caption = "pairwise wilcox.test, ref.group = Sub-Assemblage F1",
       subtitle = paste("F1 (mean = ",
round(mean(fwc_alpha$Observed[fwc_alpha$Fun_sor_clusters=="F1"]),0),
                        ") vs. all (global mean = ", 
                        round(mean(fwc_alpha$Observed),0),
                        ")",sep = ""))
  # scale_x_discrete(breaks = c(1:10))
  # expand_limits(x= c(-1, 0))+
  # scale_x_discrete(drop=FALSE)
# scale_x_discrete(limits = "F1")
# scale_x_discrete(limits = c("F2","F2","F3","F4","F5",
#                             "F6","F7","F8","F9","F10"))

# ggsave(filename="../figures/sfaf/sfaf2025/obs_Fun_clusters.png", plot = last_plot(),height = 4.8, width=7.5,units = "in", dpi = 400, device = "png")

# ggsave(filename="../figures/sfaf/sfaf2025/obs_Fun_clustersF1.png", plot = last_plot(),height = 4.8, 
#        width=7.95*(183/484)*1.05,
#        units = "in", dpi = 400, device = "png")

ggsave(filename="../figures/sfaf/sfaf2025/obs_Fun_clusters_binwidth925_all.png", plot = last_plot(),height = 4.8, 
       width=((7.5*1.05)+3.156183),
       units = "in", dpi = 400, device = "png")
length((fwc_alpha$SampleID[fwc_alpha$Fun_sor_clusters=="F1"]))
```


```{r plot-Fun-alpha-F1, echo=FALSE, fig.height=3.5, fig.width=5, dpi=300, out.height="70%",out.width="70%", eval=FALSE}
ggviolin(subset(fwc_alpha, fwc_alpha$Fun_sor_clus2=="F1"), x = "Fun_sor_clusters", y = "Observed",
          fill = "Fun_sor_clusters", palette = col_n10[1],
         alpha = 0.85, trim = FALSE, add = c("dotplot",
                                             "boxplot"
                                           ,"mean_sd","mean"
                                           ),#size=1,
         add.params = list(alpha = 0.6, binwidth = 9.25, size=0.8),
         error.plot = "errorbar" ) +
  geom_hline(yintercept = mean(fwc_alpha$Observed), linetype = 2)+ # add mean hline
  scale_y_continuous(limits = c(0,350))+
  theme(legend.position = "none",
        axis.text.y = element_text(size=18, color = "black"),
        axis.text.x = element_text(size=20, color = col_n10[1], face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=20, color = "black"),
        plot.subtitle = element_text(size=22, color = "black"),
        plot.title = element_text(color="#E58908", size=28, face="bold"))+
  labs(x = "Fungal Sub-Assemblage",title = "Fungal Species Richness",
       y = "Observed OTUs",
       caption = "pairwise wilcox.test, ref.group = Sub-Assemblage F1",
       subtitle = paste("F1 (mean = ",
round(mean(fwc_alpha$Observed[fwc_alpha$Fun_sor_clusters=="F1"]),0),
                        ") vs. all (global mean = ", 
                        round(mean(fwc_alpha$Observed),0),
                        ")",sep = ""))
  # scale_x_discrete(breaks = c(1:10))
  # expand_limits(x= c(-1, 0))+
  # scale_x_discrete(drop=FALSE)
# scale_x_discrete(limits = "F1")
# scale_x_discrete(limits = c("F2","F2","F3","F4","F5",
#                             "F6","F7","F8","F9","F10"))

# ggsave(filename="../figures/sfaf/sfaf2025/obs_Fun_clusters.png", plot = last_plot(),height = 4.8, width=7.5,units = "in", dpi = 400, device = "png")

# ggsave(filename="../figures/sfaf/sfaf2025/obs_Fun_clustersF1.png", plot = last_plot(),height = 4.8, 
#        width=7.95*(183/484)*1.05,
#        units = "in", dpi = 400, device = "png")

ggsave(filename="../figures/sfaf/sfaf2025/obs_Fun_clusters_binwidth925_F1.png", plot = last_plot(),height = 4.8, 
       width=7.95*(183/484)*1.05, # 3.156183
       units = "in", dpi = 400, device = "png")
```




```{r plot-Fun-alpha-nonF1, echo=FALSE, fig.height=3.5, fig.width=5, dpi=300, out.height="70%",out.width="70%", eval=FALSE}
ggviolin(subset(fwc_alpha, fwc_alpha$Fun_sor_clus2=="non-F1"), x = "Fun_sor_clusters", y = "Observed",
          fill = "Fun_sor_clusters", palette = col_n10[2:10],
         alpha = 0.85, trim = FALSE, add = c("dotplot",
                                             "boxplot"
                                           ,"mean_sd","mean"
                                           ),#size=1,
         add.params = list(alpha = 0.6, binwidth = 9.25, size=0.8),
         error.plot = "errorbar" ) +
  geom_hline(yintercept = mean(fwc_alpha$Observed), linetype = 2)+ # add mean hline
  scale_y_continuous(limits = c(0,350))+
  theme(legend.position = "none",
        axis.text.y = element_text(size=18, color = "black"),
        axis.text.x = element_text(size=20, color = col_n10[2:10], face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=20, color = "black"),
        plot.subtitle = element_text(size=22, color = "black"),
        plot.title = element_text(color="#E58908", size=28, face="bold"))+
  labs(x = "Fungal Sub-Assemblage",title = "Fungal Species Richness",
       y = "Observed OTUs",
       caption = "pairwise wilcox.test, ref.group = Sub-Assemblage F1",
       subtitle = paste("F1 (mean = ",
round(mean(fwc_alpha$Observed[fwc_alpha$Fun_sor_clusters=="F1"]),0),
                        ") vs. all (global mean = ", 
                        round(mean(fwc_alpha$Observed),0),
                        ")",sep = ""))
  # scale_x_discrete(breaks = c(1:10))
  # expand_limits(x= c(-1, 0))+
  # scale_x_discrete(drop=FALSE)
# scale_x_discrete(limits = "F1")
# scale_x_discrete(limits = c("F2","F2","F3","F4","F5",
#                             "F6","F7","F8","F9","F10"))

# ggsave(filename="../figures/sfaf/sfaf2025/obs_Fun_clusters.png", plot = last_plot(),height = 4.8, width=7.5,units = "in", dpi = 400, device = "png")

# ggsave(filename="../figures/sfaf/sfaf2025/obs_Fun_clustersF1.png", plot = last_plot(),height = 4.8, 
#        width=7.95*(183/484)*1.05,
#        units = "in", dpi = 400, device = "png")

ggsave(filename="../figures/sfaf/sfaf2025/obs_Fun_clusters_binwidth925_nonF1.png", plot = last_plot(),height = 4.8, 
       width=7.5*1.05,
       units = "in", dpi = 400, device = "png")
```


# beta disp

```{r beta-disp}
# Fun_wc_sorensen_dist<-as.dist(as.matrix(beta.pair(decostand(phylo2vegan_OTU(Fun_wholecommunity), method = "pa"), index.family = "sorensen")$beta.sor))

betadisp_clus_Fun_addC <- betadisper(Fun_wc_sorensen_dist,
                            group = phylo2vegan_sd(Fun_wholecommunity)$Fun_sor_clusters,
                            type = "median", # type = "median" (distance from each sample unit to the spatial median of its group)
                            bias.adjust = T, # adjust for small sample bias in beta div estimates?
                            sqrt.dist = FALSE,
                            add = "lingoes")

betadisp_clus_Fun_addC$distances

```



```{r plot-Fun-betadisp-all, echo=FALSE, eval=FALSE}
betadisp_clus_Fun_addC$distances
betadisp_clus_Fun_addC$group

betadisp_clus_Fun_df <- data.frame("group"=betadisp_clus_Fun_addC$group,
                                   "distances"=betadisp_clus_Fun_addC$distances)


ggviolin(betadisp_clus_Fun_df, x = "group", y = "distances",
          fill = "group", palette = col_n10,
         alpha = 0.85, trim = FALSE, add = c("dotplot",
                                             "boxplot"
                                           ,"mean_sd","mean"
                                           ),#size=1,
         add.params = list(alpha = 0.6, binwidth = .01, size=0.8),
         error.plot = "errorbar" ) +
  # stat_compare_means(aes(label = after_stat(p.signif)),
  #                    method = "wilcox.test", ref.group = "F1", # Add pairwise p-value
  #                    label = "p.signif",hide.ns=T, 
  #                    paired = F, size=3,label.y = 0.9)+
  stat_compare_means(label.y = 0.95,label.x = 1.15)+     # Add global p-value
  geom_hline(yintercept = mean(betadisp_clus_Fun_df$distances), linetype = 2)+ # add mean hline
  scale_y_continuous(limits = c(0.5,0.9))+
  theme(legend.position = "none",
        axis.text.y = element_text(size=18, color = "black"),
        axis.text.x = element_text(size=20, color = col_n10, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=20, color = "black"),
        plot.caption = element_text(size=18, color = "black"),
        plot.subtitle = element_text(size=22, color = "black"),
        plot.title = element_text(color="#E58908", size=28, face="bold"))+
  labs(x = "Fungal Sub-Assemblage",title = "Fungal \u03b2-dispersion (within-group variability)",
       y = "\u03b2-dispersion (within-group\ndistance to spatial median)",
       caption = "pairwise wilcox.test, ref.group = Sub-Assemblage F1",
       subtitle = paste("F1 (mean = ",
round(mean(betadisp_clus_Fun_df$distances[betadisp_clus_Fun_df$group=="F1"]),3),
                        ") vs. all (global mean = ", 
                        round(mean(betadisp_clus_Fun_df$distances),3),
                        ")",sep = ""))

ggsave(filename="../figures/sfaf/sfaf2025/betadisp_Fun_clusters_all.png", plot = last_plot(),height = 4.85, 
       width=((7.5*1.05)+3.156183)+0.1,
       units = "in", dpi = 400, device = "png")
```

```{r plot-Fun-betadisp-F1, echo=FALSE, eval=FALSE}
betadisp_clus_Fun_addC$distances
betadisp_clus_Fun_addC$group

betadisp_clus_Fun_df <- data.frame("group"=betadisp_clus_Fun_addC$group,
                                   "distances"=betadisp_clus_Fun_addC$distances)


ggviolin(subset(betadisp_clus_Fun_df, betadisp_clus_Fun_df$group=="F1"), x = "group", y = "distances",
          fill = "group", palette = col_n10,
         alpha = 0.85, trim = FALSE, add = c("dotplot",
                                             "boxplot"
                                           ,"mean_sd","mean"
                                           ),#size=1,
         add.params = list(alpha = 0.6, binwidth = .01, size=0.8),
         error.plot = "errorbar" ) +
  # stat_compare_means(aes(label = after_stat(p.signif)),
  #                    method = "wilcox.test", ref.group = "F1", # Add pairwise p-value
  #                    label = "p.signif",hide.ns=T, 
  #                    paired = F, size=3,label.y = 0.9)+
  # stat_compare_means(label.y = 0.95,label.x = 1.15)+     # Add global p-value
  geom_hline(yintercept = mean(betadisp_clus_Fun_df$distances), linetype = 2)+ # add mean hline
  scale_y_continuous(limits = c(0.5,0.9))+
  theme(legend.position = "none",
        axis.text.y = element_text(size=18, color = "black"),
        axis.text.x = element_text(size=20, color = col_n10, face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=20, color = "black"),
        # plot.caption = element_text(size=18, color = "black"),
        plot.subtitle = element_text(size=22, color = "black"),
        plot.title = element_text(color="#E58908", size=28, face="bold"))+
  labs(x = "Fungal Sub-Assemblage",title = "Fungal \u03b2-dispersion (within-group variability)",
       y = "\u03b2-dispersion (within-group\ndistance to spatial median)",
       caption = "pairwise wilcox.test, ref.group = Sub-Assemblage F1",
       subtitle = paste("F1 (mean = ",
round(mean(betadisp_clus_Fun_df$distances[betadisp_clus_Fun_df$group=="F1"]),3),
                        ") vs. all (global mean = ", 
                        round(mean(betadisp_clus_Fun_df$distances),3),
                        ")",sep = ""))

ggsave(filename="../figures/sfaf/sfaf2025/betadisp_Fun_clusters_F1.png", plot = last_plot(),height = 4.8, 
       width=(3.156183)+0.4,
       units = "in", dpi = 400, device = "png")
```

```{r plot-Fun-betadisp-nonF1, echo=FALSE, eval=FALSE}
betadisp_clus_Fun_addC$distances
betadisp_clus_Fun_addC$group

betadisp_clus_Fun_df <- data.frame("group"=betadisp_clus_Fun_addC$group,
                                   "distances"=betadisp_clus_Fun_addC$distances)


ggviolin(subset(betadisp_clus_Fun_df, betadisp_clus_Fun_df$group!="F1"), x = "group", y = "distances",
          fill = "group", palette = col_n10[2:10],
         alpha = 0.85, trim = FALSE, add = c("dotplot",
                                             "boxplot"
                                           ,"mean_sd","mean"
                                           ),#size=1,
         add.params = list(alpha = 0.6, binwidth = .01, size=0.8),
         error.plot = "errorbar" ) +
  # stat_compare_means(aes(label = after_stat(p.signif)),
  #                    method = "wilcox.test", ref.group = "F1", # Add pairwise p-value
  #                    label = "p.signif",hide.ns=T, 
  #                    paired = F, size=3,label.y = 0.9)+
  # stat_compare_means(label.y = 0.95,label.x = 1.15)+     # Add global p-value
  geom_hline(yintercept = mean(betadisp_clus_Fun_df$distances), linetype = 2)+ # add mean hline
  scale_y_continuous(limits = c(0.5,0.9))+
  theme(legend.position = "none",
        axis.text.y = element_text(size=18, color = "black"),
        axis.text.x = element_text(size=20, color = col_n10[2:10], face = "bold"),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size=20, color = "black"),
        # plot.caption = element_text(size=18, color = "black"),
        plot.subtitle = element_text(size=22, color = "black"),
        plot.title = element_text(color="#E58908", size=28, face="bold"))+
  labs(x = "Fungal Sub-Assemblage",title = "Fungal \u03b2-dispersion (within-group variability)",
       y = "\u03b2-dispersion (within-group\ndistance to spatial median)",
       caption = "pairwise wilcox.test, ref.group = Sub-Assemblage F1",
       subtitle = paste("F1 (mean = ",
round(mean(betadisp_clus_Fun_df$distances[betadisp_clus_Fun_df$group=="F1"]),3),
                        ") vs. all (global mean = ", 
                        round(mean(betadisp_clus_Fun_df$distances),3),
                        ")",sep = ""))

ggsave(filename="../figures/sfaf/sfaf2025/betadisp_Fun_clusters_nonF1.png", plot = last_plot(),height = 4.8, 
       width=(7.5*1.1),
       units = "in", dpi = 400, device = "png")
```



```{r beta-disp-anova}
perform_anova_betadisper <- function(betadisper_object, grouping) {

  # Perform ANOVA 
  anova_result <- anova(betadisper_object)
  
  tidy_aov <- tidy(anova_result)
  tidy_df <- as.data.frame(tidy_aov)
  
  tidy_df$term[1] <- paste0(grouping)

  sum_squares_regression <- tidy_df$sumsq[1]
  sum_squares_residuals <- tidy_df$sumsq[2]

  R_squared <- sum_squares_regression /
            (sum_squares_regression + sum_squares_residuals)*100

  tidier_df <- data.frame(Term = character(),df = numeric(),
                          SumSq = numeric(),MeanSq = numeric(),
                          F.statistic = numeric(),
                          R_squared = numeric(),p.value = character(),
                          stringsAsFactors = FALSE)
  
  tidier_df[1,"Term"] <- tidy_df$term[1]
  tidier_df[2,"Term"] <- tidy_df$term[2]
  tidier_df[1:2,2] <- (tidy_df$df)[1:2]
  tidier_df$SumSq <- round(tidy_df$sumsq[1:2],2)
  tidier_df$MeanSq <- round(tidy_df$meansq[1:2],3)
  tidier_df$F.statistic <- round(tidy_df$statistic,2)
  tidier_df[1,"R_squared"] <- round(R_squared,2)
  tidier_df[2,"R_squared"] <- 100-(round(R_squared,2))
  tidier_df[1,"P.value"] <- prettyNum(tidy_df$p.value[1],digits=2)

  # Return ANOVA results for the specified taxa_of_interest
  return(tidier_df)
}
library("broom"); packageVersion("broom") #‘1.0.5’
betadisp_clus_Fun.aov<-perform_anova_betadisper(betadisp_clus_Fun_addC, "Fun_sor_clusters")
```


```{r beta-disp}
```








# Sample Map + Convex Hulls

```{r fig-map-hulls, echo=F,fig.width=7, fig.height=5}

# Convex hulls by group
hull_df <- sd %>%
  group_by(Fun_sor_clus2) %>%
  dplyr::slice(chull(Longitude, Latitude))

# Ensure Site is a factor
sd$Site <- as.factor(sd$Site)

# Get centroid positions for labeling
centroids <- sd %>%
  group_by(Site) %>%
  dplyr::summarise(
    Longitude = mean(Longitude),
    Latitude = mean(Latitude)
  )

centroids$Site <- as.factor(centroids$Site)
centroids_HPGCPR <- subset(centroids, Site %in% c("HPG", "CPR"))
centroids_BLMHPGCPR <- subset(centroids, Site %in% c("BLM","HPG", "CPR"))

# centroids[5,] #CPR
# centroids[12,] #HPG
centroids_not_BLMHPGCPR <- as.data.frame(rbind(centroids[2:4,],centroids[6:11,]
                                            ,centroids[13:22,]))

spatial_extent_plot <- ggplot(sd, aes(x = Longitude, y = Latitude, 
                                      color = Fun_sor_clus2, fill = Fun_sor_clus2)) +
  geom_polygon(data = hull_df, aes(group = Fun_sor_clus2), alpha = 0.4, color = NA) +
  geom_point(size = 5, alpha = 0.95) +
  geom_point(data = centroids, aes(x = Longitude, y = Latitude,group = Site), 
             alpha = 0.8, color = "black", size = 7.5, shape = 1,
             show.legend = F, stroke = 1.25, inherit.aes = F) +

  geom_text(data = centroids_not_BLMHPGCPR, aes(x = Longitude, y = Latitude, label = Site), 
            alpha = 1, color = "black", size = 4.5, 
             show.legend = F, nudge_y = 0.6, #nudge_x = 0.15, 
            check_overlap = F, inherit.aes = F) +

  geom_text(data = centroids_BLMHPGCPR, aes(x = Longitude, y = Latitude, label = Site), 
            alpha = 1, color = "black", size = 4.5, 
             show.legend = F, nudge_y = c(0.1,0.5,0.6), nudge_x = c(0.88,-0.6, 0.4), 
            check_overlap = F, inherit.aes = F) +
  
  scale_color_manual(values = c("F1" = "#8E0152", "non-F1" = "#7FBC41")) +
  scale_fill_manual(values = c("F1" = "#8E0152", "non-F1" = "#7FBC41")) +
  theme_minimal(base_size = 15, base_family = "Arial") +
  coord_fixed() +
  labs(
    title = "Site Locations with Convex Hulls",
    subtitle = "Group-specific spatial extent",
    x = "Longitude", y = "Latitude",
     color = "Fungal Subassemblage", fill = "Fungal Subassemblage"
  ) +
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 17),
        plot.title = element_blank(),
        plot.subtitle = element_blank(),
        axis.title = element_blank(),
        plot.background = element_rect(fill = NA, colour = NA))
spatial_extent_plot

# saveRDS(spatial_extent_plot,file = "../figures/sfaf/sfaf2025/spatial_extent_plot_OG.rds")

spatial_extent_plot <- spatial_extent_plot_OG +
  theme(legend.position = "inside",
    legend.position.inside = c(0.9,0.95),
        legend.title = element_blank(),
        legend.text = element_text(size = 17),
        plot.title = element_blank(),
        plot.subtitle = element_blank(),
        axis.title = element_blank(),
        plot.background = element_rect(fill = NA, colour = NA))
spatial_extent_plot

ggsave("../figures/sfaf/sfaf2025/spatial_extent_plot_F1nonF1.png", 
       plot = last_plot(),
       # width = 5, 
       height = 6, units = "in", device = "png", dpi = 400)
```

\
\
\





# explan var

```{r, include=FALSE, eval=FALSE}

# Fun ----
Fun_ev_tab1<-data.frame("Explanatory Variable"=adonis.out.array_Fun_wc.r2$expl_var,
"Sub-assemblage"=adonis.out.array_Fun_wc.r2$comm,"R^2"=adonis.out.array_Fun_wc.r2$R2,"P-value"=adonis.out.array_Fun_wc.r2$`Pr(>F)`, check.names = F)

Fun_ev_tab1$`R^2`[Fun_ev_tab1$`P-value` > 0.05] <- 0
  
Fun_ev_tab2<-subset(Fun_ev_tab1,select = -c(4))
Fun_ev_tab3<-reshape(Fun_ev_tab2, idvar = "Explanatory Variable", timevar = "Sub-assemblage", direction = "wide")

colnames(Fun_ev_tab3) <-c("Explanatory Variable","Whole Community","F1","F3","F4","F5","F6","F7","F8","F9","F10")
rownames(Fun_ev_tab3) <- NULL
max(Fun_ev_tab3[,2:11])
Fun.xc <- seq(min(Fun_ev_tab3[,2:11]), max(Fun_ev_tab3[,2:11]), length.out = 10)
Fun.pal <- seq_gradient_pal("#E5A60C1a", "#E5A60C")(Fun.xc)

# setSpecFun <- function(y){
#   kableExtra::cell_spec(y, "html", background = Fun.pal[cut(y, breaks=Fun.pal, include.lowest = TRUE)])
# }
# apply(Fun_ev_tab3[,2:11], 2, setSpecFun) %>% knitr::kable("html", escape = FALSE,caption = "adonis2 outputs for single-term models",align='rcccccccccc') %>%

Fun_ev_tab3 %>%
  # mutate_if(is.numeric, ~ +(setSpecFun(.))) %>%
  kbl("html", escape = F,caption = "adonis2 outputs for single-term models",align='rcccccccccc') %>%
  pack_rows("Edaphic", 1, 5) %>%
  pack_rows("Climate", 6, 7) %>%
  pack_rows("Plant", 8, 10) %>%
  kable_styling() %>%
  scroll_box(width = "90%", height = "600px")
# Fun ----


# Bac ----
Bac_ev_tab1a<-data.frame("Explanatory Variable"=adonis.out.array_Bac_wc.r2$expl_var,
"Sub-assemblage"=adonis.out.array_Bac_wc.r2$comm,"R^2"=adonis.out.array_Bac_wc.r2$R2,"P-value"=adonis.out.array_Bac_wc.r2$`Pr(>F)`, check.names = F)

Bac_ev_tab1<-subset(Bac_ev_tab1a,`Sub-assemblage` != "Bac_wc_B6")
Bac_ev_tab1$`Sub-assemblage`<-factor(Bac_ev_tab1$`Sub-assemblage`)
Bac_ev_tab1$`R^2`[Bac_ev_tab1$`P-value` > 0.05] <- 0
  
Bac_ev_tab2<-subset(Bac_ev_tab1,select = -c(4))
Bac_ev_tab3<-reshape(Bac_ev_tab2, idvar = "Explanatory Variable", timevar = "Sub-assemblage", direction = "wide")

colnames(Bac_ev_tab3) <-c("Explanatory Variable","Whole Community","B1","B2","B3","B4","B5","B7",
                          "B8","B9","B10","B11","B12","B13","B14")
rownames(Bac_ev_tab3) <- NULL
max(Bac_ev_tab3[,2:15])
Bac.xc <- seq(min(Bac_ev_tab3[,2:11]), max(Bac_ev_tab3[,2:11]), length.out = 10)
Bac.pal <- seq_gradient_pal("#006e901a", "#006e90")(Bac.xc)

# setSpecBac <- function(y){
#   kableExtra::cell_spec(y, "html", background = Bac.pal[cut(y, breaks=Bac.pal, include.lowest = TRUE)])
# }
# apply(Bac_ev_tab3[,2:11], 2, setSpecBac) %>% knitr::kable("html", escape = FALSE,caption = "adonis2 outputs for single-term models",align='rcccccccccc') %>%

Bac_ev_tab1 %>%
  kbl("html", escape = F,caption = "adonis2 outputs for single-term models",align='rcccccccccc') %>%
  pack_rows("Edaphic", 1, 5) %>%
  pack_rows("Climate", 6, 7) %>%
  pack_rows("Plant", 8, 10) %>%
  kable_styling() %>%
  scroll_box(width = "90%", height = "600px")

Bac_ev_tab3 %>%
  # mutate_if(is.numeric, ~ +(setSpecBac(.))) %>%
  kbl("html", escape = F,caption = "adonis2 outputs for single-term models",align='rcccccccccc') %>%
  pack_rows("Edaphic", 1, 5) %>%
  pack_rows("Climate", 6, 7) %>%
  pack_rows("Plant", 8, 10) %>%
  kable_styling() %>%
  scroll_box(width = "90%", height = "600px")
# Bac ----
```